### 2016-01-30 abstract.tex
Initial check in, FSE format

When are two program executions similar?  In most contexts, the answer is that they are similar when they satisfy or violate correctness properties for the same reasons.  This paper proposes a novel metric for measuring the distance between two failing test cases, and automated debugging methods based on that metric.  The metric is based on a causal concept of distance:  executions are similar to the degree that changes in the program itself, called mutants, cause similar changes in the correctness of the executions.  Specifically, if two failing test cases become successful due to the same mutant, they are more likely to be due to the same fault.  We evaluate the mutant-based distance metric by applying it to problems of fault identification and automated debugging assistance for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods.

### 2016-01-30 advantages.tex
Initial check in, FSE format

\subsection{Key Novel Features of Localization}

One key point is that the localization information provided by
repairing mutants is different in kind from that typically provided by
fault localization, particularly statistically-driven localizations.
Namely, the relationship between a mutant and the underlying fault is
\emph{always} causal.  There are not ``accidental'' repairs; at
minimum, a mutant modifies program behavior such that the faulty code
no longer causes a fault, which it would have under other
circumstances.  This means that unlike traditional localizations,
where reading ``incorrect'' suggestions (e.g., ones that are not an
actual faulty location) is often of no value, every mutant that
repairs a program provides some information about the fault, however
obscure.  Some repairs, of course, are of little interest because they repair too much.

A second point is that unlike the use of program modifications in
automatic patch generation, our technique is not defeated by weak
oracles.  Of course, if an oracle is too weak to detect failures at
all, it cannot be applied, but given the ability to detect failures,
it is not critical that a suite be able to check correctness.  Our
repairing mutants are not meant to serve as actual program patches,
but 1) as features to drive a fuzzer taming algorithm and 2) as error
explanations and fault localization to ease the debugging of complex
compiler faults.

### 2016-01-30 conclusions.tex
Initial check in, FSE format

\section{Conclusions and Future Work}
\label{conc}

When are two failing test cases due to the same fault?  This paper
demonstrates that
``when they are repaired by the same program mutants'' is a useful answer to
this question.  The problem of causality in executions is hard, and
has long been tied to a notion of similarity of
executions \cite{NearNeighbor,GroceError}.  By using program changes
(produced by mutants) as
causes to determine distance, it is possible to improve on efforts
to identify the set of faults in a large set of redundant failures
\cite{PLDI13} and discover the source code locations of faults
\cite{MUSE,multilingual}.

As future work, we plan to apply the mutation-based metric to other
fuzzer taming and fault localization problems.  The highly varying
results for even the best fault localization methods in our
experiments demonstrate that further advances are required before
compiler debugging can consistently be made less onerous through automated
assistance.  One idea is to combine the metric approach and the ideas
in MUSE and MUSEUM, in some way, to exploit passing tests and perhaps
additional coverage information.

%We also plan to explore other applications of mutant-based metrics.
Many software engineering techniques rely on measuring distance
between program executions.  Not all of these concern failing
executions.  For example, Zhang et al. \cite{issta14} show that FPF based on a simple
Hamming distance over branches covered can significantly improve the
effectiveness of seeded symbolic execution
\cite{Zesti,PersonSeed,BugRedux}.  Mutation response is likely too
expensive to use for this purpose without modification, but sampling
mutants to refine distance when distances plateau
may be practical.  Mutants may also be useful in determining
execution peers \cite{Sumner2011}, for example to help operators of
spacecraft find similar behaviors to telemetry downloads in testbed
history \cite{KlausRajeev,scriptstospecs}.

### 2016-01-30 experiments.tex
Initial check in, FSE format

\section{Experimental Results}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.02\textwidth]{jscurve}
        \caption{Discovery curves for SpiderMonkey 1.6 faults}
        \label{jscurves}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.02\textwidth]{gcccurve}
        \caption{Discovery curves for GCC 4.3.0 wrong-code faults}
        \label{gcccurves}
    \end{subfigure}
    \caption{Discovery curves compared to best curves from PLDI 2013}
\end{figure*}

%\begin{figure*}
%\centering
%\includegraphics[width=2\columnwidth]{jscurve}
%\caption{Discovery curves for SpiderMonkey 1.6 faults}
%\label{jscurves}
%\end{figure*}

%\begin{figure*}
%\centering
%\includegraphics[width=2\columnwidth]{gcccurve}
%\caption{Discovery curves for GCC 4.3.0 wrong-code faults}
%\label{gcccurves}
%\end{figure*}

%\subsection{General Experimental Framework}

All experiments in this paper are based on subjects used in the only previous study of compiler fuzzer taming \cite{PLDI13}.  Of the three data sets examined in that paper, this paper considers two:  faults and tests cases for SpiderMonkey 1.6, Mozilla's JavaScript engine, with tests generated by {\tt jsfunfuzz} \cite{jsfunfuzz}, and \emph{wrong code} faults and test cases for GCC 4.3.0, with tests generated by Csmith \cite{csmith}.  GCC 4.3.0 crash faults were essentially perfectly localized by previous approaches.  Faults that produce a crash are also likely to be easier to debug than purely semantic problems, such as many of the SpiderMonkey faults and all GCC wrong code faults.

All mutants were produced using the C mutation tool written by Jamie Andrews \cite{mutant}, used in many previous experiments in mutation analysis and software testing.  Andrews' tool applies only four operators: statement deletion, conditional negation, operator replacement, and constant replacement, chosen as a small set that still produces good results for C code.  The tool was chosen to keep the number of mutants as small as possible to control the cost of mutation analysis.  Coverage was analyzed with gcov.   All experiments were performed on a MacBook Pro with 16GB RAM and dual-core 3.1GHz Intel Core i7 (not using the second core); GCC executed on a VirtualBox-hosted Ubuntu 11.04.  For GCC, some test cases from the 2013 PLDI paper no longer failed, presumably due to unknown differences in execution environment, OS, or memory layout. Discarding these reduced the set of test cases from 1,275 to 1,117 and the number of distinct faults to 27 rather than 35.  For SpiderMonkey, all 1,749 test cases from the PLDI 2013 data failed in the new environment, representing 28 distinct faults.

These data sets, though similar in that both compilers are written in C, provide some interesting variance for testing our hypotheses.  The oracle for SpiderMonkey executions is the set of checks built in to {\tt jsfunfuzz} plus the requirement that the execution not crash.   If the {\tt jsfunfuzz} produced test case executes to the end and prints {\tt TEST SUCCESSFULLY COMPLETED} it is considered successful.  This is only a moderately strong oracle, and allows serious deviations in behavior from both the JavaScript language specification and normal SpiderMonkey behavior, while constraining other some of execution fairly tightly, such as {\tt eval} round-trips.  For GCC, the oracle is a differential check on a hash code:  failure was defined as producing a executable with the -O3 flag that, when executed, produced a different checksum than code compiled by either of GCC 4.9.3 or clang 7.0.0, both using -O0.  A repairing mutant must enable GCC 4.3.0 to actually compile code correctly, which is a very strict correctness property in one sense, making coincidental correctness \cite{CCT} highly unlikely --- the only weakness is that there is no check that optimizations were applied, or that compile time was not excessive.

We only show results for the first 50 tests in the ranking, and computed areas under curves for the same limit, since it seems very unlikely that a user will examine many more than 50 tests, especially given the decreasing slope of the discovery curves.  In practice, after 50 tests, fixing a few faults and then re-running tests and FPF seems the most likely recourse, and we confirmed that after removing random subsets of faults, our metrics still outperform the previous best.  We applied X-means \cite{xmeans} in an attempt to use clustering, but, as with previous results \cite{PLDI13} clustering did not compare well with FPF, and the runtime was much worse than for FPF-based methods.  This tends to confirm earlier work's \cite{PLDI13} conclusion that clustering (at least based on X-means) simply does not work well in a setting with the kind of extreme disparity in instances of faults found in fuzzer taming.  A further difficulty for our setting is that X-means and most off-the-shelf clustering tools take vectors, not a distance metric.  X-means is using the raw repair data, not weighted by the fact that 0-0 agreement is much less informative than other possible matchings (recall that a Euclidean over 0-1 vectors is just the square root of Hamming distance).

\subsection{Mozilla SpiderMonkey 1.6 Results}

There were 96,828 SpiderMonkey mutants, based on 69,634 lines of code in C and header files.  Of these mutants, 12,666 were covered by some failing test case.  Of these mutants, 10,525 survived a basic set of SpiderMonkey quick tests \cite{icst2014}.  Of these mutants, 1,326 (12.6\%) repaired at least one test case.  Figure \ref{jscurves} shows the discovery curves for the mutant repair metric compared to the ideal discovery curve and the best curve from previous work using FPF, which used a normalized Levenshtein distance \cite{lev} over the failure output and test case text \cite{PLDI13}.  The APFD (Average Percent Faults Detected) values shown in the graphs are based on the measure  introduced by Rothermel et al. to evaluate test case prioritization methods \cite{APFD}.  This is a somewhat better summary of results than the raw curve areas used in previous work \cite{PLDI13}.  APFD, as the name suggests, shows the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's overall area to an ideal curve, with interpolation\footnote{The details of APFD calculation are slightly involved, and the interpolation and perfect curve are not exact fits for fuzzer taming; nonetheless the basic method summarizes curves fairly well and is standard in the testing literature for similar problems \cite{issta14}.}.

APFD results are useful summaries, but we suggest that the curve itself is also worth examining, since (1) a very good early curve is perhaps most important to developers and (2) long sequences of tests with no new faults may discourage users more than an overall less effective but steadily climbing curve with few ``plateaus'' as we call these uninformative sequences of tests.  The SpiderMonkey curve climbs very rapidly in the early portion, with perfect discovery for the first 12 faults. The largest plateau is 5 tests without a new fault, ignoring the long period after the 31st test.  In fact, a plateau at the end of the curve is not problematic.  Assuming that few users will examine more than 50 test cases without fixing some faults, the user may give up after seeing 10 or more tests without a new fault based on the mutant repair curve, and in fact lose nothing by doing so.  The PLDI 2013 curve, in contrast, has a plateau of size 17 after the 17th fault.  If we assume a user stops examining tests after a size 10 or greater plateau, the user will only see 17 faults using the PLDI 2013 metric, vs. 19 with our metric.  Dropping the stopping criteria to a plateau of size 5 produces the same results.

Over all mutants, and all pairs of test cases repaired by the same mutant (so the same pair may count many times, if many mutants repair both test cases), the probability of being due to the same fault was 42.77\%, and the probability of being due to the same fault if two test cases disagreed on a repair (the mutant repaired one test case but not the other) was only 20.33\%.  The baseline rate for same-fault for test pairs was 33.64\%.  Just knowing that two test cases have \emph{one} shared repairing mutant makes it 1.27x more likely that they are the same fault, and knowing they differ for one mutants makes it 0.6 times as likely they are due to the same fault.  As suspected, while matching non-repair provided some evidence, that evidence was quite weak:  only a 34.14\% chance of matching fault.

An additional measure of effectiveness is to consider how effective a metric is in producing matched nearest neighbors.  That is, how often is the nearest neighbor of a failure (that is not due to a singleton fault --- a fault detected by only one test case) due to the same fault?  For reduced \cite{DD,PLDI13,
CReduce} test cases, we assume that for a perfect metric, the nearest neighbor should almost always share the same fault, since there is little or no extraneous semantic content to each test beyond the cause of failure.  For the SpiderMonkey failures, 96.3\% of non-singleton failures matched their nearest neighbor(s).  For mutants that repaired any failures, the mean and median number of test cases repaired were 120.6 and 4.0, respectively.  Most mutants repaired a small number of tests, but a few mutants repaired a very large number of tests.  A few mutants repaired \emph{all} SpiderMonkey faults; obviously these mutants were not actual fault fixes of any kind, but effectively disabled some mechanism {\tt jsfunfuzz} used to detect failure. The mean and median homogeneity for repairing mutants (percent of failures repaired corresponding to the most common fault repaired) were 77.96\% and 86.36\%, respectively.

\subsubsection{Fault Localization}

Table \ref{jstable} shows a comparison of three fault localization methods for the SpiderMonkey faults.  The first column shows, for each fault, identified by the commit number of the code change that first ``fixed'' the failing tests associated with the fault, its FPF ranking: this is \emph{not} a fault localization ranking, just a way to see that the localization success seems to be somewhat independent of the ease of distinguishing the failures for a fault.  The next four columns show localization rankings.  A dash for a column means that localization did not rank any faulty statements, or assigned all faulty statements suspiciousness of 0.0.  The four rankings are: the {\bf Coverage} localization, the {\bf Repair} localization, the MUSE \cite{MUSE} localization, and the MUSEUM \cite{multilingual} localization, which uses the same formula as MUSE, but works better for multiple faults because it uses only one failure.  The MUSE/MUSEUM formulas normally make use of information from passing tests as well: when mutating a statement makes a passing test case fail, it makes the statement \emph{less} suspicious, by a weighted amount. The weight assigned to information from passing tests in our setting would likely be low (due to the ratios of repairs to mutation kills).  In a limited sense, information from passing tests is already incorporated in our results.  Throwing out all mutants that are killed by any passing test as potential repairs/localizations ensures that the rankings of all statements that are in MUSE/MUSEUM rankings are correct, relative to each other.  By definition, passing tests have no influence on the suspiciousness of these mutants.  However, there may be other mutants that 1) repair a failure and 2) are killed by some test case: these could, if enough repaired a faulty statement, improve MUSE/MUSEUM results.  We reject such mutants in part to keep costs low, and in part because we think that the causal information contained in such mutants, that repair a failure but also break some passing test, is problematic.  They seem likely to be less useful as explanations of the causes of a failure, since they do not impact the program semantics in a way that is only known to be beneficial, and could potentially even mislead a user if used as explanations.

Discussion with the MUSE/MUSEUM authors confirms  that adding information for passing tests, while costly, would likely improve the MUSE and MUSEUM results, given the weakness of our oracles, particularly for SpiderMonkey, and should be considered essential for ideal application of their approach.  Because the cost of recording the full mutant analysis matrix for all passing tests (rather than considering a mutant killed as soon as one test fails, as we do) is very high, and we would like to produce a comparison over a fixed computational cost, we give results for MUSE/MUSEUM over failing tests only, cautioning that we are not sure what the impact of this choice is.  For similar reasons of keeping costs low, we also used the mutation operators of Andrews vs. the more extensive Proteum \cite{Proteum} operators used in MUSE/MUSEUM's evaluations.  Even with these restrictions, running all mutants on passing tests took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.

Results are reported as absolute ranks, rather than more sophisticated \cite{MUSE} localization evaluations because our localizations do not assign suspiciousness scores, only rankings, and in accord with the proposal of Parnin and Orso that only localizations ranking a fault in the top few statements may be useful to users \cite{AutoHelp}.  We expect users to examine coverage changes and mutants.  Blind, unaided pointers to portions of compiler code without more semantic information about why the statement is suspicious seems unlikely to work well, based on our experience with complex compiler faults, and the results of Parnin and Orso \cite{AutoHelp}.

\begin{table}
\centering
{\scriptsize
\begin{tabular}{|c||c||c|c|c|c|}
\hline
& & \multicolumn{4}{|c|}{Localization Rank} \\
\hline
Fault & FPF & Coverage & Repair & MUSE & MUSEUM \\
\hline
880 & 1 & 2,255 & - & - & - \\
95 & 6 & 720 & - & 173 & -\\
1294 & 9 & 190 & 1 & 22 & 4 \\
60 & 10 & - & - & - & -\\
1543 & 15 & 18 & 19 & 118 & 26\\
1172 & 19 & 46 & 28 & 87 & 23\\
1561 & 106 & - & - & - & - \\
115 & 192 & 6 & 3 & 133 & 5 \\
\hline
\end{tabular}
}
\caption{Fault localization results for SpiderMonkey}
%\vspace{-0.2in}
\label{jstable}
\end{table}

\begin{figure}
{\scriptsize
\begin{code}
{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):}
...
<             vp[1] = OBJECT\_TO\_JSVAL(thisp);
<         \} else \{
<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);
<         \}
...
<                   a->avail = (jsuword)sp;
<               \}
...
---
>         RESTORE\_SP(fp);
...
{\bf Mutant information:}
Failure repaired by 148 mutants.
...
\#3: delete statement mutant of jsinterp.c:944:

            vp[1] = OBJECT\_TO\_JSVAL(thisp);
        \} else \{
/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/
        \}

{\bf Coverage change information:}
...
\#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:
                   a->avail = (jsuword)sp; /* ADDED */
\end{code}
}
\caption{Patch and explanation for fault 115}
%\vspace{-0.18in}
\label{fig:explain}
\end{figure}

The first thing to note about results is that Table \ref{jstable} only includes 8 of the 28 SpiderMonkey faults.  This is due to a problem discussed in the paper introducing this data set:  producing ground truth patches for a large application, in the sense of finding a minimal, clearly fault-fixing code change that can be back-ported to the original code is difficult \cite{PLDI13}.  While we believe the 28 faults identified in the PLDI 2013 data set are correct, it is very difficult to produce a valid patch of version 1.6 that captures the fix for most of these faults.  In some cases the final commit that caused tests to stop failing appeared to only be the end of a complex series of changes that converged on a correct fix.  In other cases, the code was modified so extensively before the fix that identifying ``the incorrect part'' of the original code seemed to owe more to guesswork than certainty.  The evaluation of localization therefore only examines the 8 faults for which we could be reasonably certain that the patch to 1.6 was correct and characterized the fault in question accurately.  The smallest patch required modification of two statements, and was thus larger than any mutants applied.  In no case was the patch equivalent to one of the mutants.

No method performs extremely well --- for half the faults, no methods produced what we consider a useful localization.   Compiler faults are hard to debug, and any help is useful, but it seems unlikely developers will really benefit from a localization when it does not rank at least one faulty statement in (at most) the first 20 to 30 statements.  By this measure, MUSE only provides a helpful localization once, and performs worst of the methods.  {\bf Coverage} performs very badly for three faults, and does not manage to be useful by our definition for another, but in the remaining two cases provides a good localization, in one case the best.  {\bf Repair} and MUSEUM both perform well, with {\bf Repair} slightly better.

\subsubsection{Error Explanation}

Figure \ref{fig:explain} shows part of the patch for SpiderMonkey fault 115, and the {\bf Coverage} and {\bf Repair} outputs that successfully localize the fault (the 6th coverage change output, and the 3rd mutant output), to give some idea of the information our methods present.  For {\bf Coverage}, it is surprising that the sixth highest ranked change only appeared in 5 of 148 repairs. This change ranked high out of thousands of coverage changes because it was an unusual coverage change, appearing for only 10 repairs in the entire SpiderMonkey analysis.

The precise form an explanation based on our techniques should take is not clear.  One possibility, suggested by the results here, is a method focused on repairs, but using common coverage changes.  For each repairing mutant, in rank order, the highest ranking coverage changes for that mutant could also be shown, to give an idea of unusual (and potentially localizing) changes induced by repair.

\subsection{GCC 4.3.0 Wrong Code Results}

There were 377,679 GCC mutants, based on 424,186 lines of code in C and header files. Of these mutants, 73,016 were covered by some failing test case.  Of these mutants, 41,385 survived a GCC bootstrap build (used to produce our compiled versions for repair checks), compiled hello world correctly, and passed a GCC test suite.  Of these mutants, 3,232 (7.8\%) repaired at least one test case.  Figure \ref{gcccurves} shows discovery curves for the mutant repair metric vs. the best PLDI 2013 curve, Euclidean distance over function coverage vectors \cite{PLDI13}.  GCC's APFD improvement is larger than that for SpiderMonkey, but its early curve is worse compared to the PLDI 2013 curve.  Using the hypothetical model where a user stops examining tests after a plateau of size 10, the user of the mutant-based taming will see 18 distinct faults by examining 25 tests, and the user of the function coverage based taming will see 20 faults after examining 45 tests.  If a user gives up after a plateau of size 5, our approach again lets a user examine 18 faults over 25 tests.  Function coverage provides only 16 faults over 31 tests.

For GCC wrong code faults, the baseline chance two test cases shared an underlying fault was 38.97\%.  Knowing that two test cases shared a repairing fault raised the chance to 68.5\%, and knowing they disagreed on a fault lowered it to only 19.01\%.  The respective increase and decrease in probability of matching fault compared to baseline was thus 1.75 times greater for matching repairs, and less than 0.5 the chance of being the same fault, given one mismatched repair.  Knowing two test cases were both not fixed by a mutant again provided marginal evidence of sharing a fault: 39.2\% chance of matching faults.
For 99.3\% (all but 8) of the 1,090 non-singleton fault failures in GCC wrong code, the closest test case(s) by the repair metric matched fault.  The best previous reported FPF metric, function coverage vectors, had a matching rate of 92.2\%.   The mean and median numbers of repaired tests per mutant (for mutants that repaired any tests at all) were 18.4 and 3.0, respectively.  A few mutants fixed a very large number of tests, up to 1,050\footnote{These all appear to be turning off all optimizations --- forcing GCC to run in -O0 mode.}, but most mutants fixed only a few tests, to a greater extent than was true with SpiderMonkey.    The mean and median homogeneity (\% of repaired tests corresponding to the most commonly repaired fault) were 79.4\% and 100\%, respectively --- over half of all repairing mutants only repaired instances of the same fault.

\subsubsection{Fault Localization}

\begin{table}
\centering
{\scriptsize
\begin{tabular}{|c||c||c|c|c|c|}
\hline
& & \multicolumn{4}{|c|}{Localization Rank} \\
\hline
Fault & FPF & Coverage & Repair & MUSE & MUSEUM \\
\hline
139094 & 1 & 44,561 & - & 157 & -\\
133004 & 2 & 75 & 1 & 99 & 14\\
158782 & 4 & - & - & 146 & -\\
156496 & 5 & 3,067 & - & - & -\\
143677 & 7 & 186 & - & 109 & -\\
158555 & 8 & 490 & - & - & -\\
134321 & 10 & - & - & - & -\\
133940 & 14 & 260 & 4 & 158 & 18\\
141195 & 15 & 465 & - & - & -\\
156795 & 17 & 19 & - & - & -\\
139709 & 19 & 43,903 & 10 & 167 & 6\\
155698 & 22 & 85 & 1 & 172 & 195\\
138646 & 24 & 2 & - & - & -\\
140795 & 25 & 302 & 34 & 170 & 125\\
136501 & 44 & 223 & 5 & 159 & 31\\
134322 & 45 & 639 & - & - & -\\
\hline
\end{tabular}
}
\caption{Fault localization results for GCC}
\label{gcctable}
\end{table}

Table \ref{gcctable} shows localization results for GCC 4.3.0 wrong code faults.  For the same reasons as with SpiderMonkey, only some faults were deemed to have a strong enough ground truth patch for localization evaluation.   MUSE performed poorly, with no useful (by the standard of having at least one faulty statement in the top 30) localizations.  {\bf Coverage} produced one very good localization, and was the only method to localize that fault at all, but this was its only useful localization.  MUSEUM provided two respectable localizations, but no very high quality localizations.  Only {\bf Repair} performed extremely well:  while it only localized three faults, those localizations were all extremely useful.

\subsection{Discussion}

In one sense, the discovery curve improvements here are practically useful, but not spectacular.  The SpiderMonkey curve has an APFD only slightly more than 2\% better than the best result from previous FPF efforts.  The GCC wrong code curve improves APFD by 6.4\%.  However, this is comparing the best curves from many curves computed in previous work to a single curve using a universal metric.  The comparison ignores the human effort required to find a good metric for any given program using just the techniques of our previous fuzzer taming work.

For fault localization and error explanation, the results show possible improvement on state-of-the-art approaches, in the absence of complete information from mutant analysis over passing tests (which can be very expensive to obtain).  A practical impact of the results reported is that we suggest users of our techniques examine only the first few (at most 10, and we propose as few as 5) mutants and coverage-changing statements, and ignore localization if none of these results are helpful.  This is analogous to our suggestion that a user abandon the FPF curve if 10 test cases in a row fail to reveal any new faults, as a heuristic.  We suspect the information in the mutants/coverage changes is probably helpful in some cases where they do not localize a faulty line, but this is simply based on our highly incomplete understanding of the faults and patches in question, and not a solidly established claim.  The expertise of compiler developers would be needed to confirm or reject this belief.  Our suggestion that mutations themselves provide interesting error explanation is also applicable to MUSE and MUSEUM.  There may be some potential for confusing explanations, however, if repairs include mutants that also cause some failures, as in the standard MUSE/MUSEUM approach.

One obvious question is:  why does {\bf Repair} perform best in this setting?  It is impossible to be sure, but one  possibility is power to distinguish repairs. If each failure was only repaired by one mutant, MUSEUM and {\bf Repair} would both rank that mutant maximally, and we speculate it would usually be faulty.  But such cases are rare:  for SpiderMonkey the mean/median number of mutants repairing each failure were 90.8 and 78, respectively, and for GCC the mean and median were 54 and 43.  If there are many repairs for a failure, how can the most important ones be distinguished?  MUSEUM relies on a large set of mutation operators, and assumes that faulty statements will have multiple repairs, compared to non-faulty statements, or that faulty statements will be less likely to cause passing tests to fail when mutated.  Restricting ourselves to cases where {\bf Repair} and MUSEUM both provided a localization, we know that some faulty statement was in a repairing mutant, but that passing tests could not distinguish this mutant  from other repairs.  Many of the faulty statement repairs (as expected for optimizing compilers) were statement deletions and conditional negations.  For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.  {\bf Repair} can still distinguish such mutants by relying on the onion-ring structure:  when a repair also repairs test cases that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults.  The hypothesis behind {\bf Repair} is that \emph{repairs specific to the fault causing a failure are most likely to modify code involved in the fault itself}.  This seems to be true, when such a repair exists.  If any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.  MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.  Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this appears unlikely due to the nature of the repairs.

Finally, in contrast to many previously reported results in fault localization, typically over much simpler programs and faults than subtle compiler semantics bugs, all methods frequently failed to provide a useful localization at all.  In one sense, {\bf Coverage} is the most intriguing method presented here.  While it did not perform nearly as well as {\bf Repair}, it did manage, in many cases, to rank some faulty line highly relative to the extremely large number of statements executed in each failing GCC run (on the order of 50K-100K+ statements).  In future work, we plan to experiment with different methods for weighting frequency of coverage, perhaps using methods from machine learning, or of pruning changed coverage that is not likely to be relevant.  For example, perhaps the {\bf Repair} approach can be applied, and coverage changes present in very distant failures can be removed, though the argument is less clear, since a coverage change isn't directly associated with the onion-ring structure, and does not directly repair a set of failures.  There is much to be done if automated debugging is ever to make the lives of compiler developers easier in most cases \cite{AutoHelp}.

\subsubsection{Mutation Analysis Costs}

Mutation analysis is not cheap.  The experiments in this paper were run over a long period of time.  However, there are numerous ways to reduce costs.  First, the primary cost is compiling the mutated versions of the SUT.  Running each test case under each mutated version of the compiler is cheap: each execution requires around 0.05 seconds for SpiderMonkey, and 0.12 seconds for GCC, on average.  These executions can obviously be performed in parallel, on low-cost virtual machines in the cloud.  The cost is further reduced by a few more facts: most failures do not cover most mutants, and vice versa, so the total number of repair checks needed is far less than the product of mutants and failures.  The practical approach to fuzzer taming for a large, ongoing project is likely to maintain a set of ``hot'' pre-compiled mutants.  These can be pruned at will by running them on passing test cases, and removing mutants that fail on any test case.  Keeping hot mutants for a heavily modified program could be costly (since each mutant must be recompiled, if not re-tested, for each modification to the code), but for stable releases this is a simple approach to keep costs low.

Second, and most critically, most program mutants never need to be considered as repairs, because we only consider mutants that cause no passing test case to fail.  MUSE and MUSEUM use kills by passing tests to help rank such tests as repairs, which is finer grained but more expensive.  As we show, MUSE and MUSEUM \emph{can}  be used without passing tests or these mutants, and MUSEUM still produces good results even with a weak oracle and few operators.  MUSE, MUSEUM, and our approach all run analysis without coverage instrumentation; we re-run repaired tests under gcov when using the {\bf Coverage} localization.

Mutation cost reduction techniques are also applicable.  E.g., trivial compiler equivalence (TCE) \cite{TCE} can reject equivalent and redundant mutants, removing almost 30\% of mutants for benchmark subjects. Other techniques, such as using mutant schemas to avoid having to compile and store each mutant, are also applicable --- almost all of the techniques characterized by Offutt and Untch \cite{offutt2001mutation} as \emph{do faster} approaches apply.   Whether sampling mutants \cite{RahulISSRE} is applicable is less clear.  Even if the FPF curve remains the same, a critical mutant for localization might be omitted.

\subsection{Threats to Validity}

The largest threat to validity is that this paper's conclusions rely on only two data sets, used in a previous paper.  The fault ground truths are possibly imperfect, due to the complex change history, and it was not possible to produce good patches for all faults, limiting our analysis of localization and explanation.  Generalizing to all compilers, or to software systems that are not compilers, would be unwarranted.  However, for compilers, we believe the underlying arguments supporting the probability of discovering many meaningful repairing mutants should hold in the general case.  There is no particular reason to believe that Csmith and {\tt jsfunfuzz} produce unusual random test cases, and both have often been used to detect faults in compilers other than the two considered in this paper.  For other kinds of software, more experiments are definitely warranted.  It is also possible that faults in the framework used for experiments introduced errors into the data.  In order to check for these problems, we will make the raw data for a re-analysis available on request.

Second, the comparison with MUSE and MUSEUM is limited.  MUSE is not really intended for multiple faults.  With our oracles, MUSE and MUSEUM may need the full results from passing tests, and some mutant we omit might be both highly ranked and a good localization.  A full comparison would include a complete analysis of passing tests also, but would likely require far more than 10x the computation time.  Finally, these experiments use Andrews' mutation operators, while MUSE/MUSEUM experiments use Proteum \cite{Proteum}, which has a much larger set of mutation operators, and their results indicate that the large set of operators is important.  However, running all Proteum GCC mutants (especially over all passing tests) would likely be very expensive.

### 2016-01-30 introduction.tex
Initial check in, FSE format

\section{Introduction}

Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in compilers.  Unfortunately, the blessing of many failing test cases can quickly become a curse for a compiler developer.  Simply discovering which test cases in a large set of failures represent distinct faults (and possibly already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.  For simple crashes and assertion violations, bucketing failures is sometimes easy.  When a failure is detected by differential testing \cite{Differential}, where two compilers or optimization levels disagree, the problem is so hard that it may turn potential users away from automated test generation \cite{PLDI13}.  Even once a fault is identified, fixing the problem can be very difficult.  There is reason to suspect, due to complexity of intermediate representation, interaction of a very large number of optimizations, subtle aspects of language semantics, and other issues, that debugging compiler faults may usually be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
We propose the use of mutants \cite{mutant}, small syntactic changes to source code, to address some of the problems of debugging based on large sets of automatically generated test cases.  The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault. Repairs can also provide  information for understanding and debugging discovered faults. This approach also suggests a possible metric for comparing any two program executions, not just failing test cases.

\subsection{Fuzzer Taming and Debugging Assistance}

Automated test generators tend to produce many failures, which correspond to far fewer faults.  The distribution of faults in a set of failures tends to follow a strong power-law curve, where a few faults account for most failures, and many or even most faults have only 1 or 2 associated failures (even in a set of 1,000 or more test cases) \cite{PLDI13}.  Looking at all the failures to find  distinct faults is impractical.  It is also impractical to apply the iterative approach:  first, fix one fault; then re-run all test cases, knowing that the remaining failures are due to different, unfixed faults, repeating until all failures go away.  First, there may be some faults that are very difficult to fix --- a large compiler project typically has many open bugs.  As of 11/20/2015, LLVM clang had 89 bugs assigned but not resolved, some dating to 2008.  Knowing if a randomly generated test case is an instance of a known failure can be difficult, and fixing these problems is unlikely.    A research group with a novel fault-detection technique \cite{csmith,ISSTA12,ZhendongPLDI} may dump dozens of new faults on a project at once, producing more open problems.

Chen et al. defined the \emph{fuzzer taming problem}: ``Given a potentially large collection of test cases, each of which triggers a bug, rank them in such a way that test cases triggering distinct bugs are early in the list.'' \cite{PLDI13}.  The solution Chen et al. proposed to this problem was to use the Furthest-Point-First \cite{Gonzalez} (FPF) algorithm.  An FPF ranking of test cases requires a distance metric $d$ on test cases, and ranks test cases so that dissimilar tests appear earlier.  The hypothesis of Chen et al. was that dissimilar tests, by a well-chosen metric, will also fail due to different faults.
FPF is a greedy algorithm that proceeds by repeatedly adding the item with the maximum minimum distance to all previously ranked items. Given an initial seed item $r_0$, a set
$S$ of items to rank, and a distance metric $d$, FPF computes $r_i$ as $s \in S: \forall s' \in S: min_{ j < i}(d(s,r_j)) \geq min_{j < i}(d(s',r_j))$.

Fuzzer taming results using FPF were promising, but limited.  Rather than proposing a universal metric for use in fuzzer taming, Chen et al. offered many possible metrics (based on ideas in the literature, such as edit distances \cite{lev} or coverage spectra \cite{RepsSpectra} methods), none of which performed notably well across all three benchmark data sets, and some of which were not even applicable to all of the data sets.  For the two most difficult benchmarks, none of the metrics performed extremely well. FPF-based approaches remain in need of a universal but effective metric, particularly for the critical first 50 test cases examined \cite{PLDI13}.

One reason fuzzer taming is needed is that debugging compiler faults is, as noted, difficult.  If developers were able to fix faults quickly once they were identified, it would not be as important to have very good fault identification.  Many duplicate failures could be removed by fixing the underlying faults.  Ideally, a technique able to provide effective fuzzer taming should be able to translate its underlying rationale into some form of automated debugging assistance.   A variety of automated debugging methods \cite{NearNeighbor,GroceError} have been based on distance, so there is some reason to expect that a high-quality metric might also contribute to that problem.

\subsection{Comparing Executions by Response to Mutation}

The key insight of this paper is a hypothesis with numerous possible applications, including a universal fuzzer taming metric that can also provide assistance in debugging.

\begin{quote}
{\bf Hypothesis 1:} If two test cases that fail can be made to succeed by the same syntactic modification to the original program $P$, this provides evidence that they fail due to the same fault.
\end{quote}

The core idea is hardly controversial:  most developers determine if they have fixed all faults in a set of failing test cases by re-running the tests after they fix at least one fault.  As a method for fuzzer taming or debugging assistance, however, this is useless.  Fortunately, developer-provided patches are not the only source of syntactic changes to a program.  Mutation analysis \cite{demillo1978hints,budd1980theoretical} applies large numbers of syntactic changes to a program, usually to evaluate a test suite \cite{mutant,justmutants}.  Some recent work has also considered these changes as a source of fixes or localizations for program faults \cite{achour,multilingual,MUSE,DebroyMutant}.  Our experience, as well as empirical data \cite{GopinathMutants}, suggests that very few compiler faults are due to the kind of simple syntactic changes found in mutants.  However, this does not matter for {\bf Hypothesis 1};  what matters is that failures, which correspond to faults, can be made to succeed.  When a test case $t$ fails for program $P$ but succeeds for mutant $m$ of $P$ we say that $m$ \emph{repairs} $t$.  In most cases this does not mean that $m$ is the inverse of a fault, but that $m$ somehow avoids the consequence of a fault.  It is not important that repairs actually fix anything, if all we seek is a way to decide whether failures are due to the same fault.

Hypothesis 1 is in fact a consequence of a more general (and difficult to investigate) hypothesis.  What is interesting about a program execution?  For many purposes, the only interesting properties of an execution are those that in some way relate to the correctness properties of the program, assuming we have a good set of correctness properties.  These may include assertions, differential testing \cite{Differential,ICSEDiff}, timing constraints, temporal logic formulas, or any other checkable properties.  What, then, makes two executions similar?  Our proposal is to say that:

\begin{quote}
{\bf Hypothesis 2:} Two executions of $P$ are similar to the degree that changes to $P$ produce the same changes in correctness of the executions.
\end{quote}

This leads to the idea of comparing executions by their response to mutants.   In the presence of multiple correctness properties, any two executions can be compared, not just two failures, since a mutant can cause a failure to violate additional or different properties.
%Where there are multiple, independent, correctness properties, this even allows us to compare failing and successful executions, since a mutant can cause a failure to violate a different property.
{\bf Hypothesis 2} at first may suggest that this idea subsumes past notions of similarity, such as spectra of code coverage, since if two executions do not cover a statement, they both cannot alter correctness if that statement is altered.  The metrics are in fact simply incomparable:  for example, differences in coverage that never have an effect on correctness cannot contribute to similarity in our approach; with weak correctness properties most executions are very similar.  This is similar to the idea of \emph{checked coverage}:  code that cannot impact correctness is not interesting \cite{CheckCov}.  Hypothesis 2 also suggests an additional hypothesis related to hypothesis 1.  This is {\bf Hypothesis 1a:} If two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.

\subsection{The Onion-Ring and Causality}

These hypothesis do not claim that being repaired by the same mutant perfectly predicts fault equivalence.  To understand why that seems highly unlikely, consider why a mutant of a compiler might cause a test case to stop failing.  Many, likely most, compiler faults are due to incorrect optimizations.  Turning an optimization off can cause a program to compile correctly.  Optimizations are generally guarded by conditional switches --- in some cases, so that the optimization can be turned off by a user, in other cases when a property of the code in question prevents the optimization.  If two failures become successes due to turning off an optimization, it is certainly some level of evidence they are the same fault, compared to the baseline probability.  Intuitively, the evidence provided seems much stronger than the fact that, e.g., the two test cases execute a shared line of code \cite{RepsSpectra}.  However, optimization guards range from guarding very specific fragments of code to guarding entire modules with many functions and related sub-optimizations.  If a mutant prevents a very specific optimization (consisting of a few lines of code) from executing, and repairs two tests, that is powerful evidence they are related.  If a mutant turns off a large number of optimizations, it is weak evidence.  Characterizing the granularity of program mutants is likely difficult, but fortunately it is not necessary.

\begin{figure}
\centering
\includegraphics[width=0.6\columnwidth]{onionring}
\caption{\scriptsize{The onion-ring model of repairs, showing all repairs for the center-most test case, and all other test cases repaired by those mutants.  Circles represent repairing mutants, and diamonds repaired test cases.  Diamond pattern symbolizes fault.  For the innermost ring of repairing mutant(s), all test cases are due to the same fault.  As rings expand to include more failures, the set of faults repaired grows (as does distance).}}
\label{fig:onion}
\end{figure}

Consider the structure of mutants repairing a failure due to some optimization.  If many mutants repair this failure, by turning off the optimization or a whole family of optimizations, some will likely repair failures due to different faults.  Repairs naturally produce an \emph{onion-ring} structure (Figure \ref{fig:onion}), where a failure is repaired by (1) some mutants that are very specific to its fault (the inner rings) and repair few failures, as well as other mutants (2) that repair more and more faults, up to a hypothetical mutant (3) that turns off all optimization in the compiler (e.g., by effectively adding {\tt -O0}), the outer ring of the onion.  This does not introduce confusion, because the more repairs two tests have in common, the more likely they are to be due to the same fault, and the more they disagree on, the less likely they are to be due to the same fault.  While especially intuitive as a consequence of disabling compiler optimizations, it seems likely this structure appears for repairs in many programs.  Disabling an optimization is a code change that (usually) does not cause any test case that only concerns semantics, not speed, to fail, but can avoid some failures.  For some oracles, many changes may fall into this classification:  for example, if the only oracle is to detect crashes, a fault can be avoided by aborting the program early.  Turning off a data-structure invariant checker is another possible false repair that might cover many different faults.  In all these cases, so long as there are other, more interesting and more specific ways to repair some faults, a distance metric can still distinguish faults effectively.

The idea of measuring the distance between failures (or other test cases) by their response to mutants is appealing.  The most important aspect of most applications of distances between executions is some aspect of causality \cite{LewisCause,ZellerBook}.  Since at least the work of Hume \cite{Hume1}, causality has been understood as answering some question of the form ``what would have made a difference in this effect?''  A program mutant is a very direct way to answer this question.  The problem with determining causes of events in computer programs is that, as in most problems of causality, too many things can make a difference, most of which are not interesting.  Program mutants are much more likely to be interesting causes for events in a program than changes to test cases, at least for fault identification and debugging purposes, because the practical use of knowing a cause is to change the cause and avoid or produce an effect.  Debugging by changing test cases is not useful; debugging by altering the program is precisely how we seek to use the causes we discover.  Using mutants as causes in order to measure similarity essentially inverts the ideas of David Lewis \cite{LewisCause,LewisCount}, where similarity is used to determine causality.

\subsection{Contributions}

The contributions of this paper are: hypotheses about how to measure distance between program executions, and practical proposals for using such a metric in fuzzer taming and in novel fault localization and explanation methods.  Evaluating the effectiveness of these methods also provides experimental evidence as to the validity of our hypotheses.  Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over previous results \cite{PLDI13}, in multiple ways.  Our preliminary localization approaches compare favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual}, for the mutation analysis budget in our experiments.

### 2016-01-30 localizations.tex
Initial check in, FSE format

\section{Localization and Explanation}

Automated fault localization \cite{FaultSurvey,Jones2002} provides users with
information about the likely location of a fault in a program's source
code.  Error explanation \cite{GroceError} also provides information
about the causal structure of a fault --- how it can be avoided, what
triggers it --- a kind of ``fault story.''  Parnin and Orso recently
suggested that fault localization methods are not actually providing a
great deal of assistance to real users in debugging \cite{AutoHelp}.
Automated fault localization has not been widely adopted in industry.

Parnin and Orso note that most popular methods are
statistical and tend to simply provide an ordered list of statements
to examine.  These statements are not guaranteed to be essentially
related to the fault --- with many techniques, the statements are,
too often, code that executes as a consequence of the fault, code
accidentally executed with the fault, and so forth.  Second, most
evaluations of fault localization have been performed over programs
where it is not clear that debugging is extremely hard, vs. the time
spent applying automated localization and establishing effective
(automated) testing to support automated localization.  Compiler
debugging, on the other hand, especially for subtle optimization-based
problems that produce incorrect code (vs. simple crashes) is generally
known to be difficult.
Researchers reporting subtle compiler bugs often observe the time from
reporting a fault to its correction, even once the fault is assigned,
to be lengthy, and the resulting patches are sometimes thousands of
lines \cite{PLDI13}.  Important compilers, whether JITs where security is paramount
or C compilers used to compile core systems code, also tend to already
be the
subjects of extensive automated testing \cite{jsfunfuzz,csmith}.

By definition, when a mutant repairs a test case, it changes the
behavior of the test.  In some cases, this behavioral change is
limited to changes in program state and data values.  However, the
vast majority of mutant repairs also modify the code coverage of a
test; in our experiments, there were no repairs that did not change
coverage.  If a fault can be avoided by, for example, avoiding a
problematic optimization in a compiler, this naturally leads to
changed coverage.  If the fault is due to a bad conditional, where a
program path should include additional code, the repair will often
change the execution to include the omitted code.  As a localization
for a fault, then, we can examine the source statements whose
  coverage changed most often in repairs.  Early experiments with this
  approach showed that a useful coverage change localization must  account for the frequency of a
change.  Some coverage changes are extremely common across all
failures, faults, and repairing mutants,
and therefore likely not related to any specific fault.

\begin{quote}
The {\bf Coverage} localization for a failing test case $f$ in the set
of failures $F$ ranks
statements by the function $r$ (for rank) where:
%$$U = s : \forall m . \forall t \in F . c(m,t,s) = 1$$
$$C(s) = \sum_{t \in F} (\sum_{m \in R(t)} c(m,t,s))$$
%\[
$$r(s,f) = \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s)$$
%\begin{cases}
% \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s) & s \not\in U\\
%   0 & s \in U
%\end{cases}
%\]
$c(m,t,s)$ is 1 if coverage of $s$ changes for test $t$
with mutant $m$, and 0 otherwise.  $R(f)$ is the set of mutants repairing $f$.
\end{quote}

That is, the ranking is based on the number of times a statement's
coverage changes its value (covered or not covered) from the original
execution of $f$ in executions of mutants that repair $f$.  Changes are weighted by their inverse frequency over all
failures and all mutants.
Higher values for $r$ result in a higher ranking (more likely to
be faulty).  
%Computing $U$ accounts for the fact that in our
%experiments, a few statements were added or removed for every single
%r

The repairing mutants themselves can also be used for localization and
explanation.  While few will be equivalent to actual fixes, they
will sometimes overlap the incorrect code, and will often be closely
related to incorrect code, semantically.  Since many failures are
fixed by a large number of mutants, it is important to rank the
mutants to produce a localization.  Based on the onion-ring model proposed above, this is simple: each mutant is ranked as
a localization/explanation based on the \emph{most distant failure it also
  repairs.}  If a given mutant only repairs failures that have similar
repair vectors to the test case being localized, it will be highly
ranked.  If it repairs even one very dissimilar failure, then it will
be considered a poor localization.

\begin{quote}
The {\bf Repair} localization for a failing test case $f$ ranks
statements by the function $r$ where:
%$$dmax(m,f) = max_{t : m \in R(t)} d(t,f)$$
\[
r(m,f) = 
\begin{cases}
\frac{1}{1 + max_{t \in F: m \in R(t)} d(t,f)} & m \in R(f)\\
0 & m \not\in R(f)\\
\end{cases}
\]
%$$\text{leastfar} = m \in R(f) : \neg\exists m' : \text{far}(m',f) <
%\text{far}(m,f)$$
\end{quote}

This ranks a mutant $m$ (and associated statement) by the inverse of the distance
from $f$ to the most distant statement also repaired by $m$, adding 1
so that if $m$ only repairs failures at distance 0 $r$ is well defined
(and maximal).
If $m$ does not repair $f$, it is not ranked at all.

One important aspect of these approaches is that they 
can provide qualitative information for debugging, known as an \emph{error
  explanation} \cite{GroceError}, in addition to localization.  For any
interesting line of code in the coverage-change-based localization,
the user can investigate a particular instance of the change,
by selecting a single mutant and stepping through the execution of the
failing test case in a debugger to the point of change.  The user can also
browse through all mutants that produce this coverage change for the test
case.  Mutants that repair a failure obviously provide some insight
into the nature of the fault, and, we hypothesize, more information the
closer they are to the center of the onion-ring.

Another possible advantage over most statistical approaches to
localization, even some recently proposed, more accurate, mutant-based
approaches, is that our methods are not potentially confused in
settings with multiple faults.  Each localization/explanation is based
on a single failing test case, which in most cases fails due to only one
fault.  In fact, other faults help both methods rank localizations.

### 2016-01-30 metrics.tex
Initial check in, FSE format

\section{A Mutant-Based Distance Metric}

For every failure, the response to mutants defines a bit-vector, with
length equal to the number of mutants that repaired any test case,
where a 1 bit indicates that the mutant in question repairs the test
case.

The simplest distance metric for two bit-vectors of equal length is a
Hamming distance.  However, raw Hamming distance has some unfortunate
properties.  Consider two test cases, $t_1$ and $t_2$, both repaired
by 100 mutants.  If these two test cases share 90 common repairing
mutants, and disagree for 10, they have a raw Hamming
distance of 20.  This makes them ``less similar'' than two test cases,
$t_1'$ and $t_2'$, where $t_1'$ is repaired by 5 mutants, $t_2'$ is
repaired by 5 mutants, and \emph{none of those mutants overlap.}   Hamming distance for the first pair of tests is twice as
large, despite the fact we have no evidence $t_1'$ and
$t_2'$ are due to the same fault, in contrast to strong evidence for $t_1$ and $t_2$.  A Jaccard distance, used in some fault
localization algorithms \cite{Liu06}, fits the situation more
effectively.  For bitvectors $u$ and $v$ of length $n$:
%\vspace{-0.08in}
\[d(u,v) = 
\begin{cases}
\frac{\sum_{i=0}^{n-1} 1\ \text{if}\ u_i \neq v_i\ \text{else}\ 0}{\sum_{i=0}^{n-1} 1
\  \text{if}\ u_i \neq 0 \vee v_i \neq 0\ \text{else}\ 0} & \exists i:u_i\neq 0 \vee v_i\neq 0\\
0 & otherwise
\end{cases}
\]

This metric matches our intuitions about the above comparisons.  It
essentially formalizes the intuition of the onion-ring model, where
more matching repairs indicates a much higher probability two failures
are due to the same fault, even if not all repairs match.

Given such a metric, the FPF \cite{Gonzalez} approach to fuzzer taming \cite{PLDI13}
begins by selecting an arbitrary test case.   FPF then proceeds by selecting the test
case, of all unranked tests, that has the largest minimum distance
from all already ranked test cases, repeating this until all tests are
ranked.  In other words, FPF always chooses a test case that is as dissimilar as
possible from all previously chosen test cases, based on the closest
chosen test case (rather than an average distance or other measure).  FPF is boundedly close
to optimal under certain assumptions \cite{Gonzalez}.

More complex metrics are possible, also combining information such as
coverage and language features or test case tokens \cite{PLDI13}.
However, mixtures of metrics (combining different features) other than
Levenshteins \cite{lev} over test case + output never performed well
compared to more focused metrics in earlier work \cite{PLDI13}.
Therefore, in order to evaluate the contribution of our repair-based
metric we compare it, alone, to the previous best metrics.

In Section \ref{conc} we briefly discuss possible uses of $d$
requiring comparison of both failing and passing test cases. The only
change required is to use three-valued vectors with a value for each
property and mutant, where 1 indicates a failing property succeeds (as
in our $d$) and -1 indicates a succeeding property fails.

### 2016-01-30 relatedwork.tex
Initial check in, FSE format

\section{Related Work}
\label{sec:related}

%There are four primary areas of related work.  First, fuzzer taming and related problems of determining the set of faults from a large set of redundant failures have been investigated in a few papers.  Second, some recent efforts to localize faults have begun to incorporate the use of program mutants, and the broader problem of fault localization has been very extensively investigated.  Program mutants have also been used in efforts to produce automatic repairs for program faults.  Finally, our work fits into the general framework of efforts to define distances between program executions.

The most closely related work is that of Chen et. al \cite{PLDI13}.  The present paper uses two of the same data sets, applies the FPF \cite{Gonzalez} algorithm, and has the same empirical focus of determining how to identify faults for complex compilers.  The primary difference is that the distance metrics used with FPF in the previous work were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.  This paper expands on Chen et al. by putting the FPF approach in the context of a much richer causal comparison of program executions that applies in theory to any executions, not just failing test cases.  Prior to the FPF-based work, Francis et al. \cite{Podgurski04}, Podgurski et al. \cite{Podgurski03} and others \cite{Liu06,Liblit05},used clustering to attempt to solve similar problems in identifying distinct faults, on simpler human-generated data sets.  The work of Chen et. al \cite{PLDI13} provides an in-depth comparison of clustering and ranking, and the advantages of ranking (as well as some possible limitations).  Jones et al. discuss general issues in debugging multiple faults efficiently \cite{Jones07}.

Fault localization \cite{FaultSurvey,NearNeighbor,Liblit03,Liblit05,Cleve05,Jones2002,Jones05,Jones07,GroceError,ChakiLev,Liu06,SPIN03,Santelices:ICSE:2009,Abreu:2006:PRDC,xuan2014test,DD}, even specifically for compilers \cite{Whalley94}, is a long-standing field of research.  Recently, fault localization researchers have used program mutants to improve statistical \cite{Jones2002} fault localization techniques \cite{MUSE,DebroyMutant,FasterMutant,multilingual,Metallaxis}.
%The core of this work has been the belief that faulty program locations are (1) more likely to change a failing test to successful when mutated and (2) less likely to change a successful test case to faulty when mutated.  This clearly connects to our general notion that how two test cases respond to a set of mutants is a good measure of their semantic similarity.
The work of Hong et al. \cite{multilingual} on MUSEUM and Moon et al. \cite{MUSE} on MUSE provides a good summary of other work along these lines, and is perhaps most similar to ours in assumptions, in particular ``Conjecture 1'' that a test case that used to fail on a program is more likely to pass on a mutation of a faulty statement.  While we suggest some localizations as a welcome and interesting side effect of our distance metric and FPF analysis, our primary goal is the discovery of all faults. Much fault localization work is largely (in many cases exclusively) based on single-fault scenarios; e.g., the MUSE \cite{MUSE} evaluation includes only 2 of 14 scenarios with multiple faults, and those scenarios include only a few faults.   The statistical approaches may owe some of their superiority in experiments over other methods to the use of mostly single-fault evaluations \cite{Jones05}.  By focusing on a single failing test case (for us, selected by fuzzer taming) as the locus of analysis, our work, like some early efforts \cite{NearNeighbor,Cleve05,GroceError} may be more suited for situations involving many faults.  MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single fault.

Using program modifications or mutants broadly defined to repair programs has been a popular topic of recent work as well \cite{GenProg,AutoRep,WeiFix}.  These \emph{generate-and-validate} approaches to patching aim to fix programs by generating a large variety of potential fixes, then using test cases to prune out invalid fixes.  Qi et al. \cite{achour} discuss some serious limitations of early work on this topic, but also demonstrate that effective patching is possible using a restricted search space that focuses on removing functionality.

Some approaches to localization (and all approaches to patching) assume the possibility of a mutant \emph{actually} fixing a fault.  In our experiments, we found this to be highly unlikely -- most of our patches were far larger than the largest higher-order-mutants \cite{jia2009higher} anyone usually considers applying, and none were equivalent to a repair.  Previous work has shown that, across a large body of open source programs in many languages, most fault fixes were larger in size than even low-order higher-order mutants \cite{GopinathMutants}.
%We speculate that fixes for compilers may be likely to be even more complex than the average, given that compilers are particularly complex software systems.  
%Our primary assumption is simply that similar executions respond similarly to mutants, and in particular executions due to the same fault can be ``repaired'' by similar program changes.  These changes are likely \emph{related} to the actual fault, but are unlikely to be the actual fault, for reasons of comparative complexity of mutants and real fixes.

The use of distance metrics in software engineering includes some of the localization and fault identification efforts discussed above.  Vangala et al. proposed using distance to cluster test cases to improve diversity and eliminate duplicates \cite{VangalaDist}.  Adaptive random testing uses distances (usually between inputs) to choose tests \cite{Chen,ARTChen,ISSTAART}, and Artzi et al. guided test generation for fault localization using path constraint metrics \cite{ArtziDirected}.  Sumner et al. evaluated approaches to generating execution peers for debugging and understanding \cite{Sumner2011} using tree edit distances (and evaluated other distances).  Xin et al. \cite{Xin2008} examined methods for indexing program executions, useful in all methods involving execution alignments.  To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra} (that is, using counts of coverage of code entities).    Our work differs in using a metric that is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.

### 2016-02-10 abstract.tex
minor fixes

@@
 This paper proposes a novel metric for measuring the distance between two failing test cases, and automated debugging methods based on that metric.
-The metric is based on a causal concept of distance:  executions are similar to the degree that changes in the program itself, called mutants, cause similar changes in the correctness of the executions.
+The metric is based on a causal concept of distance:  executions are similar to the degree that changes in the program itself, introduced using mutation operators, cause similar changes in the correctness of the executions.
 Specifically, if two failing test cases become successful due to the same mutant, they are more likely to be due to the same fault.

### 2016-02-10 conclusions.tex
minor fixes

@@
 expensive to use for this purpose without modification, but sampling
-mutants to refine distance when distances plateau
+mutants to refine distance when coarser metrics plateau
 may be practical.  Mutants may also be useful in determining

### 2016-02-10 experiments.tex
minor fixes

@@
 The GCC wrong code curve improves APFD by 6.4\%.
-However, this is comparing the best curves from many curves computed in previous work to a single curve using a universal metric.
-The comparison ignores the human effort required to find a good metric for any given program using just the techniques of our previous fuzzer taming work.
+However, the comparison is with the very best curve chosen after running more than 16 different metrics.
+In practice, users do not know ground truth to rank competing curves, so this is not a practical approach to fuzzer taming.
+The best methods for different subjects in previous work varied widely, even in such peculiar ways as less-fine-grained coverage providing better results in some cases, worse in other cases.
+In practice, previous work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique, and included methods such as comparing output signatures that are inapplicable to many settings.
+This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work.
 
@@
 A practical impact of the results reported is that we suggest users of our techniques examine only the first few (at most 10, and we propose as few as 5) mutants and coverage-changing statements, and ignore localization if none of these results are helpful.
-This is analogous to our suggestion that a user abandon the FPF curve if 10 test cases in a row fail to reveal any new faults, as a heuristic.
+This is analogous to our suggestion that a user abandon the FPF curve if 5-10 test cases in a row fail to reveal any new faults, as a heuristic.
 We suspect the information in the mutants/coverage changes is probably helpful in some cases where they do not localize a faulty line, but this is simply based on our highly incomplete understanding of the faults and patches in question, and not a solidly established claim.
@@
 MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.
-Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this appears unlikely due to the nature of the repairs.
+Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this is unlikely due to the nature of the repairs.
 

### 2016-02-10 relatedwork.tex
minor fixes

@@
 \cite{Podgurski04}, Podgurski et al.
-\cite{Podgurski03} and others \cite{Liu06,Liblit05},used clustering to attempt to solve similar problems in identifying distinct faults, on simpler human-generated data sets.
+\cite{Podgurski03} and others \cite{Liu06,Liblit05} used clustering to attempt to solve similar problems in identifying distinct faults, on simpler human-generated data sets.
 The work of Chen et.
@@
 By focusing on a single failing test case (for us, selected by fuzzer taming) as the locus of analysis, our work, like some early efforts \cite{NearNeighbor,Cleve05,GroceError} may be more suited for situations involving many faults.
-MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single fault.
+MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single failure.
 
@@
 \cite{Xin2008} examined methods for indexing program executions, useful in all methods involving execution alignments.
-To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra} (that is, using counts of coverage of code entities).
-Our work differs in using a metric that is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.
+To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra} (using counts of coverage of code entities), while our metric is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.

### 2016-02-17 experiments.tex
aiming to FSE

@@
 A further difficulty for our setting is that X-means and most off-the-shelf clustering tools take vectors, not a distance metric.
-X-means is using the raw repair data, not weighted by the fact that 0-0 agreement is much less informative than other possible matchings (recall that a Euclidean over 0-1 vectors is just the square root of Hamming distance).
+X-means is using the raw repair data, not weighted by the fact that 0-0 agreement is much less informative than other possible matchings (a Euclidean over 0-1 vectors is just the square root of Hamming distance).
 
@@
 This is a somewhat better summary of results than the raw curve areas used in previous work \cite{PLDI13}.
-APFD, as the name suggests, shows the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's overall area to an ideal curve, with interpolation\footnote{The details of APFD calculation are slightly involved, and the interpolation and perfect curve are not exact fits for fuzzer taming; nonetheless the basic method summarizes curves fairly well and is standard in the testing literature for similar problems \cite{issta14}.}.
+APFD, as the name suggests, shows the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's area to an ideal curve, with interpolation\footnote{The details of APFD calculation are slightly involved, and the interpolation and perfect curve are not exact fits for fuzzer taming; nonetheless the basic method summarizes curves fairly well and is standard in the testing literature for similar problems \cite{issta14}.}.
 
-APFD results are useful summaries, but we suggest that the curve itself is also worth examining, since (1) a very good early curve is perhaps most important to developers and (2) long sequences of tests with no new faults may discourage users more than an overall less effective but steadily climbing curve with few ``plateaus'' as we call these uninformative sequences of tests.
+APFD results are useful summaries, but the curve itself is also worth examining, since (1) a very good early curve is perhaps most important to developers and (2) long sequences of tests with no new faults may discourage users more than an overall less effective but steadily climbing curve with few ``plateaus'' as we call these uninformative sequences of tests.
 The SpiderMonkey curve climbs very rapidly in the early portion, with perfect discovery for the first 12 faults.
@@
 19 with our metric.
-Dropping the stopping criteria to a plateau of size 5 produces the same results.
+Stopping at a plateau of size 5 produces the same results.
 
@@
 Most mutants repaired a small number of tests, but a few mutants repaired a very large number of tests.
-A few mutants repaired \emph{all} SpiderMonkey faults; obviously these mutants were not actual fault fixes of any kind, but effectively disabled some mechanism {\tt jsfunfuzz} used to detect failure.
-The mean and median homogeneity for repairing mutants (percent of failures repaired corresponding to the most common fault repaired) were 77.96\% and 86.36\%, respectively.
+A few mutants repaired \emph{all} SpiderMonkey faults; obviously these mutants were not actual fault fixes, but effectively disabled some mechanism {\tt jsfunfuzz} used to detect failure.
+The mean and median homogeneity for repairing mutants (\% of failures repaired corresponding to the most common fault repaired) were 77.96\% and 86.36\%, respectively.
 
@@
 No method performs extremely well --- for half the faults, no methods produced what we consider a useful localization.
-Compiler faults are hard to debug, and any help is useful, but it seems unlikely developers will really benefit from a localization when it does not rank at least one faulty statement in (at most) the first 20 to 30 statements.
+Compiler faults are hard to debug, and any help is useful, but it seems unlikely developers will really benefit from a localization when it does not rank at least one faulty statement in the first 20 to 30 statements.
 By this measure, MUSE only provides a helpful localization once, and performs worst of the methods.
-{\bf Coverage} performs very badly for three faults, and does not manage to be useful by our definition for another, but in the remaining two cases provides a good localization, in one case the best.
+{\bf Coverage} performs very badly for three faults, does not manage to be useful by our definition for another, and in the remaining two cases provides a good localization, in one case the best.
 {\bf Repair} and MUSEUM both perform well, with {\bf Repair} slightly better.
@@
 For {\bf Coverage}, it is surprising that the sixth highest ranked change only appeared in 5 of 148 repairs.
-This change ranked high out of thousands of coverage changes because it was an unusual coverage change, appearing for only 10 repairs in the entire SpiderMonkey analysis.
+This change ranked high out of thousands of changes because it was unusual, appearing for only 10 repairs in the entire SpiderMonkey analysis.
 
@@
 Finally, these experiments use Andrews' mutation operators, while MUSE/MUSEUM experiments use Proteum \cite{Proteum}, which has a much larger set of mutation operators, and their results indicate that the large set of operators is important.
-However, running all Proteum GCC mutants (especially over all passing tests) would likely be very expensive.
+However, running all Proteum GCC mutants over all passing tests would likely be prohibitively expensive.

### 2016-02-17 experiments.tex
fits now

@@
 For GCC, the oracle is a differential check on a hash code:  failure was defined as producing a executable with the -O3 flag that, when executed, produced a different checksum than code compiled by either of GCC 4.9.3 or clang 7.0.0, both using -O0.
-A repairing mutant must enable GCC 4.3.0 to actually compile code correctly, which is a very strict correctness property in one sense, making coincidental correctness \cite{CCT} highly unlikely --- the only weakness is that there is no check that optimizations were applied, or that compile time was not excessive.
+A repairing mutant must enable GCC 4.3.0 to actually compile code correctly, which is a very strict correctness property, making coincidental correctness \cite{CCT} highly unlikely.
+The only weakness is that there is no check on speed of or time to produce compiled code.
 
@@
 The mean and median numbers of repaired tests per mutant (for mutants that repaired any tests at all) were 18.4 and 3.0, respectively.
-A few mutants fixed a very large number of tests, up to 1,050\footnote{These all appear to be turning off all optimizations --- forcing GCC to run in -O0 mode.}, but most mutants fixed only a few tests, to a greater extent than was true with SpiderMonkey.
-The mean and median homogeneity (\% of repaired tests corresponding to the most commonly repaired fault) were 79.4\% and 100\%, respectively --- over half of all repairing mutants only repaired instances of the same fault.
+A few mutants fixed a very large number of tests, up to 1,050.
+These all appear to be turning off all optimizations --- running in -O0 mode.
+Most mutants fixed only a few tests, to a greater extent than was true with SpiderMonkey.
+The mean and median homogeneity (\% of repaired tests corresponding to most common repaired fault) were 79.4\% and 100\%, respectively.
 
@@
 The SpiderMonkey curve has an APFD only slightly more than 2\% better than the best result from previous FPF efforts.
-The GCC wrong code curve improves APFD by 6.4\%.
+For GCC, APFD improves by 6.4\%.
 However, the comparison is with the very best curve chosen after running more than 16 different metrics.
@@
 
-One obvious question is:  why does {\bf Repair} perform best in this setting?
+ Why does {\bf Repair} perform best in this setting?
 It is impossible to be sure, but one  possibility is power to distinguish repairs.
@@
 Even if the FPF curve remains the same, a critical mutant for localization might be omitted.
+We performed some limited experiments, and found that for a few sample sizes ranging from 13\% of mutants to 50\% of mutants or using only some mutation operators (with the exception of operator-replacement-only), APFD was always better than PLDI 13 best.
+However, the improvement in some cases was \emph{much} smaller than for the full set, and varied considerably (in some cases it was up to 2\% better than our full results).
+Fault localization results were considerably worse, however (with relative rankings preserved) for both our techniques and MUSE/MUSEUM.
 

### 2016-02-17 localizations.tex
fits now

@@
 browse through all mutants that produce this coverage change for the test
-case.  Mutants that repair a failure obviously provide some insight
+case.  Mutants that repair a failure provide insight
 into the nature of the fault, and, we hypothesize, more information the

### 2016-02-17 metrics.tex
fits now

@@
 where a 1 bit indicates that the mutant in question repairs the test
-case.
-
-The simplest distance metric for two bit-vectors of equal length is a
-Hamming distance.  However, raw Hamming distance has some unfortunate
-properties.  Consider two test cases, $t_1$ and $t_2$, both repaired
+case.  Hamming distance between vectors is not suitable for our purposes.  Consider two test cases, $t_1$ and $t_2$, both repaired
 by 100 mutants.  If these two test cases share 90 common repairing
-mutants, and disagree for 10, they have a raw Hamming
+mutants, and disagree for 10, they have a Hamming
 distance of 20.  This makes them ``less similar'' than two test cases,
@@
 $t_2'$ are due to the same fault, in contrast to strong evidence for $t_1$ and $t_2$.  A Jaccard distance, used in some fault
-localization algorithms \cite{Liu06}, fits the situation more
-effectively.  For bitvectors $u$ and $v$ of length $n$:
+localization algorithms \cite{Liu06}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
 %\vspace{-0.08in}
@@
 ranked.  In other words, FPF always chooses a test case that is as dissimilar as
-possible from all previously chosen test cases, based on the closest
-chosen test case (rather than an average distance or other measure).  FPF is boundedly close
+possible from all previously chosen test cases, based on the \emph{closest
+already chosen} test case.  FPF is boundedly close
 to optimal under certain assumptions \cite{Gonzalez}.
@@
 requiring comparison of both failing and passing test cases. The only
-change required is to use three-valued vectors with a value for each
-property and mutant, where 1 indicates a failing property succeeds (as
-in our $d$) and -1 indicates a succeeding property fails.
+change required is to use three-valued vectors, where 1 indicates a failing property succeeds and -1 indicates a succeeding property fails.

### 2016-02-20 abstract.tex
new abstract

@@
-When are two program executions similar?
-In most contexts, the answer is that they are similar when they satisfy or violate correctness properties for the same reasons.
-This paper proposes a novel metric for measuring the distance between two failing test cases, and automated debugging methods based on that metric.
+Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and used in many proposed testing and debugging algorithms.
+This paper proposes a novel metric for measuring distance between executions, and specializes it to the most commonly needed case, failing test cases.
 The metric is based on a causal concept of distance: executions are similar to the degree that changes in the program itself, introduced using mutation operators, cause similar changes in the correctness of the executions.
-Specifically, if two failing test cases become successful due to the same mutant, they are more likely to be due to the same fault.
-We evaluate the mutant-based distance metric by applying it to problems of fault identification and automated debugging assistance for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods.
+Specifically, if two failing test cases (for the original program) become successful due to the same mutant, they are more likely to be due to the same fault.
+This metric both improves an existing method for identifying test cases failing due to unique faults and enables novel fault localization approaches.
+We evaluate our methods using more than 50 faults and 2,500 test cases for two complex, widely-used, real-world compilers, and demonstrate improvements over state-of-the-art methods.

### 2016-02-20 abstract.tex
abstract intro changes

@@
-Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and used in many proposed testing and debugging algorithms.
+Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and useful in many test generation and debugging algorithms.
 This paper proposes a novel metric for measuring distance between executions, and specializes it to the most commonly needed case, failing test cases.

### 2016-02-20 introduction.tex
abstract intro changes

@@
 
-Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in compilers.
+Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers.
 Unfortunately, the blessing of many failing test cases can quickly become a curse for a compiler developer.

### 2016-02-20 introduction.tex
abstract intro changes

--- 2016-01-30 introduction.tex
@@
 
-Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in compilers.
-Unfortunately, the blessing of many failing test cases can quickly become a curse for a compiler developer.
-Simply discovering which test cases in a large set of failures represent distinct faults (and possibly already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.
+Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}  Unfortunately, the blessing of many failing test cases can become a curse for a developer.
+Simply discovering which test cases in a large set of failures represent distinct faults (or already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.
 For simple crashes and assertion violations, bucketing failures is sometimes easy.
@@
 Even once a fault is identified, fixing the problem can be very difficult.
-There is reason to suspect, due to complexity of intermediate representation, interaction of a very large number of optimizations, subtle aspects of language semantics, and other issues, that debugging compiler faults may usually be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
-We propose the use of mutants \cite{mutant}, small syntactic changes to source code, to address some of the problems of debugging based on large sets of automatically generated test cases.
+Due to complexity of intermediate representation, interaction of optimizations, subtle language semantics, and other issues, debugging compiler faults may often be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
+We propose the use of program mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.
 The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault.

### 2016-02-21 experiments.tex
another small fix

@@
 In practice, users do not know ground truth to rank competing curves, so this is not a practical approach to fuzzer taming.
-The best methods for different subjects in previous work varied widely, even in such peculiar ways as less-fine-grained coverage providing better results in some cases, worse in other cases.
-In practice, previous work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique, and included methods such as comparing output signatures that are inapplicable to many settings.
+The best methods for different subjects in previous work varied widely, even in such difficult-to-understand ways as less-fine-grained coverage providing better results in some cases, but worse in other cases.
+In practice, previous work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique.
+It also included methods such as comparing output signatures that are inapplicable to differential properties.
 This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work.

### 2016-02-21 experiments.tex
edited

@@
 Just knowing that two test cases have \emph{one} shared repairing mutant makes it 1.27x more likely that they are the same fault, and knowing they differ for one mutants makes it 0.6 times as likely they are due to the same fault.
-As suspected, while matching non-repair provided some evidence, that evidence was quite weak:  only a 34.14\% chance of matching fault.
+Matching non-repair provides very weak evidence:  only a 34.14\% chance of matching fault.
 
@@
 {\bf Coverage} produced one very good localization, and was the only method to localize that fault at all, but this was its only useful localization.
-MUSEUM provided two respectable localizations, but no very high quality localizations.
-Only {\bf Repair} performed extremely well:  while it only localized three faults, those localizations were all extremely useful.
+MUSEUM provided three useful localizations, but no very high quality localizations.
+Only {\bf Repair} performed extremely welwith useful localizations for 5 faults, and localizations with the fault in the top-5 statements for 4 of these.
 

### 2016-02-21 introduction.tex
new text

@@
 Looking at all the failures to find  distinct faults is impractical.
-It is also impractical to apply the iterative approach:  first, fix one fault; then re-run all test cases, knowing that the remaining failures are due to different, unfixed faults, repeating until all failures go away.
-First, there may be some faults that are very difficult to fix --- a large compiler project typically has many open bugs.
+It is also impractical to apply the iterative approach:  first, fix one fault; then re-run all tests, knowing that the remaining failures are due to unfixed faults, repeating until all tests pass.
+However, there may be some (possibly known) faults that are very difficult to fix.
 As of 11/20/2015, LLVM clang had 89 bugs assigned but not resolved, some dating to 2008.
@@
 A research group with a novel fault-detection technique \cite{csmith,ISSTA12,ZhendongPLDI} may dump dozens of new faults on a project at once, producing more open problems.
+High priority faults may be hidden by stubborn low-priority faults.
 

### 2016-02-23 introduction.tex
latest

@@
 
-Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}  Unfortunately, the blessing of many failing test cases can become a curse for a developer.
+Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}.
+Unfortunately, the blessing of many failing test cases can become a curse for a developer.
 Simply discovering which test cases in a large set of failures represent distinct faults (or already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.

### 2016-02-23 experiments.tex
more edits

@@
 If the {\tt jsfunfuzz} produced test case executes to the end and prints {\tt TEST SUCCESSFULLY COMPLETED} it is considered successful.
-This is only a moderately strong oracle, and allows serious deviations in behavior from both the JavaScript language specification and normal SpiderMonkey behavior, while constraining other some of execution fairly tightly, such as {\tt eval} round-trips.
+This is only a moderately strong oracle, and allows serious deviations in behavior from both the JavaScript language specification and normal SpiderMonkey behavior, while constraining other aspects of execution fairly tightly, such as {\tt eval} round-trips.
 For GCC, the oracle is a differential check on a hash code:  failure was defined as producing a executable with the -O3 flag that, when executed, produced a different checksum than code compiled by either of GCC 4.9.3 or clang 7.0.0, both using -O0.

### 2016-03-01 abstract.tex
more edits

@@
 Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and useful in many test generation and debugging algorithms.
-This paper proposes a novel metric for measuring distance between executions, and specializes it to the most commonly needed case, failing test cases.
+This paper proposes a novel metric for measuring distance between executions, and specializes it to the most common application, determining similarity of failing test cases.
 The metric is based on a causal concept of distance: executions are similar to the degree that changes in the program itself, introduced using mutation operators, cause similar changes in the correctness of the executions.
 Specifically, if two failing test cases (for the original program) become successful due to the same mutant, they are more likely to be due to the same fault.
-This metric both improves an existing method for identifying test cases failing due to unique faults and enables novel fault localization approaches.
-We evaluate our methods using more than 50 faults and 2,500 test cases for two complex, widely-used, real-world compilers, and demonstrate improvements over state-of-the-art methods.
+We evaluate our metric using more than 50 faults and 2,500 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.

### 2016-03-01 introduction.tex
more edits

@@
 
-Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}.
+The goal of this paper, an effective metric for computing distance between program executions (in particular, failing test cases), is motivated in large part by progress in automated testing.
+Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}.
 Unfortunately, the blessing of many failing test cases can become a curse for a developer.
@@
 Repairs can also provide  information for understanding and debugging discovered faults.
-This approach also suggests a possible metric for comparing any two program executions, not just failing test cases.
+The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.
 
@@
 It is also impractical to apply the iterative approach:  first, fix one fault; then re-run all tests, knowing that the remaining failures are due to unfixed faults, repeating until all tests pass.
-However, there may be some (possibly known) faults that are very difficult to fix.
+There may be some (possibly known) faults that are very difficult to fix.
 As of 11/20/2015, LLVM clang had 89 bugs assigned but not resolved, some dating to 2008.
@@
 Ideally, a technique able to provide effective fuzzer taming should be able to translate its underlying rationale into some form of automated debugging assistance.
-A variety of automated debugging methods \cite{NearNeighbor,GroceError} have been based on distance, so there is some reason to expect that a high-quality metric might also contribute to that problem.
+A variety of automated debugging methods \cite{NearNeighbor,GroceError} have been based on distance, so there is reason to expect that a high-quality metric might also contribute to fault localization.
 

### 2016-03-01 metrics.tex
more edits

@@
 
-For every failure, the response to mutants defines a bit-vector, with
+For every failing test case, its response to mutants defines a bit-vector, with
 length equal to the number of mutants that repaired any test case,
-where a 1 bit indicates that the mutant in question repairs the test
+where a 1 bit indicates that the mutant in question repairs this test
 case.  Hamming distance between vectors is not suitable for our purposes.  Consider two test cases, $t_1$ and $t_2$, both repaired
@@
 
-This metric matches our intuitions about the above comparisons.  It
+This metric, the portion of mismatches over bits that are 1 in either
+vector, matches our intuitions about the above comparisons.  It
 essentially formalizes the intuition of the onion-ring model, where

### 2016-03-01 introduction.tex
fixes

@@
 
-Chen et al.
+In a PLDI 2013 paper, Chen et al.
 defined the \emph{fuzzer taming problem}: ``Given a potentially large collection of test cases, each of which triggers a bug, rank them in such a way that test cases triggering distinct bugs are early in the list.'' \cite{PLDI13}.
-The solution Chen et al.
-proposed to this problem was to use the Furthest-Point-First \cite{Gonzalez} (FPF) algorithm.
+Their proposed solution was to use the Furthest-Point-First \cite{Gonzalez} (FPF) algorithm.
 An FPF ranking of test cases requires a distance metric $d$ on test cases, and ranks test cases so that dissimilar tests appear earlier.
@@
 Rather than proposing a universal metric for use in fuzzer taming, Chen et al.
-offered many possible metrics (based on ideas in the literature, such as edit distances \cite{lev} or coverage spectra \cite{RepsSpectra} methods), none of which performed notably well across all three benchmark data sets, and some of which were not even applicable to all of the data sets.
+offered many possible metrics (based on ideas in the literature, such as edit distances \cite{lev} or coverage spectra \cite{RepsSpectra} methods), none of which performed notably well across all three benchmark data sets, and some of which were not even applicable to all data sets.
 For the two most difficult benchmarks, none of the metrics performed extremely well.

### 2016-03-01 relatedwork.tex
small fix related
fixes

@@
 The most closely related work is that of Chen et.
-al \cite{PLDI13}.
-The present paper uses two of the same data sets, applies the FPF \cite{Gonzalez} algorithm, and has the same empirical focus of determining how to identify faults for complex compilers.
+al \cite{PLDI13} (which shares one author with the present paper).
+We use two of the same data sets, apply the FPF \cite{Gonzalez} algorithm, and have the same empirical focus of determining how to identify faults for complex compilers.
 The primary difference is that the distance metrics used with FPF in the previous work were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.

### 2016-03-01 conclusions.tex
ready submit

@@
 Many software engineering techniques rely on measuring distance
-between program executions.  Not all of these concern failing
+between program executions \cite{BallConcept,NearNeighbor}.  Not all
+of these concern only failing
 executions.  For example, Zhang et al. \cite{issta14} show that FPF based on a simple
@@
 execution peers \cite{Sumner2011}, for example to help operators of
-spacecraft find similar behaviors to telemetry downloads in testbed
+spacecraft find similar behaviors to telemetry downlinks in testbed
 history \cite{KlausRajeev,scriptstospecs}.

### 2016-03-01 introduction.tex
ready submit

@@
 The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.
+Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).
 
@@
 
-\subsection{Comparing Executions by Response to Mutation}
+\subsection{Comparison by Response to Mutation}
 
@@
 Debugging by changing test cases is not useful; debugging by altering the program is precisely how we seek to use the causes we discover.
-Using mutants as causes in order to measure similarity essentially inverts the ideas of David Lewis \cite{LewisCause,LewisCount}, where similarity is used to determine causality.
+Using mutants as causes in order to measure similarity ``inverts'' the ideas of Lewis \cite{LewisCause,LewisCount}, where similarity determines causality.
 

### 2016-03-01 metrics.tex
fixes
ready submit

@@
 where a 1 bit indicates that the mutant in question repairs this test
-case.  Hamming distance between vectors is not suitable for our purposes.  Consider two test cases, $t_1$ and $t_2$, both repaired
+case.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
 by 100 mutants.  If these two test cases share 90 common repairing
@@
 large, despite the fact we have no evidence $t_1'$ and
-$t_2'$ are due to the same fault, in contrast to strong evidence for $t_1$ and $t_2$.  A Jaccard distance, used in some fault
-localization algorithms \cite{Liu06}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
+$t_2'$ are due to the same fault, in contrast to strong evidence for
+$t_1$ and $t_2$.  A Jaccard distance, also used (as similarity) in some fault
+localization algorithms \cite{Pinpoint}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
 %\vspace{-0.08in}

### 2016-03-01 relatedwork.tex
ready submit

@@
 al \cite{PLDI13} (which shares one author with the present paper).
-We use two of the same data sets, apply the FPF \cite{Gonzalez} algorithm, and have the same empirical focus of determining how to identify faults for complex compilers.
-The primary difference is that the distance metrics used with FPF in the previous work were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.
-This paper expands on Chen et al.
-by putting the FPF approach in the context of a much richer causal comparison of program executions that applies in theory to any executions, not just failing test cases.
+We use two of the same compiler data sets, and both apply the FPF \cite{Gonzalez} algorithm.
+The distance metrics used with FPF in the previous work, however, were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.
+This paper uses the FPF approach in the context of a more principled causal comparison of program executions that applies in theory to any executions, not just failing test cases, and introduces novel fault localizations.
 Prior to the FPF-based work, Francis et al.
@@
 
-The use of distance metrics in software engineering includes some of the localization and fault identification efforts discussed above.
+The use of distance metrics in software engineering includes some of the localization and fault identification efforts discussed above (e.g., \cite{NearNeighbor,GroceError,ChakiLev,Liu06}).
 Vangala et al.
@@
 \cite{Xin2008} examined methods for indexing program executions, useful in all methods involving execution alignments.
-To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra} (using counts of coverage of code entities), while our metric is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.
+To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra,BallConcept} (using counts over structural entities), while our metric is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.

### 2016-03-01 metrics.tex
ready submit

@@
 
-Given such a metric, the FPF \cite{Gonzalez} approach to fuzzer taming \cite{PLDI13}
-begins by selecting an arbitrary test case.   FPF then proceeds by selecting the test
+Given such a metric, FPF \cite{Gonzalez} fuzzer taming \cite{PLDI13}
+begins by ranking as first an arbitrary test case.   FPF next ranks the test
 case, of all unranked tests, that has the largest minimum distance
-from all already ranked test cases, repeating this until all tests are
-ranked.  In other words, FPF always chooses a test case that is as dissimilar as
-possible from all previously chosen test cases, based on the \emph{closest
-already chosen} test case.  FPF is boundedly close
+from all ranked test cases --- the test case most dissimilar by \emph{closest}
+already ranked test --- until all tests are ranked.  FPF is boundedly close
 to optimal under certain assumptions \cite{Gonzalez}.

### 2016-03-01 introduction.tex
ready submit

@@
 \caption{\scriptsize{The onion-ring model of repairs, showing all repairs for the center-most test case, and all other test cases repaired by those mutants.
-Circles represent repairing mutants, and diamonds repaired test cases.
+Rings represent repairing mutants, and diamonds repaired test cases.
 Diamond pattern symbolizes fault.
-For the innermost ring of repairing mutant(s), all test cases are due to the same fault.
-As rings expand to include more failures, the set of faults repaired grows (as does distance).}}
+For the innermost ring, all failures are due to the same fault.
+As rings expand to include more failures, the set of faults grows (as does distance).}}
+\vspace{-0.1in}
 \label{fig:onion}
@@
 Turning off a data-structure invariant checker is another possible false repair that might cover many different faults.
-In all these cases, so long as there are other, more interesting and more specific ways to repair some faults, a distance metric can still distinguish faults effectively.
+In all these cases, assuming more interesting and specific ways to repair faults exist, a distance metric can still work.
 
@@
 Using mutants as causes in order to measure similarity ``inverts'' the ideas of Lewis \cite{LewisCause,LewisCount}, where similarity determines causality.
+One pleasant side-effect is that execution representation \cite{NearNeighbor,RepsSpectra} is not a difficulty for this kind of metric, since an execution is always represented as a set of correctness evaluations.
 

### 2016-03-07 experiments.tex
Alex ready

@@
 In practice, after 50 tests, fixing a few faults and then re-running tests and FPF seems the most likely recourse, and we confirmed that after removing random subsets of faults, our metrics still outperform the previous best.
-We applied X-means \cite{xmeans} in an attempt to use clustering, but, as with previous results \cite{PLDI13} clustering did not compare well with FPF, and the runtime was much worse than for FPF-based methods.
-This tends to confirm earlier work's \cite{PLDI13} conclusion that clustering (at least based on X-means) simply does not work well in a setting with the kind of extreme disparity in instances of faults found in fuzzer taming.
-A further difficulty for our setting is that X-means and most off-the-shelf clustering tools take vectors, not a distance metric.
+We applied X-means \cite{xmeans} in an attempt to use clustering, but, as with previous results \cite{PLDI13} it did not compare well with FPF, and the runtime was much higher.
+This tends to confirm the conclusion \cite{PLDI13} that clustering (at least with X-means) does not work well in a setting with extreme disparity in instance counts for faults.
+%A further difficulty for our setting is that X-means and most off-the-shelf clustering tools take vectors, not a distance metric.
 X-means is using the raw repair data, not weighted by the fact that 0-0 agreement is much less informative than other possible matchings (a Euclidean over 0-1 vectors is just the square root of Hamming distance).

### 2016-03-12 experiments.tex
Alex ready to ship

@@
 The baseline rate for same-fault for test pairs was 33.64\%.
-Just knowing that two test cases have \emph{one} shared repairing mutant makes it 1.27x more likely that they are the same fault, and knowing they differ for one mutants makes it 0.6 times as likely they are due to the same fault.
+Just knowing that two test cases have \emph{one} shared repairing mutant makes it 1.27x more likely that they are the same fault, and knowing they differ for one mutant makes it 0.6 times as likely they are due to the same fault.
 Matching non-repair provides very weak evidence:  only a 34.14\% chance of matching fault.
@@
 We expect users to examine coverage changes and mutants.
-Blind, unaided pointers to portions of compiler code without more semantic information about why the statement is suspicious seems unlikely to work well, based on our experience with complex compiler faults, and the results of Parnin and Orso \cite{AutoHelp}.
+Blind, unaided pointers to portions of compiler code without more semantic information about why the statement is suspicious seem unlikely to work well, based on our experience with complex compiler faults, and the results of Parnin and Orso \cite{AutoHelp}.
 
@@
 MUSE performed poorly, with no useful (by the standard of having at least one faulty statement in the top 30) localizations.
-{\bf Coverage} produced one very good localization, and was the only method to localize that fault at all, but this was its only useful localization.
-MUSEUM provided three useful localizations, but no very high quality localizations.
-Only {\bf Repair} performed extremely welwith useful localizations for 5 faults, and localizations with the fault in the top-5 statements for 4 of these.
+{\bf Coverage} produced one very good localization, and was the only method to localize that fault at all, but this was one of only two useful localizations.
+MUSEUM provided 3 useful localizations, but no very high quality localizations.
+Only {\bf Repair} performed extremely well, with useful localizations for 5 faults, and the fault was in the top 5 statements for 4 of these.
 
@@
 However, the comparison is with the very best curve chosen after running more than 16 different metrics.
-In practice, users do not know ground truth to rank competing curves, so this is not a practical approach to fuzzer taming.
-The best methods for different subjects in previous work varied widely, even in such difficult-to-understand ways as less-fine-grained coverage providing better results in some cases, but worse in other cases.
+In practice, users do not know ground truth to rank curves, so PLDI 13 does not really give a practical approach to distance metric selection.
+The best methods for different subjects varied widely, even in such difficult-to-understand ways as less-fine-grained coverage providing better results in some cases, but worse in other cases.
 In practice, previous work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique.

### 2016-03-12 experiments.tex
Alex ready to ship

@@
 MUSEUM relies on a large set of mutation operators, and assumes that faulty statements will have multiple repairs, compared to non-faulty statements, or that faulty statements will be less likely to cause passing tests to fail when mutated.
-Restricting ourselves to cases where {\bf Repair} and MUSEUM both provided a localization, we know that some faulty statement was in a repairing mutant, but that passing tests could not distinguish this mutant  from other repairs.
+Restricting ourselves to cases where {\bf Repair} and MUSEUM both provided a localization, we know that some faulty statement was in a repairing mutant, but that passing tests could not distinguish this mutant  from other repairs (since we only use repairs that cause no passing tests to fail).
 Many of the faulty statement repairs (as expected for optimizing compilers) were statement deletions and conditional negations.
 For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.
-{\bf Repair} can still distinguish such mutants by relying on the onion-ring structure:  when a repair also repairs test cases that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults.
+
+{\bf Repair} can distinguish such mutants by relying on the onion-ring structure:  when a repair also repairs test cases that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults.
 The hypothesis behind {\bf Repair} is that \emph{repairs specific to the fault causing a failure are most likely to modify code involved in the fault itself}.

### 2016-05-02 response.txt
responses to reviewers

We thank the reviewers for very helpful comments.

R1:

Q: ...do you observe a correlation between distinct test cases and distinct invariants?

Even reduced test cases are hard to connect to invariants, tend to execute 10K+ LOC (see https://github.com/agroce/mutants16/tree/master/tests).  There probably is a correlation, but it's hard to spot for compiler faults.  We'd definitely like to investigate further.

Q: ...all bugs... due to... optimizations. Is this correct? ...do these... the majority of bugs...? ...results to change for non-compiler benchmarks?

For GCC ~90% of faults detected by changing from -O0 to -O3, so seems to be the rule for GCC.  For SpiderMonkey, "optimization" is hard to separate from odd JavaScript semantics, or attempts to speed up SpiderMonkey in general (not neccessarily code optimization exactly).  Results of Qi et al. suggest many programs _can_ be "patched" by turning off a feature, which is promising.

3) Section 1.2: “if two failures.., baseline probability*. What does this mean?

(Baseline) B = chance two failing tests are due to same fault.  Knowing two failures are "fixed" by turning off same optimization indicates probability they are same fault is > B.  We quantify (for our subjects) how much this holds (though not for optimization changes alone).

R2:

Q: ...fault of the center... repairs many others? Will that fault be discovered?

Really depends on the overall structure (what repairs those other tests).

Q: ...faults not discovered in the experiment...?

No obvious patterns, but differences would be complex properties of C/JavaScript and GCC/SpiderMonkey.  We'll investigate further.

Q: ...sampling does alter the results... undesirable happening in terms of the onion ring model.

Sampling hurts localization, there are probably ways to do good sampling for fuzzer taming (we are investigating).  Our guess: localization is bad when nothing in the middle of the onion ring is left, and that set could be small.

We'll provide numbers on runtimes (JavaScript not bad, GCC pretty bad, but we ran build/mutants in a very inefficient way).

R3:

Q: ...what kind of changes... ...does that count?  ...insert System.exit(0) at the beginning...?

Anything such that the test oracle passes is allowed.  For a very weak oracle, you could get away with this.  Due to onion ring, letting some "silly" changes make tests pass is probably harmless, so long as other more interesting mutants are also used.

Q: ...don’t quite understand... evaluation results...

Figure 2 shows _only_ fuzzer taming results, Tables 1-2 are _only_ for localization results.  These are two different goals, and evaluated separately.  FPF value only in Tables to show that how well a fault localizes and how well it is identified seem (surprisingly) unrelated.

Q: ...role of passing tests is "lower" than that of failing tests?

We believe so: probably for fuzzer taming, possibly for localization (note MUSEUM and other localizations agree).  Faulty code often executes in many passing tests; determining conditions for failure is difficult.  There is more information _about faults_ in failures, in some sense: e.g., single failing test can show a fault exists, which all passing tests cannot!

Q: ...how to interpret... Figure 3...

992 (commented "ADDED") is part of the code in the patch, as is the deleted statement (line #s in patch would help).  We'll clarify this much better, sorry!  3/6 is indeed ranked change.

Q:  ...confirm that these 8 faults are correct...?

For these 8 (and 16 GCC faults), we had a backpatch, based on commit data and comments, and tested it.  The other faults had such complex change history this was not feasible, making localization (not taming) evaluation impossible.  Backpatch sets were produced by colleagues, not us, before we began this work (so are not cherry-picked).

### 2016-05-02 response.txt
response changes
response edit

@@
 
-Even reduced test cases are hard to connect to invariants, tend to execute 10K+ LOC (see https://github.com/agroce/mutants16/tree/master/tests).
+Even reduced test cases are hard to connect to invariants, tend to execute ~10-50K+ LOC (see https://github.com/agroce/mutants16/tree/master/tests).
 There probably is a correlation, but it's hard to spot for compiler faults.
@@
 Figure 2 shows _only_ fuzzer taming results, Tables 1-2 are _only_ for localization results.
-These are two different goals, and evaluated separately.
+These are _different_ goals, both using the metric.
 FPF value only in Tables to show that how well a fault localizes and how well it is identified seem (surprisingly) unrelated.

### 2016-05-02 response.txt
tuning response

@@
 
-Sampling hurts localization, there are probably ways to do good sampling for fuzzer taming (we are investigating).
-Our guess: localization is bad when nothing in the middle of the onion ring is left, and that set could be small.
+Sampling hurts localization, maybe mostly ok for taming (we're investigating more).
+Our guess: localization is bad when nothing in the middle of the onion ring is left (could be small set), which taming can handle.
 
-We'll provide numbers on runtimes (JavaScript not bad, GCC pretty bad, but we ran build/mutants in a very inefficient way).
+We're very interested in alternatives to general mutation, thanks for ideas!
+
+Will provide runtimes (JavaScript ok, GCC bad, but we ran/built mutants in inefficiently).
 

### 2016-05-02 response.txt
tuning response
response edits

@@
 
-For GCC ~90% of faults detected by changing from -O0 to -O3, so seems to be the rule for GCC.
-For SpiderMonkey, "optimization" is hard to separate from odd JavaScript semantics, or attempts to speed up SpiderMonkey in general (not neccessarily code optimization exactly).
+~90% of GCC faults detected by changing from -O0 to -O3, so seems to be the rule for GCC.
+For SpiderMonkey, "optimization" hard to separate from odd JavaScript semantics, or attempts to speed up SpiderMonkey in general (not always traditional optimization).
 Results of Qi et al.
-suggest many programs _can_ be "patched" by turning off a feature, which is promising.
+suggest many programs _can_ be "patched" by turning off a feature, at least; MUSEUM also suggests generalization plausible.
 
@@
 Knowing two failures are "fixed" by turning off same optimization indicates probability they are same fault is > B.
-We quantify (for our subjects) how much this holds (though not for optimization changes alone).
+We quantify (for our subjects) how much this holds (though not for optimizations alone).
 
@@
 
-Will provide runtimes (JavaScript ok, GCC bad, but we ran/built mutants in inefficiently).
+Will provide runtimes (JavaScript ok, GCC bad, but we ran/built inefficiently).
 
@@
 
-Q: ...what kind of changes... ...does that count?  ...insert System.exit(0) at the beginning...?
+TACAS04 uses a _very_ different metric, in a very different way, we believe.
+
+Q: ...what kind of changes... ...insert System.exit(0)...?
 
@@
 
-Q: ...role of passing tests is "lower" than that of failing tests?
+Q: ...role of passing tests... "lower" than... failing tests?
 
-We believe so: probably for fuzzer taming, possibly for localization (note MUSEUM and other localizations agree).
-Faulty code often executes in many passing tests; determining conditions for failure is difficult.
-There is more information _about faults_ in failures, in some sense: e.g., single failing test can show a fault exists, which all passing tests cannot!
+We believe so: probably for fuzzer taming, possibly for localization (MUSEUM and other localizations agree).
+Faulty code often executes in many passing tests; determining conditions for failure is difficult; thus more information _about faults_ in failures, in some sense: e.g., single failing test can show a fault exists, which all passing tests cannot!
 

### 2016-05-02 response.txt
response edits

@@
 Knowing two failures are "fixed" by turning off same optimization indicates probability they are same fault is > B.
-We quantify (for our subjects) how much this holds (though not for optimizations alone).
+We quantify (for our subjects) how much this holds (just not for optimizations alone).
 
@@
 
-TACAS04 uses a _very_ different metric, in a very different way, we believe.
+TACAS04 uses a completely different metric, in a very different way, we believe.
 

### 2016-05-03 response.txt
response change

@@
 
-(Baseline) B = chance two failing tests are due to same fault.
+(Baseline) B = chance _any_ two failing tests are due to same fault.
 Knowing two failures are "fixed" by turning off same optimization indicates probability they are same fault is > B.
@@
 The other faults had such complex change history this was not feasible, making localization (not taming) evaluation impossible.
-Backpatch sets were produced by colleagues, not us, before we began this work (so are not cherry-picked).
+Backpatch sets were produced by colleagues, not us, before we began this work, to avoid bias.

### 2016-05-03 response.txt
deal with weird easychair word count
ready for fse submit

@@
 
-Q: ...do you observe a correlation between distinct test cases and distinct invariants?
+...significant over PLDI 2013...
 
-Even reduced test cases are hard to connect to invariants, tend to execute ~10-50K+ LOC (see https://github.com/agroce/mutants16/tree/master/tests).
-There probably is a correlation, but it's hard to spot for compiler faults.
-We'd definitely like to investigate further.
+Using our samplings and top-N PLDI 13 results (non-ideal method, hence not in paper), results _are_ _statistically_ significant (know not what you meant, but worth noting).
 
-Q: ...all bugs... due to... optimizations. Is this correct? ...do these... the majority of bugs...? ...results to change for non-compiler benchmarks?
+Q: ...correlation between... test cases and... invariants?
+
+Even reduced test cases are hard to connect to invariants, tend to execute ~10-50K+ LOC (https://github.com/agroce/mutants16/tree/master/tests).
+Probably is a correlation, but it's hard to spot for compiler tests.
+We'll definitely investigate further!
+
+Q: ...optimizations... majority of bugs...? ...non-compiler benchmarks?
 
@@
 Results of Qi et al.
-suggest many programs _can_ be "patched" by turning off a feature, at least; MUSEUM also suggests generalization plausible.
+suggest many programs _can_ be "patched" by turning off a feature; MUSEUM results also suggest generalization plausible.
 
-3) Section 1.2: “if two failures.., baseline probability*. What does this mean?
+3) ...baseline probability. What does this mean?
 
@@
 Knowing two failures are "fixed" by turning off same optimization indicates probability they are same fault is > B.
-We quantify (for our subjects) how much this holds (just not for optimizations alone).
+We quantify (in 4.1, 4.2) this claim (though not for optimizations alone).
 
@@
 
-Q: ...fault of the center... repairs many others? Will that fault be discovered?
+Q: ...center... repairs many others? Will that fault be discovered?
 
-Really depends on the overall structure (what repairs those other tests).
+Really depends on the overall structure (what repairs which other tests).
 
@@
 
-No obvious patterns, but differences would be complex properties of C/JavaScript and GCC/SpiderMonkey.  We'll investigate further.
+No obvious patterns, but differences would be complex properties of C/JavaScript and GCC/SpiderMonkey.  We'll investigate further!
 
@@
 
-Sampling hurts localization, maybe mostly ok for taming (we're investigating more).
-Our guess: localization is bad when nothing in the middle of the onion ring is left (could be small set), which taming can handle.
+Sampling hurts localization, maybe ok for taming (we're investigating).
+Our guess: localization is bad when nothing in the middle of the onion ring is left (could be small set), which hurts taming less.
 
@@
 
-TACAS04 uses a completely different metric, in a very different way, we believe.
+TACAS04 is completely different metric, used very differently, we believe.
 
@@
 
-Figure 2 shows _only_ fuzzer taming results, Tables 1-2 are _only_ for localization results.
-These are _different_ goals, both using the metric.
-FPF value only in Tables to show that how well a fault localizes and how well it is identified seem (surprisingly) unrelated.
+Figure 2 shows _only_ fuzzer taming results, no localization; Tables show _only_  localization results.
+These are _independent_ goals, using same metric.
+FPF column only in Tables to show that how well a fault localizes and how well it is identified seem (surprisingly) unrelated.
 
@@
 Faulty code often executes in many passing tests; determining conditions for failure is difficult; thus more information _about faults_ in failures, in some sense: e.g., single failing test can show a fault exists, which all passing tests cannot!
+We use passing tests to prune mutants.
 

### 2016-10-19 localizations.tex
more edits
fixed bad exp of localization

@@
 test; in our experiments, there were no repairs that did not change
-coverage.  If a fault can be avoided by, for example, avoiding a
+coverage.  If a fault can be avoided by, for example, disabling a
 problematic optimization in a compiler, this naturally leads to
@@
 This ranks a mutant $m$ (and associated statement) by the inverse of the distance
-from $f$ to the most distant statement also repaired by $m$, adding 1
+from $f$ to the most distant failng test also repaired by $m$, adding 1
 so that if $m$ only repairs failures at distance 0 $r$ is well defined

### 2016-10-20 experiments.tex
Add stats

@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{comparegcc}
+  \caption{APFD values for GCC suite slices}
+  \label{comparegcc}
+\end{figure}%
+
 There were 377,679 GCC mutants, based on 424,186 lines of code in C and header files.
@@
 Function coverage provides only 16 faults over 31 tests.
+
+Slicing the tests into 30 random equal-sized groups and using the PLDI 13 best metric (function coverage) and another competitive metric (line coverage), the difference between the repair-based metric and the previous metrics is statistically validated (Figure \ref{comparegcc}).
+The differences in APFD values are statistically significant between the mutation-based metric and both function and line coverage, by Wilcoxon test, with $p < 0.002$, and the mean APFD improvement, even for only 37 tests, exposing only 5-11 faults, is still over 3\%.
 

### 2016-10-20 experiments.tex
typos fixed

--- 2016-03-12 experiments.tex
@@
 
+Slicing the tests into 30 random equal-sized groups and comparing with function and line coverage metrics, the difference is statistically validated (Figure \ref{comparegcc}): differences in APFD values are statistically significant between our metric and both PLDI 13 metrics, by Wilcoxon test, with $p < 0.002$, and the mean APFD improvement, even for only 37 tests, exposing only 5-11 faults, is still over 3\%.
+
 For GCC wrong code faults, the baseline chance two test cases shared an underlying fault was 38.97\%.
-Knowing that two test cases shared a repairing fault raised the chance to 68.5\%, and knowing they disagreed on a fault lowered it to only 19.01\%.
+Knowing that two test cases shared a repairing mutant raised the chance to 68.5\%, and knowing they disagreed on a fault lowered it to only 19.01\%.
 The respective increase and decrease in probability of matching fault compared to baseline was thus 1.75 times greater for matching repairs, and less than 0.5 the chance of being the same fault, given one mismatched repair.

### 2016-10-21 experiments.tex
new exp and MUSEUM stuff

@@
 We applied X-means \cite{xmeans} in an attempt to use clustering, but, as with previous results \cite{PLDI13} it did not compare well with FPF, and the runtime was much higher.
-This tends to confirm the conclusion \cite{PLDI13} that clustering (at least with X-means) does not work well in a setting with extreme disparity in instance counts for faults.
+This tends to confirm the conclusion \cite{PLDI13} that clustering is of limited value in fuzzer taming.
+% (at least with X-means) does not work well in a setting with extreme disparity in instance counts for faults.
 %A further difficulty for our setting is that X-means and most off-the-shelf clustering tools take vectors, not a distance metric.
@@
 This is a somewhat better summary of results than the raw curve areas used in previous work \cite{PLDI13}.
-APFD, as the name suggests, shows the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's area to an ideal curve, with interpolation\footnote{The details of APFD calculation are slightly involved, and the interpolation and perfect curve are not exact fits for fuzzer taming; nonetheless the basic method summarizes curves fairly well and is standard in the testing literature for similar problems \cite{issta14}.}.
+APFD, as the name suggests, shows the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's area to an ideal curve, with interpolation.
+%\footnote{The details of APFD calculation are slightly involved, and the interpolation and perfect curve are not exact fits for fuzzer taming; nonetheless the basic method summarizes curves fairly well and is standard in the testing literature for similar problems \cite{issta14}.}.
 
@@
 Stopping at a plateau of size 5 produces the same results.
+
+In order to check whether the improvement is statistically reliable for SpiderMonkey 1.6 faults, we sliced the PLDI 13 tests into 20 randomly selected distinct subsets.
+The repair metric had a mean APFD for these subsets (about 90 tests, with a mean of just over 10 distinct faults) 2.85\% better than the mean APFD for the PLDI 13 metric, and the result was statistically significant ($p < 0.05$ by Wilcoxon test).
 
@@
 
+\begin{comment}
 \begin{figure}
@@
 \end{figure}
+\end{comment}
 
@@
 
+\begin{comment}
 \subsubsection{Error Explanation}
@@
 For each repairing mutant, in rank order, the highest ranking coverage changes for that mutant could also be shown, to give an idea of unusual (and potentially localizing) changes induced by repair.
+\end{comment}
 
@@
 
-Slicing the tests into 30 random equal-sized groups and comparing with function and line coverage metrics, the difference is statistically validated (Figure \ref{comparegcc}): differences in APFD values are statistically significant between our metric and both PLDI 13 metrics, by Wilcoxon test, with $p < 0.002$, and the mean APFD improvement, even for only 37 tests, exposing only 5-11 faults, is still over 3\%.
+Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both PLDI 13 metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement, even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults), is more than 3.8\% over either PLDI 13 metric.
 
@@
 
-In one sense, the discovery curve improvements here are practically useful, but not spectacular.
+In one sense, the discovery curve improvements here are practically very useful, but not exremely large in a relative sense.
 The SpiderMonkey curve has an APFD only slightly more than 2\% better than the best result from previous FPF efforts.
 For GCC, APFD improves by 6.4\%.
-However, the comparison is with the very best curve chosen after running more than 16 different metrics.
+However, the comparison is with the very \emph{best} curve chosen after running more than 16 different metrics.
 In practice, users do not know ground truth to rank curves, so PLDI 13 does not really give a practical approach to distance metric selection.
@@
 This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work.
+Moreover, the improvement is, for our subject programs, reliable:  choosing random subsets of the full data set, the difference in mean APFD is around 3\%, and statistically significant.
 
@@
 For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.
+We applied our repair localization to the MUSEUM data set, using the full Proteum mutant set and added MUSEUM-specific mutants, and for all but one of the 8 bugs in the data set, our results were equivalent to those for MUSEUM (in one case our approach fails to produce a localization).
 
-{\bf Repair} can distinguish such mutants by relying on the onion-ring structure:  when a repair also repairs test cases that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults.
+{\bf Repair} can distinguish more repairing mutants by relying on the onion-ring structure:  when a repair also repairs test cases that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults.
 The hypothesis behind {\bf Repair} is that \emph{repairs specific to the fault causing a failure are most likely to modify code involved in the fault itself}.
@@
 Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this is unlikely due to the nature of the repairs.
+For the MUSEUM data set, there are no cases with multiple faults for our method to use to further distinguish repairs, so its similarity to MUSEUM in results is not surprising (in this context, our approach is essentially MUSEUM modified to throw out any repairs that cause previously passing tests to fail).
 
 Finally, in contrast to many previously reported results in fault localization, typically over much simpler programs and faults than subtle compiler semantics bugs, all methods frequently failed to provide a useful localization at all.
+This is presumably due to the difficulty of compiler bugs:  our approach, and MUSEUM, performed very well on complex multi-lingual faults in the MUSEUM data set.
 In one sense, {\bf Coverage} is the most intriguing method presented here.
-While it did not perform nearly as well as {\bf Repair}, it did manage, in many cases, to rank some faulty line highly relative to the extremely large number of statements executed in each failing GCC run (on the order of 50K-100K+ statements).
+While it did not perform nearly as well as {\bf Repair}, it did manage, in many cases, to rank \emph{some} faulty line highly relative to the extremely large number of statements executed in each failing GCC run (on the order of 50K-100K+ statements).
 In future work, we plan to experiment with different methods for weighting frequency of coverage, perhaps using methods from machine learning, or of pruning changed coverage that is not likely to be relevant.

### 2016-10-21 localizations.tex
new exp and MUSEUM stuff

@@
 
-Another possible advantage over most statistical approaches to
-localization, even some recently proposed, more accurate, mutant-based
-approaches, is that our methods are not potentially confused in
+Another advantage over many statistical approaches to
+localization is that our methods are not potentially confused in
 settings with multiple faults.  Each localization/explanation is based

### 2016-10-21 experiments.tex
new exp and MUSEUM stuff

@@
 Andrews' tool applies only four operators: statement deletion, conditional negation, operator replacement, and constant replacement, chosen as a small set that still produces good results for C code.
-The tool was chosen to keep the number of mutants as small as possible to control the cost of mutation analysis.
 Coverage was analyzed with gcov.
@@
 
-In order to check whether the improvement is statistically reliable for SpiderMonkey 1.6 faults, we sliced the PLDI 13 tests into 20 randomly selected distinct subsets.
-The repair metric had a mean APFD for these subsets (about 90 tests, with a mean of just over 10 distinct faults) 2.85\% better than the mean APFD for the PLDI 13 metric, and the result was statistically significant ($p < 0.05$ by Wilcoxon test).
-
 Over all mutants, and all pairs of test cases repaired by the same mutant (so the same pair may count many times, if many mutants repair both test cases), the probability of being due to the same fault was 42.77\%, and the probability of being due to the same fault if two test cases disagreed on a repair (the mutant repaired one test case but not the other) was only 20.33\%.
@@
 The mean and median homogeneity for repairing mutants (\% of failures repaired corresponding to the most common fault repaired) were 77.96\% and 86.36\%, respectively.
+
+In order to check whether the improvement is statistically reliable for SpiderMonkey 1.6 faults, we sliced the PLDI 13 tests into 20 randomly selected equal-sized (as much as possible) distinct subsets.
+The repair metric had a mean APFD (89.23\%) for these subsets (about 90 tests, with a mean of just over 10 distinct faults) 2.85\% better than the mean APFD for the PLDI 13 metric (86.76\%), and the result was statistically significant ($p < 0.05$ by Wilcoxon test).
+Median repair APFD was 90.05\%, vs.
+84.75\% for PLDI 13's best metric.
 
@@
 
-Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both PLDI 13 metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement, even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults), is more than 3.8\% over either PLDI 13 metric.
-
 For GCC wrong code faults, the baseline chance two test cases shared an underlying fault was 38.97\%.
@@
 The mean and median homogeneity (\% of repaired tests corresponding to most common repaired fault) were 79.4\% and 100\%, respectively.
+
+Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both PLDI 13 metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement, even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults), is more than 3.8\% over either PLDI 13 metric.
 
@@
 These can be pruned at will by running them on passing test cases, and removing mutants that fail on any test case.
+\begin{comment}
 Keeping hot mutants for a heavily modified program could be costly (since each mutant must be recompiled, if not re-tested, for each modification to the code), but for stable releases this is a simple approach to keep costs low.
+\end{comment}
 

### 2016-10-21 experiments.tex
new exp and MUSEUM stuff

@@
 The largest threat to validity is that this paper's conclusions rely on only two data sets, used in a previous paper.
-The fault ground truths are possibly imperfect, due to the complex change history, and it was not possible to produce good patches for all faults, limiting our analysis of localization and explanation.
+The fault ground truths are possibly imperfect, due to the complex change history, and it was not possible to produce good patches for all faults, limiting our analysis of fault localization.
 Generalizing to all compilers, or to software systems that are not compilers, would be unwarranted.

### 2016-12-19 experiments.tex
more fixes and such

@@
 The oracle for SpiderMonkey executions is the set of checks built in to {\tt jsfunfuzz} plus the requirement that the execution not crash.
-If the {\tt jsfunfuzz} produced test case executes to the end and prints {\tt TEST SUCCESSFULLY COMPLETED} it is considered successful.
-This is only a moderately strong oracle, and allows serious deviations in behavior from both the JavaScript language specification and normal SpiderMonkey behavior, while constraining other aspects of execution fairly tightly, such as {\tt eval} round-trips.
+This is only a moderately strong oracle, and allows serious deviations from both the JavaScript language specification and normal SpiderMonkey behavior, while checking some complex details, such as {\tt eval} round-trips.
 For GCC, the oracle is a differential check on a hash code:  failure was defined as producing a executable with the -O3 flag that, when executed, produced a different checksum than code compiled by either of GCC 4.9.3 or clang 7.0.0, both using -O0.
-A repairing mutant must enable GCC 4.3.0 to actually compile code correctly, which is a very strict correctness property, making coincidental correctness \cite{CCT} highly unlikely.
-The only weakness is that there is no check on speed of or time to produce compiled code.
+A repairing mutant must enable GCC 4.3.0 to actually compile code correctly, which is a very strict correctness property, making coincidental correctness \cite{CCT} highly unlikely --- only missed optimizations are allowed.
 
@@
 
-In order to check whether the improvement is statistically reliable for SpiderMonkey 1.6 faults, we sliced the PLDI 13 tests into 20 randomly selected equal-sized (as much as possible) distinct subsets.
-The repair metric had a mean APFD (89.23\%) for these subsets (about 90 tests, with a mean of just over 10 distinct faults) 2.85\% better than the mean APFD for the PLDI 13 metric (86.76\%), and the result was statistically significant ($p < 0.05$ by Wilcoxon test).
+In order to check our results statistically, we sliced the PLDI 13 tests into 20 randomly selected equal-sized (as much as possible) distinct subsets.
+The repair metric had a mean APFD (89.23\%) for these subsets that was 2.85\% better than the mean APFD for the PLDI 13 metric (86.76\%), over sets of $\sim 10$ faults.
+The result was statistically significant ($p < 0.05$ by Wilcoxon test).
 Median repair APFD was 90.05\%, vs.
@@
 Of these mutants, 73,016 were covered by some failing test case.
-Of these mutants, 41,385 survived a GCC bootstrap build (used to produce our compiled versions for repair checks), compiled hello world correctly, and passed a GCC test suite.
+Of these mutants, 41,385 survived a GCC bootstrap build, compiled hello world, and passed a GCC test suite.
 Of these mutants, 3,232 (7.8\%) repaired at least one test case.
@@
 Most mutants fixed only a few tests, to a greater extent than was true with SpiderMonkey.
-The mean and median homogeneity (\% of repaired tests corresponding to most common repaired fault) were 79.4\% and 100\%, respectively.
+Mean homogeneity (\% of repaired tests corresponding to most common repaired fault) was 79.4\%, median 100\%.
 
@@
 
-Table \ref{gcctable} shows localization results for GCC 4.3.0 wrong code faults.
-For the same reasons as with SpiderMonkey, only some faults were deemed to have a strong enough ground truth patch for localization evaluation.
+Table \ref{gcctable} shows localization results for GCC 4.3.0.
+Again, only some faults were deemed to have a strong enough ground truth patches for localization evaluation.
 MUSE performed poorly, with no useful (by the standard of having at least one faulty statement in the top 30) localizations.
-{\bf Coverage} produced one very good localization, and was the only method to localize that fault at all, but this was one of only two useful localizations.
+{\bf Coverage} produced one very good localization, and was the only method to localize that fault at all, but usually performed poorly.
 MUSEUM provided 3 useful localizations, but no very high quality localizations.
-Only {\bf Repair} performed extremely well, with useful localizations for 5 faults, and the fault was in the top 5 statements for 4 of these.
+Only {\bf Repair} performed very well, with useful localizations for 5 faults, and the fault was in the top 5 statements for 4 of these.
 
@@
 The hypothesis behind {\bf Repair} is that \emph{repairs specific to the fault causing a failure are most likely to modify code involved in the fault itself}.
-This seems to be true, when such a repair exists.
+This seems to be true, when such a repair exists, and results in {\bf Repair} either providing a ``useful'' localization or no localization at all, which user studies and surveys suggest is the ideal behavior due to the cost of false positives in localization \cite{AutoHelp,Kochhar}.
 If any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
@@
 First, the primary cost is compiling the mutated versions of the SUT.
-Running each test case under each mutated version of the compiler is cheap: each execution requires around 0.05 seconds for SpiderMonkey, and 0.12 seconds for GCC, on average.
-These executions can obviously be performed in parallel, on low-cost virtual machines in the cloud.
-The cost is further reduced by a few more facts: most failures do not cover most mutants, and vice versa, so the total number of repair checks needed is far less than the product of mutants and failures.
+Running each test case under each mutated version of the compiler is cheap: each execution requires around 0.05 seconds for SpiderMonkey, and 0.12 seconds for GCC, on average; moreover, executions can be done in parallel, most failures do not cover most mutants, and vice versa, so the total number of repair checks needed is far less than the product of mutants and failures, and can be spread over many machines.
 The practical approach to fuzzer taming for a large, ongoing project is likely to maintain a set of ``hot'' pre-compiled mutants.

### 2016-12-19 introduction.tex
Alex ready
more fixes and such

@@
 The goal of this paper, an effective metric for computing distance between program executions (in particular, failing test cases), is motivated in large part by progress in automated testing.
-Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}.
+Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{We claim testing/debugging complex systems software is well represented by compiler testing as an instance.}.
 Unfortunately, the blessing of many failing test cases can become a curse for a developer.
@@
 Looking at all the failures to find  distinct faults is impractical.
-It is also impractical to apply the iterative approach:  first, fix one fault; then re-run all tests, knowing that the remaining failures are due to unfixed faults, repeating until all tests pass.
+It is also impractical to apply the iterative approach:  first, fix one test's fault; then re-run all tests, until all tests pass.
 There may be some (possibly known) faults that are very difficult to fix.
-As of 11/20/2015, LLVM clang had 89 bugs assigned but not resolved, some dating to 2008.
+As of 11/20/2015, clang had 89 unresolved, assigned bugs, some dating to 2008.
 Knowing if a randomly generated test case is an instance of a known failure can be difficult, and fixing these problems is unlikely.
-A research group with a novel fault-detection technique \cite{csmith,ISSTA12,ZhendongPLDI} may dump dozens of new faults on a project at once, producing more open problems.
-High priority faults may be hidden by stubborn low-priority faults.
+A research group with a novel fault-detection technique \cite{csmith,ISSTA12,ZhendongPLDI} may dump dozens of new faults on a project at once, and high priority faults may be hidden by stubborn low-priority faults.
 
@@
 
-\subsection{The Onion-Ring and Causality}
+%\subsection{The Onion-Ring and Causality}
 
@@
 If a mutant turns off a large number of optimizations, it is weak evidence.
-Characterizing the granularity of program mutants is likely difficult, but fortunately it is not necessary.
+Must we then characterize mutant granularity?
+No.
 
@@
 
-\subsection{Contributions}
+%\subsection{Contributions}
 

### 2016-12-19 localizations.tex
more fixes and such

@@
 (and maximal).
-If $m$ does not repair $f$, it is not ranked at all.
+If $m$ does not repair $f$, it is not ranked at all.  One possible
+advantage of this approach, unique to our knowledge, is that by
+ranking \emph{mutants} the localization offers developers not simply a
+ranked statement but a kind of ``explanation'' of the fault:  if this
+mutant were applied, the fault would not exhibit.  As Kochhar et
+al. show \cite{Kochhar}, a large majority (more than 85\%) of developers considered
+the ability of a localization tool to provide a \emph{rationale} or
+reason for a localization important.  A mutation plus the tests that
+stop failing when it is applied is a practical, concrete explanation
+of a fault localization.
 

### 2016-12-19 metrics.tex
ready submit
more fixes and such

@@
 $t_1'$ and $t_2'$, where $t_1'$ is repaired by 5 mutants, $t_2'$ is
-repaired by 5 mutants, and \emph{none of those mutants overlap.}   Hamming distance for the first pair of tests is twice as
-large, despite the fact we have no evidence $t_1'$ and
-$t_2'$ are due to the same fault, in contrast to strong evidence for
-$t_1$ and $t_2$.  A Jaccard distance, also used (as similarity) in some fault
+repaired by 5 mutants, and \emph{none of those mutants overlap.}   A Jaccard distance, also used (as similarity) in some fault
 localization algorithms \cite{Pinpoint}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
@@
 This metric, the portion of mismatches over bits that are 1 in either
-vector, matches our intuitions about the above comparisons.  It
+vector, matches our intuitions about the above comparisons\footnote{In Section \ref{conc} we discuss possible uses of $d$
+requiring comparison of both failing and passing test cases:  in this 
+case, we would use three-valued vectors, where 1 indicates a 
+failing property succeeds and -1 that a succeeding property 
+fails.}.  It
 essentially formalizes the intuition of the onion-ring model, where
@@
 metric we compare it, alone, to the previous best metrics.
-
-In Section \ref{conc} we briefly discuss possible uses of $d$
-requiring comparison of both failing and passing test cases. The only
-change required is to use three-valued vectors, where 1 indicates a failing property succeeds and -1 indicates a succeeding property fails.

### 2016-12-19 relatedwork.tex
fix for double blind
more fixes and such

@@
 The most closely related work is that of Chen et.
-al \cite{PLDI13} (which shares one author with the present paper).
+al \cite{PLDI13}.
 We use two of the same compiler data sets, and both apply the FPF \cite{Gonzalez} algorithm.
@@
 \cite{Podgurski03} and others \cite{Liu06,Liblit05} used clustering to attempt to solve similar problems in identifying distinct faults, on simpler human-generated data sets.
-The work of Chen et.
-al \cite{PLDI13} provides an in-depth comparison of clustering and ranking, and the advantages of ranking (as well as some possible limitations).
+Chen et.
+al \cite{PLDI13} provide an in-depth comparison of clustering and ranking, and the advantages of ranking (as well as some possible limitations).
 Jones et al.

### 2017-01-18 localizations.tex
Minor fixes

@@
 (and maximal).
-If $m$ does not repair $f$, it is not ranked at all.  One possible
-advantage of this approach, unique to our knowledge, is that by
+If $m$ does not repair $f$, it is not ranked at all.  One advantage of
+this approach over some statistical methods is that by
 ranking \emph{mutants} the localization offers developers not simply a
-ranked statement but a kind of ``explanation'' of the fault:  if this
-mutant were applied, the fault would not exhibit.  As Kochhar et
+ranked statement but a kind of ``explanation'' \cite{GroceError} of the fault:  if \emph{this
+mutant} were applied, the fault would not exhibit.  As Kochhar et
 al. show \cite{Kochhar}, a large majority (more than 85\%) of developers considered
@@
 stop failing when it is applied is a practical, concrete explanation
-of a fault localization.
+of a fault localization.  Mutants that repair a failure provide insight 
+into the nature of the fault, and, we hypothesize, more information the 
+closer they are to the center of the onion-ring.
 
-One important aspect of these approaches is that they 
-can provide qualitative information for debugging, known as an \emph{error
-  explanation} \cite{GroceError}, in addition to localization.  For any
+The coverage localization also provides a kind of explanation, though
+a less direct one.  For any
 interesting line of code in the coverage-change-based localization,
@@
 browse through all mutants that produce this coverage change for the test
-case.  Mutants that repair a failure provide insight
-into the nature of the fault, and, we hypothesize, more information the
-closer they are to the center of the onion-ring.
+case.
 

### 2017-01-28 introduction.tex
Minor fixes
intro fixup

@@
 There may be some (possibly known) faults that are very difficult to fix.
-As of 11/20/2015, clang had 89 unresolved, assigned bugs, some dating to 2008.
-Knowing if a randomly generated test case is an instance of a known failure can be difficult, and fixing these problems is unlikely.
+As of 1/16/2017, clang had 109 unresolved, assigned bugs, some dating to 2008 \cite{clangbugs}.
+Knowing if a randomly generated test case is an instance of a known failure can be difficult, and fixing long-standing problems to improve bug triage is unlikely.
 A research group with a novel fault-detection technique \cite{csmith,ISSTA12,ZhendongPLDI} may dump dozens of new faults on a project at once, and high priority faults may be hidden by stubborn low-priority faults.
@@
 Rather than proposing a universal metric for use in fuzzer taming, Chen et al.
-offered many possible metrics (based on ideas in the literature, such as edit distances \cite{lev} or coverage spectra \cite{RepsSpectra} methods), none of which performed notably well across all three benchmark data sets, and some of which were not even applicable to all data sets.
+offered many possible metrics (based on ideas in the literature, such as edit distances \cite{lev} or coverage spectra \cite{RepsSpectra} methods), none of which performed notably well across all three of their benchmark data sets, and some of which were not even applicable to all data sets.
 For the two most difficult benchmarks, none of the metrics performed extremely well.

### 2017-01-28 abstract.tex
alex fix

@@
 Specifically, if two failing test cases (for the original program) become successful due to the same mutant, they are more likely to be due to the same fault.
-We evaluate our metric using more than 50 faults and 2,500 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.
+We evaluate our metric using more than 50 faults and 2,800 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.

### 2017-01-28 experiments.tex
alex fix

@@
 In practice, after 50 tests, fixing a few faults and then re-running tests and FPF seems the most likely recourse, and we confirmed that after removing random subsets of faults, our metrics still outperform the previous best.
-We applied X-means \cite{xmeans} in an attempt to use clustering, but, as with previous results \cite{PLDI13} it did not compare well with FPF, and the runtime was much higher.
+We applied X-means \cite{xmeans} in an attempt to use clustering with our metrics, but, as with previous results \cite{PLDI13} it did not compare well with FPF, and the runtime was much higher.
 This tends to confirm the conclusion \cite{PLDI13} that clustering is of limited value in fuzzer taming.

### 2017-01-28 experiments.tex
more fixes

@@
 
-APFD results are useful summaries, but the curve itself is also worth examining, since (1) a very good early curve is perhaps most important to developers and (2) long sequences of tests with no new faults may discourage users more than an overall less effective but steadily climbing curve with few ``plateaus'' as we call these uninformative sequences of tests.
+APFD results are useful summaries, but the curve itself is also worth examining, since (1) long sequences of tests with no new faults may discourage users more than an overall less effective but steadily climbing curve with few ``plateaus'' as we call these uninformative sequences of tests and (2) a good early curve is important to developers.
 The SpiderMonkey curve climbs very rapidly in the early portion, with perfect discovery for the first 12 faults.
@@
 19 with our metric.
-Stopping at a plateau of size 5 produces the same results.
+Stopping at a plateau of size 6 produces the same results.
 
@@
 Just knowing that two test cases have \emph{one} shared repairing mutant makes it 1.27x more likely that they are the same fault, and knowing they differ for one mutant makes it 0.6 times as likely they are due to the same fault.
-Matching non-repair provides very weak evidence:  only a 34.14\% chance of matching fault.
+Matching non-repair, however, provides very weak evidence:  only a 34.14\% chance of matching fault, just over baseline probability.
 
@@
 A few mutants fixed a very large number of tests, up to 1,050.
-These all appear to be turning off all optimizations --- running in -O0 mode.
+These all appear to be turning off all optimizations --- essentially running gcc in -O0 mode.
 Most mutants fixed only a few tests, to a greater extent than was true with SpiderMonkey.
@@
 
-Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both PLDI 13 metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement, even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults), is more than 3.8\% over either PLDI 13 metric.
+Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both PLDI 13 metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement --- even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults) --- is more than 3.8\% better than either PLDI 13 metric.
 
@@
 However, the comparison is with the very \emph{best} curve chosen after running more than 16 different metrics.
-In practice, users do not know ground truth to rank curves, so PLDI 13 does not really give a practical approach to distance metric selection.
+In practice, users do not know ground truth to rank curves, thus Chen et al.
+\cite{PLDI13} do not really give a practical approach to distance metric selection in their PLDI work.
 The best methods for different subjects varied widely, even in such difficult-to-understand ways as less-fine-grained coverage providing better results in some cases, but worse in other cases.
-In practice, previous work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique.
-It also included methods such as comparing output signatures that are inapplicable to differential properties.
+In practice, their work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique.
+It also included methods such as comparing output signatures that are inapplicable to differential testing in most cases, the most critical compiler application.
 This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work.

### 2017-01-28 relatedwork.tex
more fixes

@@
 
-Some approaches to localization (and all approaches to patching) assume the possibility of a mutant \emph{actually} fixing a fault.
+Some approaches to localization or patching assume the possibility of a mutant/change \emph{actually} fixing a fault.
 In our experiments, we found this to be highly unlikely -- most of our patches were far larger than the largest higher-order-mutants \cite{jia2009higher} anyone usually considers applying, and none were equivalent to a repair.
@@
 proposed using distance to cluster test cases to improve diversity and eliminate duplicates \cite{VangalaDist}.
+We speculate that using FPF with our causal metric over passing tests as a test prioritization (or selection) method might be a fruitful approach \cite{YooHarman}.
 Adaptive random testing uses distances (usually between inputs) to choose tests \cite{Chen,ARTChen,ISSTAART}, and Artzi et al.

### 2017-01-29 conclusions.tex
ok

@@
 \cite{PLDI13} and discover the source code locations of faults
-\cite{MUSE,multilingual}.
+\cite{MUSE,multilingual}.  For a set of more than 2,800 tests and more
+than 50 faults, over the Mozilla SpiderMonkey 1.6 JavaScript compiler and
+GCC 4.3.0, our methods improve fault identification (``fuzzer taming''), in a statistically significant way
+(with effect size of about 3\% improvement), over
+the \emph{best} (for each subject program) of over 16 metrics considered in previous work.
+Preliminary investigation also shows improvement (for a shared
+mutation testing budget) over previous mutant-based fault localization
+methods, for the same tests and a subset of 24 faults with known
+locations.  Our {\bf Repair} method, in particular, was the only
+localization approach to provide any perfect localizations of faults
+(it produced 3),
+and provided three times as many very-high-quality (ranking the fault
+in top 5 locations) localizations as the next best method
+\cite{multilingual}  --- 6 such localizations, compared to only 2.
+For GCC wrong-code faults, only our methods produced any very-high-quality fault localizations.
 
@@
 assistance.  One idea is to combine the metric approach and the ideas
-in MUSE and MUSEUM, in some way, to exploit passing tests and perhaps
-additional coverage information.
+in MUSE and MUSEUM, or other mutant-based localizations such as
+Metallaxis \cite{Metallaxis} in some way, to exploit passing tests and perhaps
+additional coverage information.  We would also be very interested to
+compare our localizations with the Metallaxis
+\cite{Metallaxis,Papadakis} approach.
 
-%We also plan to explore other applications of mutant-based metrics.
+We also plan to explore other applications of mutant-based metrics.
 Many software engineering techniques rely on measuring distance
@@
 mutants to refine distance when coarser metrics plateau
-may be practical.  Mutants may also be useful in determining
+may be practical.  Cost may be less of a concern when applying our
+metrics as a method for prioritizing or selecting regression tests,
+where the results can be repeatedly used in test suite execution,
+without recomputing distances based on every code change
+\cite{YooHarman}.  Our metrics may also be useful in determining
 execution peers \cite{Sumner2011}, for example to help operators of

### 2017-01-29 experiments.tex
ok

@@
 Fault localization results were considerably worse, however (with relative rankings preserved) for both our techniques and MUSE/MUSEUM.
+Finally, we have preliminary confirmation that a guided-samping approach can provide good fuzzer taming (if not fault localizaton):  compute vectors for a small sample of mutants, then report highly distinct failing tests.
+These are likely to be singleton failures.
+After repairing some singletons, choose mutants to evaluate that are covered by multiple tests with few or no repairing mutants (since such tests are the ones most likely to be distinct but not yet identified).
+Such an approach could be quite cost-effective, requiring only a small number of mutant evaluations.
 

### 2017-01-29 relatedwork.tex
ok

@@
 Fault localization \cite{FaultSurvey,NearNeighbor,Liblit03,Liblit05,Cleve05,Jones2002,Jones05,Jones07,GroceError,ChakiLev,Liu06,SPIN03,Santelices:ICSE:2009,Abreu:2006:PRDC,xuan2014test,DD}, even specifically for compilers \cite{Whalley94}, is a long-standing field of research.
-Recently, fault localization researchers have used program mutants to improve statistical \cite{Jones2002} fault localization techniques \cite{MUSE,DebroyMutant,FasterMutant,multilingual,Metallaxis}.
+Recently, fault localization researchers have used program mutants to improve statistical \cite{Jones2002} fault localization techniques \cite{MUSE,DebroyMutant,FasterMutant,multilingual,Metallaxis,Papadakis}.
 %The core of this work has been the belief that faulty program locations are (1) more likely to change a failing test to successful when mutated and (2) less likely to change a successful test case to faulty when mutated.
@@
 Previous work has shown that, across a large body of open source programs in many languages, most fault fixes were larger in size than even low-order higher-order mutants \cite{GopinathMutants}.
-%We speculate that fixes for compilers may be likely to be even more complex than the average, given that compilers are particularly complex software systems.  
-%Our primary assumption is simply that similar executions respond similarly to mutants, and in particular executions due to the same fault can be ``repaired'' by similar program changes.
-These changes are likely \emph{related} to the actual fault, but are unlikely to be the actual fault, for reasons of comparative complexity of mutants and real fixes.
+We speculate that fixes for compilers may be likely to be even more complex than the average, given that compilers are particularly complex software systems.  
+Our primary assumption is simply that similar executions respond similarly to mutants, and in particular executions due to the same fault can be ``repaired'' by similar program changes.
+These changes are likely to be \emph{related} in some causal way to the actual fault, but are unlikely to be the actual fault, for reasons of comparative complexity of mutants and real fixes.
 

### 2017-01-29 experiments.tex
typo

--- 2017-01-28 experiments.tex
@@
 Fault localization results were considerably worse, however (with relative rankings preserved) for both our techniques and MUSE/MUSEUM.
+Finally, we have preliminary results indicating that a guided-sampling approach may provide good fuzzer taming (if not fault localizaton):  compute vectors over a small constant-sized \cite{RahulISSRE} sample of mutants, then report highly distinct failing tests.
+Other metrics, such as coverage-based comparisons, could provide hints as to likely distinct tests (and thus mutants covered by them) to focus on first.
+These are likely to be different faults.
+After repairing some of these, choose mutants to evaluate that are covered by multiple tests with few or no repairing mutants (since such tests are the ones most likely to be distinct but not yet identified).
+Such an approach could be quite cost-effective, requiring only a relatively small number of mutant evaluations before providing useful results.
 

### 2017-01-29 experiments.tex
spelling

@@
 
-In one sense, the discovery curve improvements here are practically very useful, but not exremely large in a relative sense.
+In one sense, the discovery curve improvements here are practically very useful, but not extremely large in a relative sense.
 The SpiderMonkey curve has an APFD only slightly more than 2\% better than the best result from previous FPF efforts.
@@
 Fault localization results were considerably worse, however (with relative rankings preserved) for both our techniques and MUSE/MUSEUM.
-Finally, we have preliminary results indicating that a guided-sampling approach may provide good fuzzer taming (if not fault localizaton):  compute vectors over a small constant-sized \cite{RahulISSRE} sample of mutants, then report highly distinct failing tests.
+Finally, we have preliminary results indicating that a guided-sampling approach may provide good fuzzer taming (if not fault localization):  compute vectors over a small constant-sized \cite{RahulISSRE} sample of mutants, then report highly distinct failing tests.
 Other metrics, such as coverage-based comparisons, could provide hints as to likely distinct tests (and thus mutants covered by them) to focus on first.

### 2017-01-29 localizations.tex
spelling

@@
 This ranks a mutant $m$ (and associated statement) by the inverse of the distance
-from $f$ to the most distant failng test also repaired by $m$, adding 1
+from $f$ to the most distant failing test also repaired by $m$, adding 1
 so that if $m$ only repairs failures at distance 0 $r$ is well defined

### 2017-04-30 abstract.tex
IEEE format

@@
 Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and useful in many test generation and debugging algorithms.
-This paper proposes a novel metric for measuring distance between executions, and specializes it to the most common application, determining similarity of failing test cases.
+This paper proposes a metric for measuring distance between executions, and specializes it to an important application, determining similarity of failing test cases, for the purpose of automated fault identification and localization in debugging based on automatically generated compiler tests.
 The metric is based on a causal concept of distance: executions are similar to the degree that changes in the program itself, introduced using mutation operators, cause similar changes in the correctness of the executions.
-Specifically, if two failing test cases (for the original program) become successful due to the same mutant, they are more likely to be due to the same fault.
+Specifically, if two failing test cases (for the original compiler) become successful due to the same mutant, they are more likely to be due to the same fault.
 We evaluate our metric using more than 50 faults and 2,800 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.

### 2017-04-30 experiments.tex
small fixes
IEEE format

@@
 In a limited sense, information from passing tests is already incorporated in our results.
-Throwing out all mutants that are killed by any passing test as potential repairs/localizations ensures that the rankings of all statements that are in MUSE/MUSEUM rankings are correct, relative to each other.
+Throwing out all mutants that are killed by any passing test as potential repairs/localizations ensures that the rankings of all statements that are in MUSE/MUSEUM rankings are correct, \emph{relative to each other}.
 By definition, passing tests have no influence on the suspiciousness of these mutants.
-However, there may be other mutants that 1) repair a failure and 2) are killed by some test case: these could, if enough repaired a faulty statement, improve MUSE/MUSEUM results.
+However, there may be other mutants that 1) repair a failure and 2) are killed by some test case: these could, if enough different ones repaired the same faulty statement, improve MUSE/MUSEUM results, though in practice this is extremely unlikely.
 We reject such mutants in part to keep costs low, and in part because we think that the causal information contained in such mutants, that repair a failure but also break some passing test, is problematic.
 They seem likely to be less useful as explanations of the causes of a failure, since they do not impact the program semantics in a way that is only known to be beneficial, and could potentially even mislead a user if used as explanations.
+Still, it is most useful to compare MUSEUM/MUSE and our approach on the cases where all methods produce a localization.
 
 Discussion with the MUSE/MUSEUM authors confirms  that adding information for passing tests, while costly, would likely improve the MUSE and MUSEUM results, given the weakness of our oracles, particularly for SpiderMonkey, and should be considered essential for ideal application of their approach.
-Because the cost of recording the full mutant analysis matrix for all passing tests (rather than considering a mutant killed as soon as one test fails, as we do) is very high, and we would like to produce a comparison over a fixed computational cost, we give results for MUSE/MUSEUM over failing tests only, cautioning that we are not sure what the impact of this choice is.
+Because the cost of recording the full mutant analysis matrix for all passing tests (rather than considering a mutant killed as soon as one test fails, as we do) is very high, and we would like to produce a comparison over a fixed computational cost, we give results for MUSE/MUSEUM over failing tests only, cautioning that we are not sure what the impact of this choice is on faults that were not localized.
 For similar reasons of keeping costs low, we also used the mutation operators of Andrews vs.
@@
 
-\begin{comment}
-\begin{figure}
-{\scriptsize
-\begin{code}
-{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):}
-...
-<             vp[1] = OBJECT\_TO\_JSVAL(thisp);
-<         \} else \{
-<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);
-<         \}
-...
-<                   a->avail = (jsuword)sp;
-<               \}
-...
----
->         RESTORE\_SP(fp);
-...
-{\bf Mutant information:}
-Failure repaired by 148 mutants.
-...
-\#3: delete statement mutant of jsinterp.c:944:
-
-            vp[1] = OBJECT\_TO\_JSVAL(thisp);
-        \} else \{
-/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/
-        \}
-
-{\bf Coverage change information:}
-...
-\#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:
-                   a->avail = (jsuword)sp; /* ADDED */
-\end{code}
-}
-\caption{Patch and explanation for fault 115}
-%\vspace{-0.18in}
-\label{fig:explain}
-\end{figure}
-\end{comment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{comment}                                               %
+% \begin{figure}                                                %
+% {\scriptsize                                                  %
+% \begin{code}                                                  %
+% {\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} %
+% ...                                                           %
+% <             vp[1] = OBJECT\_TO\_JSVAL(thisp);               %
+% <         \} else \{                                          %
+% <             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    %
+% <         \}                                                  %
+% ...                                                           %
+% <                   a->avail = (jsuword)sp;                   %
+% <               \}                                            %
+% ...                                                           %
+% ---                                                           %
+% >         RESTORE\_SP(fp);                                    %
+% ...                                                           %
+% {\bf Mutant information:}                                     %
+% Failure repaired by 148 mutants.                              %
+% ...                                                           %
+% \#3: delete statement mutant of jsinterp.c:944:               %
+%                                                               %
+%             vp[1] = OBJECT\_TO\_JSVAL(thisp);                 %
+%         \} else \{                                            %
+% /*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    %
+%         \}                                                    %
+%                                                               %
+% {\bf Coverage change information:}                            %
+% ...                                                           %
+% \#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:      %
+%                    a->avail = (jsuword)sp; /* ADDED */        %
+% \end{code}                                                    %
+% }                                                             %
+% \caption{Patch and explanation for fault 115}                 %
+% %\vspace{-0.18in}                                             %
+% \label{fig:explain}                                           %
+% \end{figure}                                                  %
+% \end{comment}                                                 %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\begin{comment}
-\subsubsection{Error Explanation}
-
-Figure \ref{fig:explain} shows part of the patch for SpiderMonkey fault 115, and the {\bf Coverage} and {\bf Repair} outputs that successfully localize the fault (the 6th coverage change output, and the 3rd mutant output), to give some idea of the information our methods present.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{comment}                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        %
+% \subsubsection{Error Explanation}                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      %
+%                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        %
+%                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        %
+% Figure \ref{fig:explain} shows part of the patch for SpiderMonkey fault 115, and the {\bf Coverage} and {\bf Repair} outputs that successfully localize the fault (the 6th coverage change output, and the 3rd mutant output), to give some idea of the information our methods present.
 For {\bf Coverage}, it is surprising that the sixth highest ranked change only appeared in 5 of 148 repairs.
 This change ranked high out of thousands of changes because it was unusual, appearing for only 10 repairs in the entire SpiderMonkey analysis.
-
-The precise form an explanation based on our techniques should take is not clear.
+%
+%                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        %
+% The precise form an explanation based on our techniques should take is not clear.
 One possibility, suggested by the results here, is a method focused on repairs, but using common coverage changes.
 For each repairing mutant, in rank order, the highest ranking coverage changes for that mutant could also be shown, to give an idea of unusual (and potentially localizing) changes induced by repair.
-\end{comment}
+%
+% \end{comment}                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 For GCC wrong code faults, the baseline chance two test cases shared an underlying fault was 38.97\%.
-Knowing that two test cases shared a repairing mutant raised the chance to 68.5\%, and knowing they disagreed on a fault lowered it to only 19.01\%.
+Knowing that two test cases shared a repairing mutant raised the chance to 68.5\%, and knowing they disagreed on a repair lowered it to only 19.01\%.
 The respective increase and decrease in probability of matching fault compared to baseline was thus 1.75 times greater for matching repairs, and less than 0.5 the chance of being the same fault, given one mismatched repair.
@@
 These can be pruned at will by running them on passing test cases, and removing mutants that fail on any test case.
-\begin{comment}
-Keeping hot mutants for a heavily modified program could be costly (since each mutant must be recompiled, if not re-tested, for each modification to the code), but for stable releases this is a simple approach to keep costs low.
-\end{comment}
 

### 2017-04-30 introduction.tex
small fixes
IEEE format

@@
 
-The key insight of this paper is a hypothesis with numerous possible applications, including a universal fuzzer taming metric that can also provide assistance in debugging.
+The key insight of this paper is based on an idea with numerous possible applications, including a universal compiler fuzzer taming metric that can also provide assistance in debugging.
 
@@
 Our experience, as well as empirical data \cite{GopinathMutants}, suggests that very few compiler faults are due to the kind of simple syntactic changes found in mutants.
-However, this does not matter for {\bf Hypothesis 1};  what matters is that failures, which correspond to faults, can be made to succeed.
+However, this does not weaken the value of the hypothesis; what matters is that failures, which correspond to faults, can be made to succeed.
 When a test case $t$ fails for program $P$ but succeeds for mutant $m$ of $P$ we say that $m$ \emph{repairs} $t$.
@@
 
-These hypothesis do not claim that being repaired by the same mutant perfectly predicts fault equivalence.
+These hypotheses do not claim that being repaired by the same mutant perfectly predicts fault equivalence.
 To understand why that seems highly unlikely, consider why a mutant of a compiler might cause a test case to stop failing.
@@
 
+Because of the difficulty in evaluating Hypotheses 1, 2, and 1a over general programs, in this paper we restrict our attention to the problem of assisting developers using automated compiler testing tools to debug production compilers.
+We therefore evaluate a more modest proposal:  using distance metrics based on the causal approach suggested by these hypotheses, we can improve, in a statistically significant way, the ability to distinguish tests based on which underlying compiler faults they trigger.
+Furthermore, fault localization based on the same metrics can successfully pinpoint the location of some compiler faults sufficiently to be likely to substantially aid debugging.
+
 %\subsection{Contributions}
 
-The contributions of this paper are: hypotheses about how to measure distance between program executions, and practical proposals for using such a metric in fuzzer taming and in novel fault localization and explanation methods.
-Evaluating the effectiveness of these methods also provides experimental evidence as to the validity of our hypotheses.
-Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over previous results \cite{PLDI13}, in multiple ways.
-Our preliminary localization approaches compare favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual}, for the mutation analysis budget in our experiments.
+The contributions of this paper include our hypotheses about how to measure distance between program executions, but most importantly are practical proposals for using such a metric in compiler fuzzer taming and fault localization and explanation.
+Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over best previous results \cite{PLDI13}, in multiple ways.
+Our best preliminary localization approach also compare favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual}, for the mutation analysis budget in our experiments.

### 2017-05-04 localizations.tex
working

@@
 
+In general, we expect {\tt Repair} to be used in the context of
+compiler (or other complex system) fuzzing, where there are numerous
+failing tests, likely from multiple faults.  In such cases, the $f$ to
+be localized is presumably selected by FPF.  However, in case our
+localization approach is used in a setting where single faults are
+expected, we propose an alternative method: let $f$ be an arbitrary test
+test with as few repairing mutants as possible.
+
 The coverage localization also provides a kind of explanation, though

### 2017-05-05 abstract.tex
need to cut a half page

@@
 Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and useful in many test generation and debugging algorithms.
-This paper proposes a metric for measuring distance between executions, and specializes it to an important application, determining similarity of failing test cases, for the purpose of automated fault identification and localization in debugging based on automatically generated compiler tests.
-The metric is based on a causal concept of distance: executions are similar to the degree that changes in the program itself, introduced using mutation operators, cause similar changes in the correctness of the executions.
+This paper proposes a metric for measuring distance between executions, and specializes it to an important application: determining similarity of failing test cases for the purpose of automated fault identification and localization in debugging based on automatically generated compiler tests.
+The metric is based on a causal concept of distance where executions are similar to the degree that changes in the program itself, introduced by mutation, cause similar changes in the correctness of the executions.
 Specifically, if two failing test cases (for the original compiler) become successful due to the same mutant, they are more likely to be due to the same fault.

### 2017-05-05 experiments.tex
need to cut a half page

@@
 
-All mutants were produced using the C mutation tool written by Jamie Andrews \cite{mutant}, used in many previous experiments in mutation analysis and software testing.
+We applied our metric to two of the datasets used in the previous study of compiler fuzzer taming \cite{PLDI13}: faults and tests cases for SpiderMonkey 1.6, Mozilla's JavaScript engine, with tests generated by {\tt jsfunfuzz} \cite{jsfunfuzz}, and \emph{wrong code} faults and test cases for GCC 4.3.0, with tests generated by Csmith \cite{csmith}.
+We did not use the crash faults from that data set, as these are in practice fairly easy to identify by hand (and perfectly identified by previous metrics), and also typically easy to debug.
+
+Our mutants were produced using the tool written by Jamie Andrews \cite{mutant}.
 Andrews' tool applies only four operators: statement deletion, conditional negation, operator replacement, and constant replacement, chosen as a small set that still produces good results for C code.
-Coverage was analyzed with gcov.
-All experiments were performed on a MacBook Pro with 16GB RAM and dual-core 3.1GHz Intel Core i7 (not using the second core); GCC executed on a VirtualBox-hosted Ubuntu 11.04.
+Experiments were performed on a MacBook Pro with 16GB RAM and dual-core 3.1GHz Intel Core i7; GCC executed on a VirtualBox-hosted Ubuntu 11.04.
 For GCC, some test cases from the 2013 PLDI paper no longer failed, presumably due to unknown differences in execution environment, OS, or memory layout.
@@
 
-These data sets, though similar in that both compilers are written in C, provide some interesting variance for testing our hypotheses.
+These data sets, though similar in that both compilers are written in C, provide some interesting variance for testing our metrics.
 The oracle for SpiderMonkey executions is the set of checks built in to {\tt jsfunfuzz} plus the requirement that the execution not crash.
@@
 We applied X-means \cite{xmeans} in an attempt to use clustering with our metrics, but, as with previous results \cite{PLDI13} it did not compare well with FPF, and the runtime was much higher.
-This tends to confirm the conclusion \cite{PLDI13} that clustering is of limited value in fuzzer taming.
+We confirm the conclusion \cite{PLDI13} that clustering is of limited value in fuzzer taming.
 % (at least with X-means) does not work well in a setting with extreme disparity in instance counts for faults.
@@
 Just knowing that two test cases have \emph{one} shared repairing mutant makes it 1.27x more likely that they are the same fault, and knowing they differ for one mutant makes it 0.6 times as likely they are due to the same fault.
-Matching non-repair, however, provides very weak evidence:  only a 34.14\% chance of matching fault, just over baseline probability.
+Matching non-repair, however, provides very weak evidence:  only a 34.14\% chance of matching fault, just over baseline.
 
@@
 
-Table \ref{jstable} shows a comparison of three fault localization methods for the SpiderMonkey faults.
-The first column shows, for each fault, identified by the commit number of the code change that first ``fixed'' the failing tests associated with the fault, its FPF ranking: this is \emph{not} a fault localization ranking, just a way to see that the localization success seems to be somewhat independent of the ease of distinguishing the failures for a fault.
-The next four columns show localization rankings.
+Table \ref{bothtable} shows a comparison of three fault localization methods for the SpiderMonkey (and GCC) faults.
+The first column shows, for each fault, an ID.
+The next three columns show localization rankings.
 A dash for a column means that localization did not rank any faulty statements, or assigned all faulty statements suspiciousness of 0.0.
-The four rankings are: the {\bf Coverage} localization, the {\bf Repair} localization, the MUSE \cite{MUSE} localization, and the MUSEUM \cite{multilingual} localization, which uses the same formula as MUSE, but works better for multiple faults because it uses only one failure.
+The three rankings are: our {\bf Repair} localization, the MUSE \cite{MUSE} localization, and the MUSEUM \cite{multilingual} localization, which uses the same formula as MUSE, but works better for multiple faults because it uses only one failure.
 The MUSE/MUSEUM formulas normally make use of information from passing tests as well: when mutating a statement makes a passing test case fail, it makes the statement \emph{less} suspicious, by a weighted amount.
@@
 the more extensive Proteum \cite{Proteum} operators used in MUSE/MUSEUM's evaluations.
-Even with these restrictions, running all mutants on passing tests took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.
+Even with these restrictions, running all mutants on passing tests until at least a single kill was observed took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.
 
+\comment{
 Results are reported as absolute ranks, rather than more sophisticated \cite{MUSE} localization evaluations because our localizations do not assign suspiciousness scores, only rankings, and in accord with the proposal of Parnin and Orso that only localizations ranking a fault in the top few statements may be useful to users \cite{AutoHelp}.
@@
 Blind, unaided pointers to portions of compiler code without more semantic information about why the statement is suspicious seem unlikely to work well, based on our experience with complex compiler faults, and the results of Parnin and Orso \cite{AutoHelp}.
+}
 
+\comment{
 \begin{table}
@@
 {\scriptsize
-\begin{tabular}{|c||c||c|c|c|c|}
+\begin{tabular}{|c||c|c|c|}
 \hline
-& & \multicolumn{4}{|c|}{Localization Rank} \\
+%& \multicolumn{3}{|c|}{Localization Rank} \\
+%\hline
+Fault & Repair & MUSE & MUSEUM \\
 \hline
-Fault & FPF & Coverage & Repair & MUSE & MUSEUM \\
-\hline
-880 & 1 & 2,255 & - & - & - \\
-95 & 6 & 720 & - & 173 & -\\
-1294 & 9 & 190 & 1 & 22 & 4 \\
-60 & 10 & - & - & - & -\\
-1543 & 15 & 18 & 19 & 118 & 26\\
-1172 & 19 & 46 & 28 & 87 & 23\\
-1561 & 106 & - & - & - & - \\
-115 & 192 & 6 & 3 & 133 & 5 \\
+880 & - & - & - \\
+95 & - & 173 & -\\
+1294 & 1 & 22 & 4 \\
+60 & - & - & -\\
+1543 & 19 & 118 & 26\\
+1172 & 28 & 87 & 23\\
+1561 & - & - & - \\
+115 & 3 & 133 & 5 \\
 \hline
@@
 \end{table}
+}
 
@@
 
-The first thing to note about results is that Table \ref{jstable} only includes 8 of the 28 SpiderMonkey faults.
-This is due to a problem discussed in the paper introducing this data set:  producing ground truth patches for a large application, in the sense of finding a minimal, clearly fault-fixing code change that can be back-ported to the original code is difficult \cite{PLDI13}.
+Table \ref{bothtable} only includes 8 of the 28 SpiderMonkey faults.
+This is due to a problem with the data set:  producing ground truth patches for a large application, in the sense of finding a minimal, clearly fault-fixing code change that can be back-ported to the original code is difficult \cite{PLDI13}.
 While we believe the 28 faults identified in the PLDI 2013 data set are correct, it is very difficult to produce a valid patch of version 1.6 that captures the fix for most of these faults.
@@
 The evaluation of localization therefore only examines the 8 faults for which we could be reasonably certain that the patch to 1.6 was correct and characterized the fault in question accurately.
-The smallest patch required modification of two statements, and was thus larger than any mutants applied.
-In no case was the patch equivalent to one of the mutants.
+In no case was the patch equivalent to one of the mutants (the smallest patch modified two statements).
 
 No method performs extremely well --- for half the faults, no methods produced what we consider a useful localization.
-Compiler faults are hard to debug, and any help is useful, but it seems unlikely developers will really benefit from a localization when it does not rank at least one faulty statement in the first 20 to 30 statements.
+Compiler faults are hard to debug, and any help is useful, but it seems unlikely developers will really benefit from a localization when it does not rank at least one faulty statement in the first 25-30 statements.
 By this measure, MUSE only provides a helpful localization once, and performs worst of the methods.
-{\bf Coverage} performs very badly for three faults, does not manage to be useful by our definition for another, and in the remaining two cases provides a good localization, in one case the best.
 {\bf Repair} and MUSEUM both perform well, with {\bf Repair} slightly better.
@@
 
+\comment{
 \begin{table}
@@
 {\scriptsize
-\begin{tabular}{|c||c||c|c|c|c|}
+\begin{tabular}{|c||c||c|c|}
 \hline
-& & \multicolumn{4}{|c|}{Localization Rank} \\
+%& & \multicolumn{4}{|c|}{Localization Rank} \\
+%\hline
+Fault & Repair & MUSE & MUSEUM \\
 \hline
-Fault & FPF & Coverage & Repair & MUSE & MUSEUM \\
-\hline
-139094 & 1 & 44,561 & - & 157 & -\\
-133004 & 2 & 75 & 1 & 99 & 14\\
-158782 & 4 & - & - & 146 & -\\
-156496 & 5 & 3,067 & - & - & -\\
-143677 & 7 & 186 & - & 109 & -\\
-158555 & 8 & 490 & - & - & -\\
-134321 & 10 & - & - & - & -\\
-133940 & 14 & 260 & 4 & 158 & 18\\
-141195 & 15 & 465 & - & - & -\\
-156795 & 17 & 19 & - & - & -\\
-139709 & 19 & 43,903 & 10 & 167 & 6\\
-155698 & 22 & 85 & 1 & 172 & 195\\
-138646 & 24 & 2 & - & - & -\\
-140795 & 25 & 302 & 34 & 170 & 125\\
-136501 & 44 & 223 & 5 & 159 & 31\\
-134322 & 45 & 639 & - & - & -\\
+139094 & - & 157 & -\\
+133004 & 1 & 99 & 14\\
+158782 & - & 146 & -\\
+156496 &  - & - & -\\
+143677 & - & 109 & -\\
+158555 & - & - & -\\
+134321 & - & - & -\\
+133940 & 4 & 158 & 18\\
+141195 & - & - & -\\
+156795 & - & - & -\\
+139709 & 10 & 167 & 6\\
+155698 & 1 & 172 & 195\\
+138646 & - & - & -\\
+140795 & 34 & 170 & 125\\
+136501 &  5 & 159 & 31\\
+134322 & - & - & -\\
 \hline
@@
 \end{table}
+}
 
-Table \ref{gcctable} shows localization results for GCC 4.3.0.
+\begin{table}
+\centering
+{\scriptsize
+\begin{tabular}{|c||c|c|c||c||c|c|c|}
+\hline
+%& & \multicolumn{4}{|c|}{Localization Rank} \\
+%\hline
+\multicolumn{4}{|c||}{SpiderMonkey}&\multicolumn{4}{|c|}{GCC}\\
+\hline
+ID& Repair & MUSE & MUSEUM & ID & Repair & MUSE & MUSEUM \\
+\hline
+1 & - & - & - &         1 & - & 157 & -\\
+2 & - & 173 & - &     2 & 1 & 99 & 14\\
+3 & 1 & 22 & 4 &       3 & - & 146 & -\\
+4 & - & - & - &         4 &  - & - & -\\
+5 & 19 & 118 & 26 & 5 & - & 109 & -\\
+6 & 28 & 87 & 23 &   6 & - & - & -\\
+7 & - & - & - &         7 & - & - & -\\
+8 & 3 & 133 & 5 &     8 & 4 & 158 & 18\\
+& & & & 9 & - & - & -\\
+& & & & 10 & - & - & -\\
+& & & & 11 & 10 & 167 & 6\\
+& & & & 12 & 1 & 172 & 195\\
+& & & & 13 & - & - & -\\
+& & & & 14 & 34 & 170 & 125\\
+& & & & 15 &  5 & 159 & 31\\
+& & & & 16 & - & - & -\\
+\hline
+\end{tabular}
+}
+\caption{Fault localization for compiler faults}
+\label{bothtable}
+\end{table}
+
+Table \ref{bothtable} shows localization results for GCC 4.3.0.
 Again, only some faults were deemed to have a strong enough ground truth patches for localization evaluation.
-MUSE performed poorly, with no useful (by the standard of having at least one faulty statement in the top 30) localizations.
-{\bf Coverage} produced one very good localization, and was the only method to localize that fault at all, but usually performed poorly.
+MUSE performed poorly, with no useful (by the standard of having at least one faulty statement in the top 25) localizations.
 MUSEUM provided 3 useful localizations, but no very high quality localizations.
 Only {\bf Repair} performed very well, with useful localizations for 5 faults, and the fault was in the top 5 statements for 4 of these.
+
+\input{othertools}
 
@@
 This seems to be true, when such a repair exists, and results in {\bf Repair} either providing a ``useful'' localization or no localization at all, which user studies and surveys suggest is the ideal behavior due to the cost of false positives in localization \cite{AutoHelp,Kochhar}.
-If any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
+For the compiler faults, if any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
 MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.
 Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this is unlikely due to the nature of the repairs.
-For the MUSEUM data set, there are no cases with multiple faults for our method to use to further distinguish repairs, so its similarity to MUSEUM in results is not surprising (in this context, our approach is essentially MUSEUM modified to throw out any repairs that cause previously passing tests to fail).
 
+\comment{
 Finally, in contrast to many previously reported results in fault localization, typically over much simpler programs and faults than subtle compiler semantics bugs, all methods frequently failed to provide a useful localization at all.
 This is presumably due to the difficulty of compiler bugs:  our approach, and MUSEUM, performed very well on complex multi-lingual faults in the MUSEUM data set.
-In one sense, {\bf Coverage} is the most intriguing method presented here.
+}
+\comment{ In one sense, {\bf Coverage} is the most intriguing method presented here.
 While it did not perform nearly as well as {\bf Repair}, it did manage, in many cases, to rank \emph{some} faulty line highly relative to the extremely large number of statements executed in each failing GCC run (on the order of 50K-100K+ statements).
@@
 There is much to be done if automated debugging is ever to make the lives of compiler developers easier in most cases \cite{AutoHelp}.
+}
 
@@
 As we show, MUSE and MUSEUM \emph{can}  be used without passing tests or these mutants, and MUSEUM still produces good results even with a weak oracle and few operators.
-MUSE, MUSEUM, and our approach all run analysis without coverage instrumentation; we re-run repaired tests under gcov when using the {\bf Coverage} localization.
 

### 2017-05-05 introduction.tex
need to cut a half page

--- 2017-01-28 introduction.tex
@@
 
-The goal of this paper, an effective metric for computing distance between program executions (in particular, failing test cases), is motivated in large part by progress in automated testing.
-Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{We claim testing/debugging complex systems software is well represented by compiler testing as an instance.}.
+The proposed techniques in this paper are motivated in large part by progress in automated testing.
+Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in modern optimizing compilers.
 Unfortunately, the blessing of many failing test cases can become a curse for a developer.
@@
 Repairs can also provide  information for understanding and debugging discovered faults.
-The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.
-Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).
+\comment{The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.
+Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).}
 
@@
 
-The key insight of this paper is a hypothesis with numerous possible applications, including a universal fuzzer taming metric that can also provide assistance in debugging.
-
-\begin{quote}
-{\bf Hypothesis 1:} If two test cases that fail can be made to succeed by the same syntactic modification to the original program $P$, this provides evidence that they fail due to the same fault.
-\end{quote}
-
-The core idea is hardly controversial:  most developers determine if they have fixed all faults in a set of failing test cases by re-running the tests after they fix at least one fault.
+The key insight of this paper is based on an idea with numerous possible applications: \emph{if two test cases that fail can be made to succeed by the same small modification to the original program $P$, this provides evidence that they fail due to the same fault}.
+The idea is hardly controversial:  most developers determine if they have fixed all faults in a set of failing test cases by re-running the tests after they fix at least one fault.
 As a method for fuzzer taming or debugging assistance, however, this is useless.
-Fortunately, developer-provided patches are not the only source of syntactic changes to a program.
-Mutation analysis \cite{demillo1978hints,budd1980theoretical} applies large numbers of syntactic changes to a program, usually to evaluate a test suite \cite{mutant,justmutants}.
+Fortunately, developer-provided patches are not the only source of changes to a program.
+Mutation analysis \cite{demillo1978hints,budd1980theoretical} applies large numbers of small syntactic changes to a program, usually to evaluate a test suite \cite{mutant,justmutants}.
 Some recent work has also considered these changes as a source of fixes or localizations for program faults \cite{achour,multilingual,MUSE,DebroyMutant}.
 Our experience, as well as empirical data \cite{GopinathMutants}, suggests that very few compiler faults are due to the kind of simple syntactic changes found in mutants.
-However, this does not matter for {\bf Hypothesis 1};  what matters is that failures, which correspond to faults, can be made to succeed.
+However, what should matter is that failures, which correspond to faults, can be made to succeed.
 When a test case $t$ fails for program $P$ but succeeds for mutant $m$ of $P$ we say that $m$ \emph{repairs} $t$.
@@
 
-Hypothesis 1 is in fact a consequence of a more general (and difficult to investigate) hypothesis.
+This idea is in fact a consequence of a more general (and difficult to investigate) hypothesis.
 What is interesting about a program execution?
@@
 What, then, makes two executions similar?
-Our proposal is to say that:
-
-\begin{quote}
-{\bf Hypothesis 2:} Two executions of $P$ are similar to the degree that changes to $P$ produce the same changes in correctness of the executions.
-\end{quote}
+Our proposal is that \emph{two executions of $P$ are similar to the degree that changes to $P$ produce the same changes in correctness of the executions.}
 
 This leads to the idea of comparing executions by their response to mutants.
-In the presence of multiple correctness properties, any two executions can be compared, not just two failures, since a mutant can cause a failure to violate additional or different properties.
+In the presence of multiple correctness properties, in principle any two executions can be compared, not just two failures, since a mutant can cause a failure to violate additional or different properties.
+However, for practical purposes the most interesting implication is that if two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.
+
+\comment{
 %Where there are multiple, independent, correctness properties, this even allows us to compare failing and successful executions, since a mutant can cause a failure to violate a different property.
@@
 This is {\bf Hypothesis 1a:} If two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.
+}
 
@@
 
-These hypothesis do not claim that being repaired by the same mutant perfectly predicts fault equivalence.
+This not a claim that being repaired by the same mutant perfectly predicts fault equivalence.
 To understand why that seems highly unlikely, consider why a mutant of a compiler might cause a test case to stop failing.
@@
 Optimizations are generally guarded by conditional switches --- in some cases, so that the optimization can be turned off by a user, in other cases when a property of the code in question prevents the optimization.
-If two failures become successes due to turning off an optimization, it is certainly some level of evidence they are the same fault, compared to the baseline probability.
+If two failures become successes due to turning off an optimization, it is presumably \emph{some} level of evidence they are the same fault, compared to the baseline probability.
 Intuitively, the evidence provided seems much stronger than the fact that, e.g., the two test cases execute a shared line of code \cite{RepsSpectra}.
@@
 
+This paper investigates the utility of using distance metrics based on the causal approach suggested above.
+We show that such a metric can improve, in a statistically significant way, the ability to distinguish tests based on which faults they trigger.
+Furthermore, fault localization based on the same metrics can successfully pinpoint the location of some compiler faults sufficiently to be likely to substantially aid debugging, and, when applicable, performs very well for some faults studied in recent papers on mutant-based fault localization.
+
 %\subsection{Contributions}
 
-The contributions of this paper are: hypotheses about how to measure distance between program executions, and practical proposals for using such a metric in fuzzer taming and in novel fault localization and explanation methods.
-Evaluating the effectiveness of these methods also provides experimental evidence as to the validity of our hypotheses.
-Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over previous results \cite{PLDI13}, in multiple ways.
-Our preliminary localization approaches compare favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual}, for the mutation analysis budget in our experiments.
+The contributions of this paper include our ideas about how to measure distance between program executions, but more importantly some practical proposals for using such a metric in compiler fuzzer taming and fault localization.
+Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over best previous results \cite{PLDI13}, in multiple ways.
+Our best preliminary localization approach also compares favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual,Metallaxis}, albeit with a more limited applicability (but a lower cost).

### 2017-05-05 localizations.tex
need to cut a half page

--- 2017-01-29 localizations.tex
@@
 information about the likely location of a fault in a program's source
-code.  Error explanation \cite{GroceError} also provides information
+code.  \comment{Error explanation \cite{GroceError} also provides information
 about the causal structure of a fault --- how it can be avoided, what
-triggers it --- a kind of ``fault story.''  Parnin and Orso recently
+triggers it --- a kind of ``fault story.''}  Parnin and Orso recently
 suggested that fault localization methods are not actually providing a
@@
 
+\comment{
 By definition, when a mutant repairs a test case, it changes the
@@
 %r
-
-The repairing mutants themselves can also be used for localization and
+}
+Repairing mutants can be used for localization and
 explanation.  While few will be equivalent to actual fixes, they
 will sometimes overlap the incorrect code, and will often be closely
-related to incorrect code, semantically.  Since many failures are
-fixed by a large number of mutants, it is important to rank the
+related to incorrect code, semantically.  Since some tests are
+repaired by a large number of mutants, it is important to rank the
 mutants to produce a localization.  Based on the onion-ring model proposed above, this is simple: each mutant is ranked as
@@
 
+An advantage of this method over some statistical approaches to
+localization is that it is not potentially confused in
+settings with multiple faults.  Each localization/explanation is based
+on a single failing test case, which in most cases fails due to only one
+fault.  In fact, other faults (and their failures) theoretically \emph{help} {\bf Repair} rank the mutants.
+
+In general, we expect {\bf Repair} to be used in the context of
+compiler (or other complex system) fuzzing, where there are numerous
+failing tests, likely from multiple faults.  In such cases, the $f$ to
+be localized is presumably selected by FPF.  However, in case our
+localization approach is used in a setting where single faults are
+expected, we propose an alternative method: let $f$ be an arbitrary
+test with as few repairing mutants as possible.
+
+\comment{
 The coverage localization also provides a kind of explanation, though
@@
 on a single failing test case, which in most cases fails due to only one
-fault.  In fact, other faults help both methods rank localizations.
+fault.  In fact, other faults help both methods rank localizations.}

### 2017-05-05 metrics.tex
need to cut a half page

@@
 where a 1 bit indicates that the mutant in question repairs this test
-case.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
+case\footnote{By \emph{repairing} a failing test we mean that a mutant
+causes the test to pass, and the mutant does not cause
+other, previously passing tests, to fail.}.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
 by 100 mutants.  If these two test cases share 90 common repairing
@@
 This metric, the portion of mismatches over bits that are 1 in either
-vector, matches our intuitions about the above comparisons\footnote{In Section \ref{conc} we discuss possible uses of $d$
+vector, matches our intuitions about the above comparisons\footnote{In Section \ref{conc} we discuss uses of $d$
 requiring comparison of both failing and passing test cases:  in this 

### 2017-05-05 othertools.tex
need to cut a half page

\subsection{Repair Localization for Non-Compiler Faults}

The {\bf Repair} localization was designed for use in compiler (or at least complex system software) fuzzing.  It is not proposed (or at least, not here evaluated) as a general-purpose fault localization method.  However, it is possible to compare {\bf Repair} with other methods, using their full mutation and experimental set, to provide a more apples-to-apples comparison than the evaluation vs. MUSE and MUSEUM over compiler faults.  We obtained the data from two evaluations of mutant-based fault localization methods:  the set of faults used to evaluate MUSEUM \cite{multilingual}, and a set of CoreBENCH \cite{CoreBENCH} faults used in an evaluation of MUSE and Metallaxis-FL by Chekal et. al \cite{Papadakis} (we were unable to obtain the original MUSE dataset, thus far).  These data sets are not ideal for evaluating {\bf Repair}: for all but 10 of the faults (all from the CoreBENCH set) there is only a single failing test case in the data.  {\bf Repair} for single failing tests is essentially MUSEUM without use of passing tests, except to reject any mutation that breaks a passing test.  {\bf Repair} is meant to operate in the case where a program either has numerous failing tests associated with open bugs (e.g., a typical production compiler or complex system software program) or where a program has just been fuzzed for the first time, so has numerous failures representing different faults.  The distance metric lets {\bf Repair} make use of these other faults, by enabling it to focus on repairing mutants most relevant to the failing test.

\subsubsection{MUSEUM Bug Results}  For the six faults used in the original MUSEUM paper (all with a single failing test), {\bf Repair} gives the same (perfect) localization for three faults.  One of the other three faults (Bug2) has no repairing mutants (that do not break some passing test) so {\bf Repair} provides no localization information at all.  For the remaining two faults, Bug5 and Bug6, {\bf Repair} gives best rankings of 12 and 9, respectively, compared to 8 and 3 for MUSEUM.

\subsubsection{CoreBENCH Results} For the full data set in the paper \cite{Papadakis}, {\bf Repair} does not perform as well as MUSE and Metallaxis-FL, as expected.   It is only able to  rank a faulty line of code for 12 of 30 faults.  However, for these faults, it performs very well indeed, perfectly localizing 7 of the 12.  {\bf Repair} is uniquely best for 5 of the 12, and is tied for best for another 6; for the remaining fault, it performs better than Metallaxis-FL but worse than MUSE.  Table \ref{otherbugs} shows the results.  The fourth column shows, for those bugs where {\bf Repair} did not rank any faulty statement, how many statements {\bf Repair} proposed for a user to examine.  Note that in most cases if {\bf Repair} was not helpful it also provided little output to waste a user's time: it only once presented more than 3 statements to examine.   In practice, we only suggest using {\bf Repair} at all in cases where 1) there is at least one repairing mutant (not killed by any passing test) for a failing test and 2) there are multiple failing tests (whether from one or multiple faults).  When the Lines column shows a bold 0, it indicates there were no repairs.  The 7 faults meeting both requirements are bolded.  For 4 of these, {\bf Repair} produces a perfect localization, and for the other three, it produces no more than 2 non-faulty statements to examine.

\begin{table}
\centering
{\scriptsize
\begin{tabular}{|c|c||c|c||c|c|}
\hline
& \#Failing & & & & \\
Fault & Tests & Repair & Lines & Metallaxis-FL & MUSE\\
\hline
{\bf Coreutils 1} & 2 & 1 & - & 1 & 1 \\
Coreutils 2 & 3 & - &  {\bf 0} & 86 & 186 \\
Coreutils 3 & 1 & 25 & - & 47 & 27 \\
Coreutils 4 & 1 & - &  {\bf 0} & 28 & 431 \\
{\bf Coreutils 5} & 5 & - &  2 & 4 & 7 \\
Coreutils 6 & 1 & - &  1 & 5 & 206 \\
Coreutils 7 & 1 & 1 & - & 5 & 9 \\
Coreutils 8 & 1 & - &  1 & 22 & 160 \\
{\bf Coreutils 9} & 5 & - &  1 & 10 & 157 \\
Coreutils 11 & 1 & - &  1 & 2 & 1 \\
Coreutils 12 & 2 & - & {\bf 0} & 6 & 9 \\
Coreutils 13 & 1 & - & {\bf 0} & 905 & 833 \\
Coreutils 14 & 1 & 14 & - & 19 & 11 \\
{\bf Coreutils 15} & 2 & 1 & - & 1 & 6 \\
Coreutils 16 & 1 & - & {\bf 0} & 569 & 447 \\
{\bf Coreutils 17} & 3 & - &  2 & 37 & 195 \\
Coreutils 18 & 1 & - &  3 & 27 & 3 \\
Coreutils 19 & 1 & 1 & - & 3 & 1 \\
{\bf Coreutils 20} & 7 & 1 & - & 12 & 3 \\
Coreutils 21 & 2 & - & {\bf 0} & 20 & 191 \\
{\bf Coreutils 22} & 2 & 1 & - & 6 & 2 \\
Findutils 27 & 1 & - &  9 & 12 & 6 \\
Findutils 32 & 1 & - &  1 & 73 & 70 \\
Findutils 33 & 1 & - &  3 & 10 & 30 \\
Findutils 35 & 1 & 1 & - & 1 & 1 \\
Findutils 36 & 1 & 9 & - & 9 & 19 \\
Findutils 37 & 1 & 11 & - & 22 & 34 \\
Grep 46 & 1 & - &  2 & 7 & 10 \\
Grep 47 & 1 & 4 & - & 4 & 4 \\
Grep 48 & 1 & - & {\bf 0} & 26 & 497 \\

\hline
\end{tabular}
}
\caption{Fault localization for CoreBENCH faults}
%\vspace{-0.2in}
\label{otherbugs}
\end{table}

### 2017-05-05 conclusions.tex
could ship in theory

@@
 \label{conc}
-
+\comment{
 When are two failing test cases due to the same fault?  This paper
@@
 has long been tied to a notion of similarity of
-executions \cite{NearNeighbor,GroceError}.  By using program changes
+executions \cite{NearNeighbor,GroceError}. } By using program changes
 (produced by mutants) as
 causes to determine distance, it is possible to improve on efforts
-to identify the set of faults in a large set of redundant failures
+to identify the set of faults in a large set of redundant compiler failures
 \cite{PLDI13} and discover the source code locations of faults
@@
 compiler debugging can consistently be made less onerous through automated
-assistance.  One idea is to combine the metric approach and the ideas
+assistance.  \comment{One idea is to combine the metric approach and fall
+the ideas
 in MUSE and MUSEUM, or other mutant-based localizations such as
@@
 compare our localizations with the Metallaxis
-\cite{Metallaxis,Papadakis} approach.
+\cite{Metallaxis,Papadakis} approach.}
 

### 2017-05-05 experiments.tex
could ship in theory

@@
 Table \ref{bothtable} shows a comparison of three fault localization methods for the SpiderMonkey (and GCC) faults.
-The first column shows, for each fault, an ID.
+The first column is a fault ID.
 The next three columns show localization rankings.
@@
 However, there may be other mutants that 1) repair a failure and 2) are killed by some test case: these could, if enough different ones repaired the same faulty statement, improve MUSE/MUSEUM results, though in practice this is extremely unlikely.
-We reject such mutants in part to keep costs low, and in part because we think that the causal information contained in such mutants, that repair a failure but also break some passing test, is problematic.
+We reject such mutants in part to keep costs low, and in part because we think that the causal information contained in such mutants, that ``fix'' a failure but also break some passing test(s), is problematic.
 They seem likely to be less useful as explanations of the causes of a failure, since they do not impact the program semantics in a way that is only known to be beneficial, and could potentially even mislead a user if used as explanations.
-Still, it is most useful to compare MUSEUM/MUSE and our approach on the cases where all methods produce a localization.
+\comment{Still, it is most useful to compare MUSEUM/MUSE and our approach on the cases where all methods produce a localization.}
 
@@
 the more extensive Proteum \cite{Proteum} operators used in MUSE/MUSEUM's evaluations.
-Even with these restrictions, running all mutants on passing tests until at least a single kill was observed took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.
+\comment{Even with these restrictions, running all mutants on passing tests until at least a single kill was observed took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.}  Below we provide some comparisons with MUSEUM and MUSE fault localization in the context of a full mutation result matrix using their preferred set of operators.
 
@@
 If a user gives up after a plateau of size 5, our approach again lets a user examine 18 faults over 25 tests.
-Function coverage provides only 16 faults over 31 tests.
+Function coverage only yields 16 faults over 31 tests.
 
@@
 Table \ref{bothtable} shows localization results for GCC 4.3.0.
-Again, only some faults were deemed to have a strong enough ground truth patches for localization evaluation.
+Again, only some faults were deemed to have strong enough ground truth patches for evaluation.
 MUSE performed poorly, with no useful (by the standard of having at least one faulty statement in the top 25) localizations.
@@
 However, the comparison is with the very \emph{best} curve chosen after running more than 16 different metrics.
-In practice, users do not know ground truth to rank curves, thus Chen et al.
+In practice, users simply do not know ground truth to rank curves, thus Chen et al.
 \cite{PLDI13} do not really give a practical approach to distance metric selection in their PLDI work.
 The best methods for different subjects varied widely, even in such difficult-to-understand ways as less-fine-grained coverage providing better results in some cases, but worse in other cases.
-In practice, their work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique.
-It also included methods such as comparing output signatures that are inapplicable to differential testing in most cases, the most critical compiler application.
-This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work.
-Moreover, the improvement is, for our subject programs, reliable:  choosing random subsets of the full data set, the difference in mean APFD is around 3\%, and statistically significant.
+In practice, their work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique, since trying multiple metrics is impractical, since a user has to examine each curve.
+\comment{It also included methods such as comparing output signatures that are inapplicable to differential testing in most cases, the most critical compiler application.}  This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work; our improvements over any single method applied to both problems are considerably larger.
+Moreover, the improvement is, for our subject programs, reliable:  choosing random subsets of the full data set, the difference in mean APFD from best-previous method is around 3\%, and statistically significant.
 
-For fault localization and error explanation, the results show possible improvement on state-of-the-art approaches, in the absence of complete information from mutant analysis over passing tests (which can be very expensive to obtain).
+For fault localization and error explanation, the results show possible improvement on state-of-the-art mutation-based approaches.
 A practical impact of the results reported is that we suggest users of our techniques examine only the first few (at most 10, and we propose as few as 5) mutants and coverage-changing statements, and ignore localization if none of these results are helpful.
@@
 
+There are two ways to consider the performance of {\bf Repair}.
+First, it is using information that other techniques are not using.
+If each failure was only repaired by one mutant, MUSEUM and {\bf Repair} would both rank that mutant maximally, and we speculate that it would usually be faulty code.
+But such cases are very rare in compilers:  for SpiderMonkey the mean/median number of mutants repairing each failure were 90.8 and 78, respectively, and for GCC the mean and median were 54 and 43.
+{\bf Repair} can distinguish repairing mutants in fine-grained ways that MUSE and MUSEUM cannot by relying on the onion-ring structure:  when a repair also repairs failures that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults, and so ranks it lower than more ``relevant'' repairs.
+
+The other way to think about {\bf Repair} for the non-compiler faults especially, is as a very \emph{cautious} localization.
+It only uses mutants that repair a failure, and, because it is clear that, in general, failing tests contain far more information about faults than passing tests, it only uses information from failures (other than using passing tests to prune mutants that kill some passing tests, again a principle of discarding potentially confusing information).
+This results in {\bf Repair} either providing a ``useful'' localization or almost no localization information at all to mislead a user, for the non-compiler faults, which user studies and surveys suggest is the ideal behavior due to the cost of false positives in localization \cite{AutoHelp,Kochhar}.
+
+\comment{
  Why does {\bf Repair} perform best in this setting?
 It is impossible to be sure, but one  possibility is power to distinguish repairs.
-If each failure was only repaired by one mutant, MUSEUM and {\bf Repair} would both rank that mutant maximally, and we speculate it would usually be faulty.
-But such cases are rare:  for SpiderMonkey the mean/median number of mutants repairing each failure were 90.8 and 78, respectively, and for GCC the mean and median were 54 and 43.
 If there are many repairs for a failure, how can the most important ones be distinguished?
@@
 For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.
-We applied our repair localization to the MUSEUM data set, using the full Proteum mutant set and added MUSEUM-specific mutants, and for all but one of the 8 bugs in the data set, our results were equivalent to those for MUSEUM (in one case our approach fails to produce a localization).
+We applied our repair localization to the MUSEUM data set, using the full Proteum mutant set and added MUSEUM-specific mutants, and for all but one of the 8 bugs in the data set, our results were equivalent to those for MUSEUM (in one case our approach fails to produce a localization).}
 
-{\bf Repair} can distinguish more repairing mutants by relying on the onion-ring structure:  when a repair also repairs test cases that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults.
-The hypothesis behind {\bf Repair} is that \emph{repairs specific to the fault causing a failure are most likely to modify code involved in the fault itself}.
-This seems to be true, when such a repair exists, and results in {\bf Repair} either providing a ``useful'' localization or no localization at all, which user studies and surveys suggest is the ideal behavior due to the cost of false positives in localization \cite{AutoHelp,Kochhar}.
-For the compiler faults, if any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
+{ For the compiler faults, if any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
 MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.
@@
 
-Mutation analysis is not cheap.
-The experiments in this paper were run over a long period of time.
-However, there are numerous ways to reduce costs.
-First, the primary cost is compiling the mutated versions of the SUT.
+The compiler experiments required a large computational budget.
+However, our approach is, by design, cheaper than MUSE or Metallaxis-FL, in that, for a program with up-to-date mutation testing results (just knowing which mutants are killed, not a full matrix), it only requires executing mutants over failing tests.
 Running each test case under each mutated version of the compiler is cheap: each execution requires around 0.05 seconds for SpiderMonkey, and 0.12 seconds for GCC, on average; moreover, executions can be done in parallel, most failures do not cover most mutants, and vice versa, so the total number of repair checks needed is far less than the product of mutants and failures, and can be spread over many machines.
-The practical approach to fuzzer taming for a large, ongoing project is likely to maintain a set of ``hot'' pre-compiled mutants.
+\comment{The practical approach to fuzzer taming for a large, ongoing project is likely to maintain a set of ``hot'' pre-compiled mutants.
 These can be pruned at will by running them on passing test cases, and removing mutants that fail on any test case.
-
-Second, and most critically, most program mutants never need to be considered as repairs, because we only consider mutants that cause no passing test case to fail.
-MUSE and MUSEUM use kills by passing tests to help rank such tests as repairs, which is finer grained but more expensive.
-As we show, MUSE and MUSEUM \emph{can}  be used without passing tests or these mutants, and MUSEUM still produces good results even with a weak oracle and few operators.
+}
 
@@
 However, the improvement in some cases was \emph{much} smaller than for the full set, and varied considerably (in some cases it was up to 2\% better than our full results).
-Fault localization results were considerably worse, however (with relative rankings preserved) for both our techniques and MUSE/MUSEUM.
-Finally, we have preliminary results indicating that a guided-sampling approach may provide good fuzzer taming (if not fault localization):  compute vectors over a small constant-sized \cite{RahulISSRE} sample of mutants, then report highly distinct failing tests.
+Fault localization results were much worse, however (with relative rankings preserved) for both our techniques and MUSE/MUSEUM.
+\comment{Finally, guided sampling may provide good fuzzer taming (if not fault localization):  compute vectors over a small constant-sized \cite{RahulISSRE} sample of mutants, then report highly distinct failing tests.
 Other metrics, such as coverage-based comparisons, could provide hints as to likely distinct tests (and thus mutants covered by them) to focus on first.
@@
 After repairing some of these, choose mutants to evaluate that are covered by multiple tests with few or no repairing mutants (since such tests are the ones most likely to be distinct but not yet identified).
-Such an approach could be quite cost-effective, requiring only a relatively small number of mutant evaluations before providing useful results.
+Such an approach could be quite cost-effective, requiring only a relatively small number of mutant evaluations before providing useful results.}
 
@@
 
-The largest threat to validity is that this paper's conclusions rely on only two data sets, used in a previous paper.
-The fault ground truths are possibly imperfect, due to the complex change history, and it was not possible to produce good patches for all faults, limiting our analysis of fault localization.
-Generalizing to all compilers, or to software systems that are not compilers, would be unwarranted.
-However, for compilers, we believe the underlying arguments supporting the probability of discovering many meaningful repairing mutants should hold in the general case.
-There is no particular reason to believe that Csmith and {\tt jsfunfuzz} produce unusual random test cases, and both have often been used to detect faults in compilers other than the two considered in this paper.
-For other kinds of software, more experiments are definitely warranted.
+The largest threat to validity is that this paper's conclusions rely on only two data sets for compilers.
+The fault ground truths are possibly imperfect, due to the complex change history, and it was not possible to produce good patches for all faults, limiting our analysis of compiler fault localization.
+The comparison with other fault localization methods is also limited:  our compiler experiments limit the mutation budget in ways that may weaken MUSE and MUSEUM effectiveness, and the results using fault and mutant data sets from other studies provide few faults, no multi-fault problems, and very few multi-failure problems, limiting {\bf Repair} effectiveness.
 It is also possible that faults in the framework used for experiments introduced errors into the data.
 In order to check for these problems, we will make the raw data for a re-analysis available on request.
-
-Second, the comparison with MUSE and MUSEUM is limited.
-MUSE is not really intended for multiple faults.
-With our oracles, MUSE and MUSEUM may need the full results from passing tests, and some mutant we omit might be both highly ranked and a good localization.
-A full comparison would include a complete analysis of passing tests also, but would likely require far more than 10x the computation time.
-Finally, these experiments use Andrews' mutation operators, while MUSE/MUSEUM experiments use Proteum \cite{Proteum}, which has a much larger set of mutation operators, and their results indicate that the large set of operators is important.
-However, running all Proteum GCC mutants over all passing tests would likely be prohibitively expensive.

### 2017-05-05 relatedwork.tex
could ship in theory

@@
 \cite{multilingual} on MUSEUM and Moon et al.
-\cite{MUSE} on MUSE provides a good summary of other work along these lines, and is perhaps most similar to ours in assumptions, in particular ``Conjecture 1'' that a test case that used to fail on a program is more likely to pass on a mutation of a faulty statement.
-While we suggest some localizations as a welcome and interesting side effect of our distance metric and FPF analysis, our primary goal is the discovery of all faults.
-Much fault localization work is largely (in many cases exclusively) based on single-fault scenarios; e.g., the MUSE \cite{MUSE} evaluation includes only 2 of 14 scenarios with multiple faults, and those scenarios include only a few faults.
+\cite{MUSE} on MUSE provides a good summary of other work along these lines.
+\comment{, and is perhaps most similar to ours in assumptions, in particular ``Conjecture 1'' that a test case that used to fail on a program is more likely to pass on a mutation of a faulty statement.
+While we suggest some localizations as a welcome and interesting side effect of our distance metric and FPF analysis, our primary goal is the discovery of all faults.} Much fault localization work is largely (in many cases exclusively) based on single-fault scenarios; e.g., the MUSE \cite{MUSE} evaluation includes only 2 of 14 scenarios with multiple faults, and those scenarios include only a few faults.
 The statistical approaches may owe some of their superiority in experiments over other methods to the use of mostly single-fault evaluations \cite{Jones05}.
@@
 
+\comment{
 Some approaches to localization or patching assume the possibility of a mutant/change \emph{actually} fixing a fault.
@@
 Our primary assumption is simply that similar executions respond similarly to mutants, and in particular executions due to the same fault can be ``repaired'' by similar program changes.
-These changes are likely to be \emph{related} in some causal way to the actual fault, but are unlikely to be the actual fault, for reasons of comparative complexity of mutants and real fixes.
+These changes are likely to be \emph{related} in some causal way to the actual fault, but are unlikely to be the actual fault, for reasons of comparative complexity of mutants and real fixes.}
 

### 2017-05-05 experiments.tex
could ship in theory

@@
 In practice, their work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique, since trying multiple metrics is impractical, since a user has to examine each curve.
-\comment{It also included methods such as comparing output signatures that are inapplicable to differential testing in most cases, the most critical compiler application.}  This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work; our improvements over any single method applied to both problems are considerably larger.
+\comment{It also included methods such as comparing output signatures that are inapplicable to differential testing in most cases, the most critical compiler application.}  This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work; our improvements over any single ``reasonable'' method applied to both problems would be considerably larger (for instance, using the most obvious basis, line coverage, we see more than 17\% improvement).
 Moreover, the improvement is, for our subject programs, reliable:  choosing random subsets of the full data set, the difference in mean APFD from best-previous method is around 3\%, and statistically significant.

### 2017-05-05 conclusions.tex
first ship

@@
 \cite{multilingual}  --- 6 such localizations, compared to only 2.
-For GCC wrong-code faults, only our methods produced any very-high-quality fault localizations.
+For GCC wrong-code faults, only our methods produced any
+very-high-quality fault localizations.  When applied to non-compiler
+faults, the metric tended to either not produce a localization,
+produce a very small incorrect localization (1-3 statements), or
+produce a very high quality localization.  Restricting to the cases
+where we consider {\bf Repair} applicable (7 faults) {\bf Repair}
+perfectly localized 4 of them, and produced a 1 or 2 statement
+incorrect localization for the others.
 

### 2017-05-05 experiments.tex
first ship

@@
 This results in {\bf Repair} either providing a ``useful'' localization or almost no localization information at all to mislead a user, for the non-compiler faults, which user studies and surveys suggest is the ideal behavior due to the cost of false positives in localization \cite{AutoHelp,Kochhar}.
+This is basically a trade-off.
+MUSE, MUSEUM, and Metallaxis-FL provide localization for many more faults, in a more diverse range of settings (with single failing tests, no repairs, etc.), and using more diverse information --- but, in cases where {\bf Repair} provides a localization, it seems to often be higher quality due to its restriction to high-signal source of information (mutants repairing failing tests, only).
+In settings, unlike compilers, where there are relatively few repairs, {\bf Repair} also produces very little incorrect information, and so has a low cost to the user in terms of wasted attention.
+Note that our heuristic to stop reading after 5 wrong localizations is not even needed for all but one CoreUTILS fault --- {\bf Repair} suggests 1-3 statements at most.
 
@@
 
-\subsubsection{Mutation Analysis Costs}
+\subsection{Mutation Analysis Costs}
 

### 2017-05-05 relatedwork.tex
first ship

@@
 \cite{Xin2008} examined methods for indexing program executions, useful in all methods involving execution alignments.
-To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra,BallConcept} (using counts over structural entities), while our metric is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.
+To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra,BallConcept} (using counts over structural entities), while our metric is essentially causal, based on mutants as counterfactual \cite{LewisCause,LewisCount} versions of a program.

### 2017-05-07 conclusions.tex
ready to ship, I think

@@
 history \cite{KlausRajeev,scriptstospecs}.
+
+{\scriptsize {\bf Acknowledgements:} The authors would like to thank the authors of
+the MUSEUM \cite{multilingual} and Metallaxis-FL vs. MUSE 
+\cite{Papadakis} data sets, who provided both data and prompt assistance with
+its analysis.  In particular, Shin Hong and Thierry Chekam were
+extremely helpful in use of these data sets.}

### 2017-05-07 experiments.tex
ready to ship, I think

@@
 
-All experiments in this paper are based on subjects used in the only previous study of compiler fuzzer taming \cite{PLDI13}.
+Our primary experiments are based on subjects used in the only previous study of compiler fuzzer taming \cite{PLDI13}.
 Of the three data sets examined in that paper, this paper considers two:  faults and tests cases for SpiderMonkey 1.6, Mozilla's JavaScript engine, with tests generated by {\tt jsfunfuzz} \cite{jsfunfuzz}, and \emph{wrong code} faults and test cases for GCC 4.3.0, with tests generated by Csmith \cite{csmith}.
@@
 Figure \ref{jscurves} shows the discovery curves for the mutant repair metric compared to the ideal discovery curve and the best curve from previous work using FPF, which used a normalized Levenshtein distance \cite{lev} over the failure output and test case text \cite{PLDI13}.
+A discovery curve is a plot of the number of distinct faults that a user, examining the tests in the ranked order, would have seen after N tests (here, N goes up to 50).
+
 The APFD (Average Percent Faults Detected) values shown in the graphs are based on the measure  introduced by Rothermel et al.
@@
 APFD, as the name suggests, shows the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's area to an ideal curve, with interpolation.
+A simpler (but less informative) way to compare the curves is to note that the mutation-based metric's curve is above the best previous curve at 60\% of data points.
 %\footnote{The details of APFD calculation are slightly involved, and the interpolation and perfect curve are not exact fits for fuzzer taming; nonetheless the basic method summarizes curves fairly well and is standard in the testing literature for similar problems \cite{issta14}.}.
@@
 APFD results are useful summaries, but the curve itself is also worth examining, since (1) long sequences of tests with no new faults may discourage users more than an overall less effective but steadily climbing curve with few ``plateaus'' as we call these uninformative sequences of tests and (2) a good early curve is important to developers.
-The SpiderMonkey curve climbs very rapidly in the early portion, with perfect discovery for the first 12 faults.
+The mutant repair curve climbs very rapidly in the early portion, with perfect discovery for the first 12 faults.
 The largest plateau is 5 tests without a new fault, ignoring the long period after the 31st test.
@@
 the more extensive Proteum \cite{Proteum} operators used in MUSE/MUSEUM's evaluations.
-\comment{Even with these restrictions, running all mutants on passing tests until at least a single kill was observed took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.}  Below we provide some comparisons with MUSEUM and MUSE fault localization in the context of a full mutation result matrix using their preferred set of operators.
+\comment{Even with these restrictions, running all mutants on passing tests until at least a single kill was observed took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.}  In Section \ref{sec:otherfaults} we provide some comparisons with MUSEUM and MUSE fault localization in the context of a full mutation result matrix using their preferred set of operators.
 
@@
 GCC's APFD improvement is larger than that for SpiderMonkey, but its early curve is worse compared to the PLDI 2013 curve.
+Over all, again the curve is better than the best PLDI 13 curve at 60\% of all points.
 Using the hypothetical model where a user stops examining tests after a plateau of size 10, the user of the mutant-based taming will see 18 distinct faults by examining 25 tests, and the user of the function coverage based taming will see 20 faults after examining 45 tests.
@@
 \begin{table}
+\caption{Fault localization for compiler faults}
 \centering
@@
 }
-\caption{Fault localization for compiler faults}
 \label{bothtable}

### 2017-05-07 othertools.tex
first ship
ready to ship, I think

@@
 \subsection{Repair Localization for Non-Compiler Faults}
+\label{sec:otherfaults}
 
@@
 
-\subsubsection{MUSEUM Bug Results}  For the six faults used in the original MUSEUM paper (all with a single failing test), {\bf Repair} gives the same (perfect) localization for three faults.
+\subsubsection{MUSEUM Results}  For the six faults used in the original MUSEUM paper (all with a single failing test), {\bf Repair} gives the same (perfect) localization for three faults.
 One of the other three faults (Bug2) has no repairing mutants (that do not break some passing test) so {\bf Repair} provides no localization information at all.
@@
 In practice, we only suggest using {\bf Repair} at all in cases where 1) there is at least one repairing mutant (not killed by any passing test) for a failing test and 2) there are multiple failing tests (whether from one or multiple faults).
-When the Lines column shows a bold 0, it indicates there were no repairs.
+When the Lines column shows a bold 0, it indicates there were no repairs (7 of the 30 faults).
 The 7 faults meeting both requirements are bolded.
@@
 \begin{table}
+\caption{Fault localization for CoreBENCH faults}
 \centering
@@
 }
-\caption{Fault localization for CoreBENCH faults}
 %\vspace{-0.2in}

### 2017-05-08 experiments.tex
ok to ship

@@
 GCC 4.3.0 crash faults were essentially perfectly localized by previous approaches.
-Faults that produce a crash are also likely to be easier to debug than purely semantic problems, such as many of the SpiderMonkey faults and all GCC wrong code faults.
+In fact, on examination, only two of the 11 crash bugs are not distinguished by simply examining the crash message.
+Faults that produce a crash are also likely to be much easier to debug than semantic problems such as all of the GCC wrong code bugs and most of the JavaScript faults --- e.g., following a dynamic slice might well suffice.
 
+\comment{
 We applied our metric to two of the datasets used in the previous study of compiler fuzzer taming \cite{PLDI13}: faults and tests cases for SpiderMonkey 1.6, Mozilla's JavaScript engine, with tests generated by {\tt jsfunfuzz} \cite{jsfunfuzz}, and \emph{wrong code} faults and test cases for GCC 4.3.0, with tests generated by Csmith \cite{csmith}.
 We did not use the crash faults from that data set, as these are in practice fairly easy to identify by hand (and perfectly identified by previous metrics), and also typically easy to debug.
+}
 
@@
 
-The APFD (Average Percent Faults Detected) values shown in the graphs are based on the measure  introduced by Rothermel et al.
-to evaluate test case prioritization methods \cite{APFD}.
-This is a somewhat better summary of results than the raw curve areas used in previous work \cite{PLDI13}.
-APFD, as the name suggests, shows the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's area to an ideal curve, with interpolation.
+The APFD (Average Percent Faults Detected) values in the graphs are based on the measure  introduced by Rothermel et al.
+for evaluating test case prioritization methods \cite{APFD}.
+APFD is a somewhat better summary of results than the raw curve areas used for evaluation in previous work \cite{PLDI13}.
+APFD, as the name suggests, measures the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's area to an ideal curve, with interpolation.
 A simpler (but less informative) way to compare the curves is to note that the mutation-based metric's curve is above the best previous curve at 60\% of data points.
@@
 the best PLDI 2013 curve, Euclidean distance over function coverage vectors \cite{PLDI13}.
-GCC's APFD improvement is larger than that for SpiderMonkey, but its early curve is worse compared to the PLDI 2013 curve.
+GCC's APFD improvement is larger than that for SpiderMonkey, but its early curve (first 15 tests) is worse compared to the PLDI 2013 curve, but still has a maximum gap of only 2 redundant tests vs.
+4 for PLDI 13.
 Over all, again the curve is better than the best PLDI 13 curve at 60\% of all points.
@@
 The best methods for different subjects varied widely, even in such difficult-to-understand ways as less-fine-grained coverage providing better results in some cases, but worse in other cases.
-In practice, their work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique, since trying multiple metrics is impractical, since a user has to examine each curve.
+In practice, their work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique, since trying multiple metrics is impractical: a user has to examine each curve.
 \comment{It also included methods such as comparing output signatures that are inapplicable to differential testing in most cases, the most critical compiler application.}  This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work; our improvements over any single ``reasonable'' method applied to both problems would be considerably larger (for instance, using the most obvious basis, line coverage, we see more than 17\% improvement).
@@
 
+For the compiler faults, if any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
+MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.
+Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this is unlikely due to the nature of the repairs.
+For the CoREBench faults, {\bf Repair} produced more than twice as many perfect localizations as the other methods, though it also did not provide any ranking for a fault in a large number of cases.
+
 There are two ways to consider the performance of {\bf Repair}.
@@
 This is basically a trade-off.
-MUSE, MUSEUM, and Metallaxis-FL provide localization for many more faults, in a more diverse range of settings (with single failing tests, no repairs, etc.), and using more diverse information --- but, in cases where {\bf Repair} provides a localization, it seems to often be higher quality due to its restriction to high-signal source of information (mutants repairing failing tests, only).
+MUSE, MUSEUM, and Metallaxis-FL provide localization for many more faults, in a more diverse range of settings (with single failing tests, no repairs, etc.), and using more diverse information --- but, in cases where {\bf Repair} provides a localization, it seems to often be higher quality due to its restriction to a high-signal source of information (mutants repairing failing tests, only).
 In settings, unlike compilers, where there are relatively few repairs, {\bf Repair} also produces very little incorrect information, and so has a low cost to the user in terms of wasted attention.
-Note that our heuristic to stop reading after 5 wrong localizations is not even needed for all but one CoreUTILS fault --- {\bf Repair} suggests 1-3 statements at most.
+Note that our heuristic to stop reading after 5 wrong localizations is not even needed for all but one CoREBench fault --- {\bf Repair} suggests 1-3 statements at most.
 
@@
 
-{ For the compiler faults, if any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
+\comment{ For the compiler faults, if any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
 MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.
 Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this is unlikely due to the nature of the repairs.
+}
 
@@
 
-The largest threat to validity is that this paper's conclusions rely on only two data sets for compilers.
-The fault ground truths are possibly imperfect, due to the complex change history, and it was not possible to produce good patches for all faults, limiting our analysis of compiler fault localization.
+The largest threat to validity is that this paper's conclusions rely on only two data sets for compilers, and a small number of non-compiler faults.
+The compiler ground truths are possibly imperfect, due to the complex change history, and it was not possible to produce good patches for all faults, limiting our analysis of compiler fault localization.
 The comparison with other fault localization methods is also limited:  our compiler experiments limit the mutation budget in ways that may weaken MUSE and MUSEUM effectiveness, and the results using fault and mutant data sets from other studies provide few faults, no multi-fault problems, and very few multi-failure problems, limiting {\bf Repair} effectiveness.

### 2017-05-08 introduction.tex
ok to ship

@@
 Automated test generators tend to produce many failures, which correspond to far fewer faults.
-The distribution of faults in a set of failures tends to follow a strong power-law curve, where a few faults account for most failures, and many or even most faults have only 1 or 2 associated failures (even in a set of 1,000 or more test cases) \cite{PLDI13}.
+The distribution of faults in a set of failures tends to follow a strong power-law curve, where a few faults account for most failures, and many or even most faults have only 1 or 2 associated failures (even in a set of 1,000 or more test cases) \cite{PLDI13,OneTest}.
 Looking at all the failures to find  distinct faults is impractical.
@@
 There may be some (possibly known) faults that are very difficult to fix.
-As of 1/16/2017, clang had 109 unresolved, assigned bugs, some dating to 2008 \cite{clangbugs}.
-Knowing if a randomly generated test case is an instance of a known failure can be difficult, and fixing long-standing problems to improve bug triage is unlikely.
-A research group with a novel fault-detection technique \cite{csmith,ISSTA12,ZhendongPLDI} may dump dozens of new faults on a project at once, and high priority faults may be hidden by stubborn low-priority faults.
+As of 5/8/2017, clang had 107 unresolved, assigned bugs, some dating to 2008 \cite{clangbugs}.
+Knowing if a randomly generated test case is an instance of a known failure can be difficult, and fixing long-standing problems is unlikely.
+Applying a novel fuzzing technique \cite{csmith,ISSTA12,ZhendongPLDI} may dump dozens of new faults on a project at once, and high priority faults may be hidden by stubborn low-priority faults.
 
@@
 
-The key insight of this paper is based on an idea with numerous possible applications: \emph{if two test cases that fail can be made to succeed by the same small modification to the original program $P$, this provides evidence that they fail due to the same fault}.
+The methods we propose are based on an idea with numerous possible applications: \emph{if two test cases that fail can be made to succeed by the same small modification to the original program $P$, this provides evidence that they fail due to the same fault}.
 The idea is hardly controversial:  most developers determine if they have fixed all faults in a set of failing test cases by re-running the tests after they fix at least one fault.
@@
 However, what should matter is that failures, which correspond to faults, can be made to succeed.
-When a test case $t$ fails for program $P$ but succeeds for mutant $m$ of $P$ we say that $m$ \emph{repairs} $t$.
+When a test case $t$ fails for program $P$ but succeeds for mutant $m$ of $P$ we say that $m$ \emph{repairs} $t$ (in practice, we will usually restrict this to unkilled mutants $m$).
 In most cases this does not mean that $m$ is the inverse of a fault, but that $m$ somehow avoids the consequence of a fault.
@@
 
-This leads to the idea of comparing executions by their response to mutants.
-In the presence of multiple correctness properties, in principle any two executions can be compared, not just two failures, since a mutant can cause a failure to violate additional or different properties.
-However, for practical purposes the most interesting implication is that if two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.
+This leads to the general idea of comparing any two executions by their response to mutants, since a mutant can cause a passing or failing test to violate more/different properties.
+However, for practical purposes the most interesting implication is that if two failing test cases can be made to differ as to whether they fail by a mutant, this provides evidence that they fail due to different faults.
 
@@
 We show that such a metric can improve, in a statistically significant way, the ability to distinguish tests based on which faults they trigger.
-Furthermore, fault localization based on the same metrics can successfully pinpoint the location of some compiler faults sufficiently to be likely to substantially aid debugging, and, when applicable, performs very well for some faults studied in recent papers on mutant-based fault localization.
+Furthermore, fault localization based on the same metrics can successfully pinpoint the location of some compiler faults sufficiently to be likely to substantially aid debugging, and, when applicable, performs very well for some faults studied in recent papers on mutant-based fault localization \cite{multilingual,Papadakis}.
 
@@
 Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over best previous results \cite{PLDI13}, in multiple ways.
-Our best preliminary localization approach also compares favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual,Metallaxis}, albeit with a more limited applicability (but a lower cost).
+Our fault localization approach also compares favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual,Metallaxis}, albeit with a more limited applicability (but also a lower cost).

### 2017-05-08 localizations.tex
ok to ship

@@
 too often, code that executes as a consequence of the fault, code
-accidentally executed with the fault, and so forth.  Second, most
+accidentally executed with the fault, and so forth.  We note that most
 evaluations of fault localization have been performed over programs
@@
 expected, we propose an alternative method: let $f$ be an arbitrary
-test with as few repairing mutants as possible.
+test with as few repairing mutants as possible, to allow the distance
+metric to focus attention on a few key mutants.
 

### 2017-05-08 othertools.tex
ok to ship

@@
 MUSE and MUSEUM over compiler faults.
-We obtained the data from two evaluations of mutant-based fault localization methods:  the set of faults used to evaluate MUSEUM \cite{multilingual}, and a set of CoreBENCH \cite{CoreBENCH} faults used in an evaluation of MUSE and Metallaxis-FL by Chekal et.
+We obtained the data from two evaluations of mutant-based fault localization methods:  the set of faults used to evaluate MUSEUM \cite{multilingual}, and a set of CoREBench \cite{CoREBench} faults used in an evaluation of MUSE and Metallaxis-FL by Chekal et.
 al \cite{Papadakis} (we were unable to obtain the original MUSE dataset, thus far).
-These data sets are not ideal for evaluating {\bf Repair}: for all but 10 of the faults (all from the CoreBENCH set) there is only a single failing test case in the data.
+These data sets are not ideal for evaluating {\bf Repair}: for all but 10 of the faults (all from the CoREBench set) there is only a single failing test case in the data.
 {\bf Repair} for single failing tests is essentially MUSEUM without use of passing tests, except to reject any mutation that breaks a passing test.
@@
 
-\subsubsection{CoreBENCH Results} For the full data set in the paper \cite{Papadakis}, {\bf Repair} does not perform as well as MUSE and Metallaxis-FL, as expected.
+\subsubsection{CoREBench Results} For the full data set in the paper \cite{Papadakis}, {\bf Repair} does not perform as well as MUSE and Metallaxis-FL, as expected.
 It is only able to  rank a faulty line of code for 12 of 30 faults.
-However, for these faults, it performs very well indeed, perfectly localizing 7 of the 12.
+However, for these faults, it performs very well indeed, perfectly localizing 7 of the 12 (vs.
+3 total perfect localizations for Metallaxis-FL and 4 for MUSE).
 {\bf Repair} is uniquely best for 5 of the 12, and is tied for best for another 6; for the remaining fault, it performs better than Metallaxis-FL but worse than MUSE.
-Table \ref{otherbugs} shows the results.
+Table \ref{otherbugs} shows full results.
 The fourth column shows, for those bugs where {\bf Repair} did not rank any faulty statement, how many statements {\bf Repair} proposed for a user to examine.
@@
 \begin{table}
-\caption{Fault localization for CoreBENCH faults}
+\caption{Fault localization for CoREBench faults}
 \centering

### 2017-05-08 relatedwork.tex
ok to ship

@@
 The distance metrics used with FPF in the previous work, however, were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.
-This paper uses the FPF approach in the context of a more principled causal comparison of program executions that applies in theory to any executions, not just failing test cases, and introduces novel fault localizations.
+This paper uses the FPF approach in the context of a more causal comparison of program executions that applies in theory to any executions, not just failing test cases, and introduces a novel fault localization.
 Prior to the FPF-based work, Francis et al.
@@
 discuss general issues in debugging multiple faults efficiently \cite{Jones07}.
+The work of Groce et al.
+\cite{OneTest} provides an alternative approach to fault identification, based on term rewriting to normalize failures.
+However, this approach is not currently applicable to compiler fuzzing, as it only support tests that are sequences of method calls.
 
@@
 By focusing on a single failing test case (for us, selected by fuzzer taming) as the locus of analysis, our work, like some early efforts \cite{NearNeighbor,Cleve05,GroceError} may be more suited for situations involving many faults.
-MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single failure.
+\comment{MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single failure.
+}
 

### 2018-03-02 introduction.tex
get ready for issre

@@
 Due to complexity of intermediate representation, interaction of optimizations, subtle language semantics, and other issues, debugging compiler faults may often be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
-We propose the use of program mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.
+This paper proposes the use of mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.
 The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault.
@@
 Some recent work has also considered these changes as a source of fixes or localizations for program faults \cite{achour,multilingual,MUSE,DebroyMutant}.
-Our experience, as well as empirical data \cite{GopinathMutants}, suggests that very few compiler faults are due to the kind of simple syntactic changes found in mutants.
+Published empirical data \cite{GopinathMutants}, as well as our data, suggests that very few compiler faults are due to the kind of simple syntactic changes found in mutants.
 However, what should matter is that failures, which correspond to faults, can be made to succeed.
@@
 \centering
-\includegraphics[width=0.6\columnwidth]{onionring}
-\caption{\scriptsize{The onion-ring model of repairs, showing all repairs for the center-most test case, and all other test cases repaired by those mutants.
-Rings represent repairing mutants, and diamonds repaired test cases.
+\includegraphics[width=0.4\columnwidth]{onionring}
+\caption{\scriptsize{Onion-ring shows all repairs for the center-most test case, and all other test cases repaired by same mutants.
+Rings represent repairing mutants, and diamonds test cases.
 Diamond pattern symbolizes fault.
-For the innermost ring, all failures are due to the same fault.
-As rings expand to include more failures, the set of faults grows (as does distance).}}
+For innermost ring, all failures are due to the same fault.
+As rings expand, the set of faults grows (as does distance).}}
 \vspace{-0.1in}

### 2018-07-16 experiments.tex
get ready for issre
PLDI 13 to FPF benchmark change

@@
         \centering
-        \includegraphics[width=1.02\textwidth]{jscurve}
+        \includegraphics[width=1.0\textwidth]{jscurve}
         \caption{Discovery curves for SpiderMonkey 1.6 faults}
@@
         \centering
-        \includegraphics[width=1.02\textwidth]{gcccurve}
+        \includegraphics[width=1.0\textwidth]{gcccurve}
         \caption{Discovery curves for GCC 4.3.0 wrong-code faults}
@@
     \end{subfigure}
-    \caption{Discovery curves compared to best curves from PLDI 2013}
+    \caption{Discovery curves compared to best curves from Chen et. al PLDI 2013 data (FPF benchmark)}
 \end{figure*}
@@
 
-Our primary experiments are based on subjects used in the only previous study of compiler fuzzer taming \cite{PLDI13}.
+Our primary experiments are based on subjects used in the only previous study of compiler fuzzer taming (from Chen et al.'s PLDI 2013 paper \cite{PLDI13}), which we hereafter refer to as the FPF benchmark.
 Of the three data sets examined in that paper, this paper considers two:  faults and tests cases for SpiderMonkey 1.6, Mozilla's JavaScript engine, with tests generated by {\tt jsfunfuzz} \cite{jsfunfuzz}, and \emph{wrong code} faults and test cases for GCC 4.3.0, with tests generated by Csmith \cite{csmith}.
@@
 Experiments were performed on a MacBook Pro with 16GB RAM and dual-core 3.1GHz Intel Core i7; GCC executed on a VirtualBox-hosted Ubuntu 11.04.
-For GCC, some test cases from the 2013 PLDI paper no longer failed, presumably due to unknown differences in execution environment, OS, or memory layout.
+For GCC, some test cases from the FPF set no longer failed, presumably due to unknown differences in execution environment, OS, or memory layout.
 Discarding these reduced the set of test cases from 1,275 to 1,117 and the number of distinct faults to 27 rather than 35.
-For SpiderMonkey, all 1,749 test cases from the PLDI 2013 data failed in the new environment, representing 28 distinct faults.
+For SpiderMonkey, all 1,749 test cases from the FPF benchmark failed in the new environment, representing 28 distinct faults.
 
@@
 Assuming that few users will examine more than 50 test cases without fixing some faults, the user may give up after seeing 10 or more tests without a new fault based on the mutant repair curve, and in fact lose nothing by doing so.
-The PLDI 2013 curve, in contrast, has a plateau of size 17 after the 17th fault.
-If we assume a user stops examining tests after a size 10 or greater plateau, the user will only see 17 faults using the PLDI 2013 metric, vs.
+The best FPF benchmark curve, in contrast, has a plateau of size 17 after the 17th fault.
+If we assume a user stops examining tests after a size 10 or greater plateau, the user will only see 17 faults using the best benchmark metric, vs.
 19 with our metric.
@@
 
-In order to check our results statistically, we sliced the PLDI 13 tests into 20 randomly selected equal-sized (as much as possible) distinct subsets.
-The repair metric had a mean APFD (89.23\%) for these subsets that was 2.85\% better than the mean APFD for the PLDI 13 metric (86.76\%), over sets of $\sim 10$ faults.
+In order to check our results statistically, we sliced the FPF benchmark tests into 20 randomly selected equal-sized (as much as possible) distinct subsets.
+The repair metric had a mean APFD (89.23\%) for these subsets that was 2.85\% better than the mean APFD for the best benchmark metric (86.76\%), over sets of $\sim 10$ faults.
 The result was statistically significant ($p < 0.05$ by Wilcoxon test).
 Median repair APFD was 90.05\%, vs.
-84.75\% for PLDI 13's best metric.
+84.75\% for the best benchmark metric.
 
@@
 This is due to a problem with the data set:  producing ground truth patches for a large application, in the sense of finding a minimal, clearly fault-fixing code change that can be back-ported to the original code is difficult \cite{PLDI13}.
-While we believe the 28 faults identified in the PLDI 2013 data set are correct, it is very difficult to produce a valid patch of version 1.6 that captures the fix for most of these faults.
+While we believe the 28 faults identified in the FPF benchmark data are correct, it is very difficult to produce a valid patch of version 1.6 that captures the fix for most of these faults.
 In some cases the final commit that caused tests to stop failing appeared to only be the end of a complex series of changes that converged on a correct fix.
@@
   \centering
-  \includegraphics[width=\columnwidth]{comparegcc}
+  \includegraphics[width=0.8\columnwidth]{comparegcc}
   \caption{APFD values for GCC suite slices}
@@
 Figure \ref{gcccurves} shows discovery curves for the mutant repair metric vs.
-the best PLDI 2013 curve, Euclidean distance over function coverage vectors \cite{PLDI13}.
-GCC's APFD improvement is larger than that for SpiderMonkey, but its early curve (first 15 tests) is worse compared to the PLDI 2013 curve, but still has a maximum gap of only 2 redundant tests vs.
-4 for PLDI 13.
-Over all, again the curve is better than the best PLDI 13 curve at 60\% of all points.
+the best benchmark curve, Euclidean distance over function coverage vectors \cite{PLDI13}.
+GCC's APFD improvement is larger than that for SpiderMonkey, but its early curve (first 15 tests) is worse compared to the benchmark curve, but still has a maximum gap of only 2 redundant tests vs.
+4 for the benchmark.
+Over all, again the curve is better than the best benchmark curve at 60\% of all points.
 Using the hypothetical model where a user stops examining tests after a plateau of size 10, the user of the mutant-based taming will see 18 distinct faults by examining 25 tests, and the user of the function coverage based taming will see 20 faults after examining 45 tests.
@@
 
-Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both PLDI 13 metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement --- even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults) --- is more than 3.8\% better than either PLDI 13 metric.
+Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both benchmark metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement --- even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults) --- is more than 3.8\% better than either benchmark metric.
 
@@
 In practice, users simply do not know ground truth to rank curves, thus Chen et al.
-\cite{PLDI13} do not really give a practical approach to distance metric selection in their PLDI work.
+\cite{PLDI13} do not really give a practical approach to distance metric selection.
 The best methods for different subjects varied widely, even in such difficult-to-understand ways as less-fine-grained coverage providing better results in some cases, but worse in other cases.
@@
 Even if the FPF curve remains the same, a critical mutant for localization might be omitted.
-We performed some limited experiments, and found that for a few sample sizes ranging from 13\% of mutants to 50\% of mutants or using only some mutation operators (with the exception of operator-replacement-only), APFD was always better than PLDI 13 best.
+We performed some limited experiments, and found that for a few sample sizes ranging from 13\% of mutants to 50\% of mutants or using only some mutation operators (with the exception of operator-replacement-only), APFD was always better than benchmark set's best metric.
 However, the improvement in some cases was \emph{much} smaller than for the full set, and varied considerably (in some cases it was up to 2\% better than our full results).

### 2018-07-16 experiments.tex
more fixes from reviewers

@@
 Most mutants repaired a small number of tests, but a few mutants repaired a very large number of tests.
-A few mutants repaired \emph{all} SpiderMonkey faults; obviously these mutants were not actual fault fixes, but effectively disabled some mechanism {\tt jsfunfuzz} used to detect failure.
+A few mutants repaired \emph{all} SpiderMonkey faults; obviously these mutants were not actual fault fixes, but effectively disabled the mechanism {\tt jsfunfuzz} used to detect failure.
 The mean and median homogeneity for repairing mutants (\% of failures repaired corresponding to the most common fault repaired) were 77.96\% and 86.36\%, respectively.
@@
 A dash for a column means that localization did not rank any faulty statements, or assigned all faulty statements suspiciousness of 0.0.
-The three rankings are: our {\bf Repair} localization, the MUSE \cite{MUSE} localization, and the MUSEUM \cite{multilingual} localization, which uses the same formula as MUSE, but works better for multiple faults because it uses only one failure.
+The three rankings are: our {\bf Repair} localization, the MUSE \cite{MUSE} localization, and the MUSEUM \cite{multilingual} localization.
+MUSEUM uses the same formula as MUSE, but works better for multiple faults because it uses only one failure.
 The MUSE/MUSEUM formulas normally make use of information from passing tests as well: when mutating a statement makes a passing test case fail, it makes the statement \emph{less} suspicious, by a weighted amount.
@@
 
-For GCC wrong code faults, the baseline chance two test cases shared an underlying fault was 38.97\%.
+For GCC wrong code faults, the baseline chance that two test cases shared an underlying fault was 38.97\%.
 Knowing that two test cases shared a repairing mutant raised the chance to 68.5\%, and knowing they disagreed on a repair lowered it to only 19.01\%.
@@
 Knowing two test cases were both not fixed by a mutant again provided marginal evidence of sharing a fault: 39.2\% chance of matching faults.
-For 99.3\% (all but 8) of the 1,090 non-singleton fault failures in GCC wrong code, the closest test case(s) by the repair metric matched fault.
+For 99.3\% (all but 8) of the 1,090 non-singleton fault failures in GCC wrong code, the closest test case(s) by the repair metric showed the same fault.
 The best previous reported FPF metric, function coverage vectors, had a matching rate of 92.2\%.

### 2018-07-16 othertools.tex
more fixes from reviewers

@@
 In practice, we only suggest using {\bf Repair} at all in cases where 1) there is at least one repairing mutant (not killed by any passing test) for a failing test and 2) there are multiple failing tests (whether from one or multiple faults).
-When the Lines column shows a bold 0, it indicates there were no repairs (7 of the 30 faults).
+When the Stmts column shows a bold 0, it indicates there were no repairs (7 of the 30 faults).
 The 7 faults meeting both requirements are bolded.
@@
 & \#Failing & & & & \\
-Fault & Tests & Repair & Lines & Metallaxis-FL & MUSE\\
+Fault & Tests & Repair & Stmts & Metallaxis-FL & MUSE\\
 \hline

### 2018-07-16 localizations.tex
ready to submit maybe

@@
 \begin{quote}
-The {\bf Coverage} localization for a failing test case $f$ in the set
-of failures $F$ ranks
+The {\bf Coverage} localization for a failing test case $f \in F$ ($F$
+is the set of all failing test cases) ranks
 statements by the function $r$ (for rank) where:
@@
 \begin{quote}
-The {\bf Repair} localization for a failing test case $f$ ranks
-statements by the function $r$ where:
+For a set $F$ of failing test cases and $m$ of mutants, 
+the {\bf Repair} localization for $f \in F$, where $R(f) \in m$ is the set
+of mutants repairing $f$, ranks
+statements by the function $r$:
 %$$dmax(m,f) = max_{t : m \in R(t)} d(t,f)$$
 \[
-r(m,f) = 
+r(f) = 
 \begin{cases}

### 2018-07-17 relatedwork.tex
fix up for ISSRE camera ready

@@
 The most closely related work is that of Chen et.
-al \cite{PLDI13}.
-We use two of the same compiler data sets, and both apply the FPF \cite{Gonzalez} algorithm.
-The distance metrics used with FPF in the previous work, however, were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.
-This paper uses the FPF approach in the context of a more causal comparison of program executions that applies in theory to any executions, not just failing test cases, and introduces a novel fault localization.
-Prior to the FPF-based work, Francis et al.
+al \cite{PLDI13}, whose FPF \cite{Gonzalez} algorithm and benchmark we use.
+We extend their work with a better, and more universal distance metric, as well as an approach to fault localization in addition to fuzzer taming.Prior to the FPF-based work, Francis et al.
 \cite{Podgurski04}, Podgurski et al.
@@
 Jones et al.
-discuss general issues in debugging multiple faults efficiently \cite{Jones07}.
+discuss general issues in debugging multiple faults \cite{Jones07}.
 The work of Groce et al.
-\cite{OneTest} provides an alternative approach to fault identification, based on term rewriting to normalize failures.
-However, this approach is not currently applicable to compiler fuzzing, as it only support tests that are sequences of method calls.
+\cite{OneTest} provides an alternative approach to fault identification, based on term rewriting to normalize failures, but is not currently applicable to compiler fuzzing, as it only supports tests that are sequences of method calls in Python \cite{ISSTA15,tstlsttt}.
 

### 2019-01-11 journal/experiments.tex
journal start

--- 2018-07-16 experiments.tex
@@
 Over all, again the curve is better than the best benchmark curve at 60\% of all points.
-Using the hypothetical model where a user stops examining tests after a plateau of size 10, the user of the mutant-based taming will see 18 distinct faults by examining 25 tests, and the user of the function coverage based taming will see 20 faults after examining 45 tests.
+Using the hypothetical model where a user stops examining tests after a plateau of size 10, the user of the mutant-based taming will see 18 distinct faults by examining 25 tests, and the user of the function coverage will see 20 faults after 45 tests.
 If a user gives up after a plateau of size 5, our approach again lets a user examine 18 faults over 25 tests.
@@
 Most mutants fixed only a few tests, to a greater extent than was true with SpiderMonkey.
-Mean homogeneity (\% of repaired tests corresponding to most common repaired fault) was 79.4\%, median 100\%.
+Mean homogeneity was 79.4\% (median 100\%).
 

### 2019-01-11 journal/metrics.tex
journal start

--- 2016-12-19 metrics.tex
@@
 where a 1 bit indicates that the mutant in question repairs this test
-case.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
+case\footnote{By \emph{repairing} a failing test we here mean that a mutant
+causes the test to pass, \emph{and} the mutant does not cause
+other, previously passing tests, to fail.}.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
 by 100 mutants.  If these two test cases share 90 common repairing
@@
 This metric, the portion of mismatches over bits that are 1 in either
-vector, matches our intuitions about the above comparisons\footnote{In Section \ref{conc} we discuss possible uses of $d$
+vector, matches our intuitions about the above comparisons.  It
+essentially formalizes the intuition of the onion-ring model, where
+more matching repairs indicates a much higher probability two failures
+are due to the same fault, even if not all repairs match.
+\comment{
+\footnote{In Section \ref{conc} we discuss uses of $d$
 requiring comparison of both failing and passing test cases:  in this 
@@
 failing property succeeds and -1 that a succeeding property 
-fails.}.  It
-essentially formalizes the intuition of the onion-ring model, where
-more matching repairs indicates a much higher probability two failures
-are due to the same fault, even if not all repairs match.
+fails.  For such an application, we could remove the requirement that
+a repair not be a killed mutant, since the passing test vectors would encode
+that information.}}
 
@@
 metric we compare it, alone, to the previous best metrics.
+
+In Section \ref{conc} we discuss uses of $d$
+requiring comparison of both failing and passing test cases:  in this 
+case, we would use three-valued vectors, with a bit for each \emph{property}, where 1 indicates a 
+failing property succeeds and -1 that a succeeding property 
+fails.  For such an application, we could remove the requirement that
+a repair not be a killed mutant, since the passing test vectors would encode
+that information

### 2019-01-11 journal/oldintro.tex
journal start

--- 2017-05-05 introduction.tex
@@
 
-The proposed techniques in this paper are motivated in large part by progress in automated testing.
-Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in modern optimizing compilers.
+Recent advances in automated test generation, reported in numerous papers and implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in, e.g., modern optimizing compilers\footnote{In this paper, we take the problem of testing and debugging compilers as representative of the general problem of testing and debugging complex critical systems software.}.
 Unfortunately, the blessing of many failing test cases can become a curse for a developer.
@@
 Repairs can also provide  information for understanding and debugging discovered faults.
-\comment{The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.
-Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).}
+The approach suggests a pThese may include assertions, differential testing \cite{Differential,ICSEDiff}, timing constraints, temporal logic formulas, or any other checkable properties.
+What, then, makes two executions similar?
+Our proposal is to say that:
 
-\subsection{Fuzzer Taming and Debugging Assistance}
-
-Automated test generators tend to produce many failures, which correspond to far fewer faults.
-The distribution of faults in a set of failures tends to follow a strong power-law curve, where a few faults account for most failures, and many or even most faults have only 1 or 2 associated failures (even in a set of 1,000 or more test cases) \cite{PLDI13}.
-Looking at all the failures to find  distinct faults is impractical.
-It is also impractical to apply the iterative approach:  first, fix one test's fault; then re-run all tests, until all tests pass.
-There may be some (possibly known) faults that are very difficult to fix.
-As of 1/16/2017, clang had 109 unresolved, assigned bugs, some dating to 2008 \cite{clangbugs}.
-Knowing if a randomly generated test case is an instance of a known failure can be difficult, and fixing long-standing problems to improve bug triage is unlikely.
-A research group with a novel fault-detection technique \cite{csmith,ISSTA12,ZhendongPLDI} may dump dozens of new faults on a project at once, and high priority faults may be hidden by stubborn low-priority faults.
-
-In a PLDI 2013 paper, Chen et al.
-defined the \emph{fuzzer taming problem}: ``Given a potentially large collection of test cases, each of which triggers a bug, rank them in such a way that test cases triggering distinct bugs are early in the list.'' \cite{PLDI13}.
-Their proposed solution was to use the Furthest-Point-First \cite{Gonzalez} (FPF) algorithm.
-An FPF ranking of test cases requires a distance metric $d$ on test cases, and ranks test cases so that dissimilar tests appear earlier.
-The hypothesis of Chen et al.
-was that dissimilar tests, by a well-chosen metric, will also fail due to different faults.
-FPF is a greedy algorithm that proceeds by repeatedly adding the item with the maximum minimum distance to all previously ranked items. Given an initial seed item $r_0$, a set
-$S$ of items to rank, and a distance metric $d$, FPF computes $r_i$ as $s \in S: \forall s' \in S: min_{ j < i}(d(s,r_j)) \geq min_{j < i}(d(s',r_j))$.
-
-Fuzzer taming results using FPF were promising, but limited.
-Rather than proposing a universal metric for use in fuzzer taming, Chen et al.
-offered many possible metrics (based on ideas in the literature, such as edit distances \cite{lev} or coverage spectra \cite{RepsSpectra} methods), none of which performed notably well across all three of their benchmark data sets, and some of which were not even applicable to all data sets.
-For the two most difficult benchmarks, none of the metrics performed extremely well.
-FPF-based approaches remain in need of a universal but effective metric, particularly for the critical first 50 test cases examined \cite{PLDI13}.
-
-One reason fuzzer taming is needed is that debugging compiler faults is, as noted, difficult.
-If developers were able to fix faults quickly once they were identified, it would not be as important to have very good fault identification.
-Many duplicate failures could be removed by fixing the underlying faults.
-Ideally, a technique able to provide effective fuzzer taming should be able to translate its underlying rationale into some form of automated debugging assistance.
-A variety of automated debugging methods \cite{NearNeighbor,GroceError} have been based on distance, so there is reason to expect that a high-quality metric might also contribute to fault localization.
-
-\subsection{Comparison by Response to Mutation}
-
-The key insight of this paper is based on an idea with numerous possible applications: \emph{if two test cases that fail can be made to succeed by the same small modification to the original program $P$, this provides evidence that they fail due to the same fault}.
-The idea is hardly controversial:  most developers determine if they have fixed all faults in a set of failing test cases by re-running the tests after they fix at least one fault.
-As a method for fuzzer taming or debugging assistance, however, this is useless.
-Fortunately, developer-provided patches are not the only source of changes to a program.
-Mutation analysis \cite{demillo1978hints,budd1980theoretical} applies large numbers of small syntactic changes to a program, usually to evaluate a test suite \cite{mutant,justmutants}.
-Some recent work has also considered these changes as a source of fixes or localizations for program faults \cite{achour,multilingual,MUSE,DebroyMutant}.
-Our experience, as well as empirical data \cite{GopinathMutants}, suggests that very few compiler faults are due to the kind of simple syntactic changes found in mutants.
-However, what should matter is that failures, which correspond to faults, can be made to succeed.
-When a test case $t$ fails for program $P$ but succeeds for mutant $m$ of $P$ we say that $m$ \emph{repairs} $t$.
-In most cases this does not mean that $m$ is the inverse of a fault, but that $m$ somehow avoids the consequence of a fault.
-It is not important that repairs actually fix anything, if all we seek is a way to decide whether failures are due to the same fault.
-
-This idea is in fact a consequence of a more general (and difficult to investigate) hypothesis.
-What is interesting about a program execution?
-For many purposes, the only interesting properties of an execution are those that in some way relate to the correctness properties of the program, assuming we have a good set of correctness properties.
-These may include assertions, differential testing \cite{Differential,ICSEDiff}, timing constraints, temporal logic formulas, or any other checkable properties.
-What, then, makes two executions similar?
-Our proposal is that \emph{two executions of $P$ are similar to the degree that changes to $P$ produce the same changes in correctness of the executions.}
+\begin{quote}
+{\bf Hypothesis 2:} Two executions of $P$ are similar to the degree that changes to $P$ produce the same changes in correctness of the executions.
+\end{quote}
 
 This leads to the idea of comparing executions by their response to mutants.
-In the presence of multiple correctness properties, in principle any two executions can be compared, not just two failures, since a mutant can cause a failure to violate additional or different properties.
-However, for practical purposes the most interesting implication is that if two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.
-
-\comment{
+In the presence of multiple correctness properties, any two executions can be compared, not just two failures, since a mutant can cause a failure to violate additional or different properties.
 %Where there are multiple, independent, correctness properties, this even allows us to compare failing and successful executions, since a mutant can cause a failure to violate a different property.
@@
 This is {\bf Hypothesis 1a:} If two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.
-}
 
-%\subsection{The Onion-Ring and Causality}
+\subsection{The Onion-Ring and Causality}
 
-This not a claim that being repaired by the same mutant perfectly predicts fault equivalence.
+These hypothesis do not claim that being repaired by the same mutant perfectly predicts fault equivalence.
 To understand why that seems highly unlikely, consider why a mutant of a compiler might cause a test case to stop failing.
@@
 Optimizations are generally guarded by conditional switches --- in some cases, so that the optimization can be turned off by a user, in other cases when a property of the code in question prevents the optimization.
-If two failures become successes due to turning off an optimization, it is presumably \emph{some} level of evidence they are the same fault, compared to the baseline probability.
+If two failures become successes due to turning off an optimization, it is certainly some level of evidence they are the same fault, compared to the baseline probability.
 Intuitively, the evidence provided seems much stronger than the fact that, e.g., the two test cases execute a shared line of code \cite{RepsSpectra}.
@@
 If a mutant turns off a large number of optimizations, it is weak evidence.
-Must we then characterize mutant granularity?
-No.
+Characterizing the granularity of program mutants is likely difficult, but fortunately it is not necessary.
 
@@
 \caption{\scriptsize{The onion-ring model of repairs, showing all repairs for the center-most test case, and all other test cases repaired by those mutants.
-Rings represent repairing mutants, and diamonds repaired test cases.
+Circles represent repairing mutants, and diamonds repaired test cases.
 Diamond pattern symbolizes fault.
-For the innermost ring, all failures are due to the same fault.
-As rings expand to include more failures, the set of faults grows (as does distance).}}
-\vspace{-0.1in}
+For the innermost ring of repairing mutant(s), all test cases are due to the same fault.
+As rings expand to include more failures, the set of faults repaired grows (as does distance).}}
 \label{fig:onion}
@@
 Turning off a data-structure invariant checker is another possible false repair that might cover many different faults.
-In all these cases, assuming more interesting and specific ways to repair faults exist, a distance metric can still work.
+In all these cases, so long as there are other, more interesting and more specific ways to repair some faults, a distance metric can still distinguish faults effectively.
 
@@
 Debugging by changing test cases is not useful; debugging by altering the program is precisely how we seek to use the causes we discover.
-Using mutants as causes in order to measure similarity ``inverts'' the ideas of Lewis \cite{LewisCause,LewisCount}, where similarity determines causality.
-One pleasant side-effect is that execution representation \cite{NearNeighbor,RepsSpectra} is not a difficulty for this kind of metric, since an execution is always represented as a set of correctness evaluations.
+Using mutants as causes in order to measure similarity essentially inverts the ideas of David Lewis \cite{LewisCause,LewisCount}, where similarity is used to determine causality.
 
-This paper investigates the utility of using distance metrics based on the causal approach suggested above.
-We show that such a metric can improve, in a statistically significant way, the ability to distinguish tests based on which faults they trigger.
-Furthermore, fault localization based on the same metrics can successfully pinpoint the location of some compiler faults sufficiently to be likely to substantially aid debugging, and, when applicable, performs very well for some faults studied in recent papers on mutant-based fault localization.
+\subsection{Contributions}
 
-%\subsection{Contributions}
-
-The contributions of this paper include our ideas about how to measure distance between program executions, but more importantly some practical proposals for using such a metric in compiler fuzzer taming and fault localization.
-Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over best previous results \cite{PLDI13}, in multiple ways.
-Our best preliminary localization approach also compares favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual,Metallaxis}, albeit with a more limited applicability (but a lower cost).
+The contributions of this paper are: hypotheses about how to measure distance between program executions, and practical proposals for using such a metric in fuzzer taming and in novel fault localization and explanation methods.
+Evaluating the effectiveness of these methods also provides experimental evidence as to the validity of our hypotheses.
+Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over previous results \cite{PLDI13}, in multiple ways.
+Our preliminary localization approaches compare favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual}, for the mutation analysis budget in our experiments.

### 2019-01-11 journal/othertools.tex
journal start

--- 2018-07-16 othertools.tex
@@
 It is not proposed (or at least, not here evaluated) as a general-purpose fault localization method.
-However, it is possible to compare {\bf Repair} with other methods, using their full mutation and experimental set, to provide a more apples-to-apples comparison than the evaluation vs.
+However, it is possible to compare {\bf Repair} with other methods, using their full mutation and experimental set, to provide a more generalizable (and fair) comparison than the evaluation vs.
 MUSE and MUSEUM over compiler faults.
@@
 The distance metric lets {\bf Repair} make use of these other faults, by enabling it to focus on repairing mutants most relevant to the failing test.
+{\bf Repair} can, in many cases, have a lower cost than MUSEUM or Metallaxis-FL because it does not need to run mutants that do not repair any failing tests on the passing tests: analysis can begin with failing tests and only analyze passing tests covering repairing mutants.
 

### 2019-01-14 journal/experiments.tex
add new graph

@@
 
-In order to check our results statistically, we sliced the FPF benchmark tests into 20 randomly selected equal-sized (as much as possible) distinct subsets.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\columnwidth]{comparejs}
+  \caption{APFD values for Spidermonkey suite slices}
+  \label{comparejs}
+\end{figure}%
+
+In order to check our results statistically, we sliced the FPF benchmark tests into 20 randomly selected equal-sized (as much as possible) distinct subsets (Figure \ref{comparejs}) .
 The repair metric had a mean APFD (89.23\%) for these subsets that was 2.85\% better than the mean APFD for the best benchmark metric (86.76\%), over sets of $\sim 10$ faults.

### 2019-01-20 journal/example.tex
mutants16 paper

\section{A Simple Example}

To better understand the causal approach taken in this paper, it is useful to consider a simplified, abstract example, and understand how the approach proposed would apply to this example.

### 2019-01-20 journal/experiments.tex
more text

@@
 We applied X-means \cite{xmeans} in an attempt to use clustering with our metrics, but, as with previous results \cite{PLDI13} it did not compare well with FPF, and the runtime was much higher.
-We confirm the conclusion \cite{PLDI13} that clustering is of limited value in fuzzer taming.
-% (at least with X-means) does not work well in a setting with extreme disparity in instance counts for faults.
-%A further difficulty for our setting is that X-means and most off-the-shelf clustering tools take vectors, not a distance metric.
+We confirm the conclusion \cite{PLDI13} that clustering (at least with X-means) does not work well in a setting with extreme disparity in instance counts for faults.
+A further difficulty for our setting is that X-means and most off-the-shelf clustering tools take vectors, not a distance metric.
 X-means is using the raw repair data, not weighted by the fact that 0-0 agreement is much less informative than other possible matchings (a Euclidean over 0-1 vectors is just the square root of Hamming distance).
@@
 APFD, as the name suggests, measures the percent of all faults discovered at the ``average'' point on the curve, by comparing the curve's area to an ideal curve, with interpolation.
-A simpler (but less informative) way to compare the curves is to note that the mutation-based metric's curve is above the best previous curve at 60\% of data points.
-%\footnote{The details of APFD calculation are slightly involved, and the interpolation and perfect curve are not exact fits for fuzzer taming; nonetheless the basic method summarizes curves fairly well and is standard in the testing literature for similar problems \cite{issta14}.}.
+A simpler (but less informative) way to compare the curves is to note that the mutation-based metric's curve is above the best previous curve at 60\% of data points\footnote{The details of APFD calculation are slightly involved, and the interpolation and perfect curve are not exact fits for fuzzer taming; nonetheless the basic method summarizes curves fairly well and is standard in the testing literature for similar problems \cite{issta14}.}.
 
@@
 
-\comment{
 Results are reported as absolute ranks, rather than more sophisticated \cite{MUSE} localization evaluations because our localizations do not assign suspiciousness scores, only rankings, and in accord with the proposal of Parnin and Orso that only localizations ranking a fault in the top few statements may be useful to users \cite{AutoHelp}.
-We expect users to examine coverage changes and mutants.
+We expect users to examine the mutants and consider how they help avoid the fault.
 Blind, unaided pointers to portions of compiler code without more semantic information about why the statement is suspicious seem unlikely to work well, based on our experience with complex compiler faults, and the results of Parnin and Orso \cite{AutoHelp}.
-}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{comment}                                               %
-% \begin{figure}                                                %
-% {\scriptsize                                                  %
-% \begin{code}                                                  %
-% {\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} %
-% ...                                                           %
-% <             vp[1] = OBJECT\_TO\_JSVAL(thisp);               %
-% <         \} else \{                                          %
-% <             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    %
-% <         \}                                                  %
-% ...                                                           %
-% <                   a->avail = (jsuword)sp;                   %
-% <               \}                                            %
-% ...                                                           %
-% ---                                                           %
-% >         RESTORE\_SP(fp);                                    %
-% ...                                                           %
-% {\bf Mutant information:}                                     %
-% Failure repaired by 148 mutants.                              %
-% ...                                                           %
-% \#3: delete statement mutant of jsinterp.c:944:               %
-%                                                               %
-%             vp[1] = OBJECT\_TO\_JSVAL(thisp);                 %
-%         \} else \{                                            %
-% /*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    %
-%         \}                                                    %
-%                                                               %
-% {\bf Coverage change information:}                            %
-% ...                                                           %
-% \#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:      %
-%                    a->avail = (jsuword)sp; /* ADDED */        %
-% \end{code}                                                    %
-% }                                                             %
-% \caption{Patch and explanation for fault 115}                 %
-% %\vspace{-0.18in}                                             %
-% \label{fig:explain}                                           %
-% \end{figure}                                                  %
-% \end{comment}                                                 %
+\begin{comment}                                               %
+\begin{figure}                                                %
+{\scriptsize                                                  %
+\begin{code}                                                  %
+{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} %
+...                                                           %
+<             vp[1] = OBJECT\_TO\_JSVAL(thisp);               %
+<         \} else \{                                          %
+<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    %
+<         \}                                                  %
+...                                                           %
+<                   a->avail = (jsuword)sp;                   %
+<               \}                                            %
+...                                                           %
+---                                                           %
+>         RESTORE\_SP(fp);                                    %
+...                                                           %
+{\bf Mutant information:}                                     %
+Failure repaired by 148 mutants.                              %
+...                                                           %
+\#3: delete statement mutant of jsinterp.c:944:               %
+                                                              %
+            vp[1] = OBJECT\_TO\_JSVAL(thisp);                 %
+        \} else \{                                            %
+/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    %
+        \}                                                    %
+                                                              %
+\end{code}                                                    %
+}                                                             %
+\caption{Patch and explanation for fault 115}                 %
+%\vspace{-0.18in}                                             %
+\label{fig:explain}                                           %
+\end{figure}                                                  %
+\end{comment}                                                 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        %
-% Figure \ref{fig:explain} shows part of the patch for SpiderMonkey fault 115, and the {\bf Coverage} and {\bf Repair} outputs that successfully localize the fault (the 6th coverage change output, and the 3rd mutant output), to give some idea of the information our methods present.
-For {\bf Coverage}, it is surprising that the sixth highest ranked change only appeared in 5 of 148 repairs.
-This change ranked high out of thousands of changes because it was unusual, appearing for only 10 repairs in the entire SpiderMonkey analysis.
+ Figure \ref{fig:explain} shows part of the patch for SpiderMonkey fault 115, and the {\bf Repair} output that successfully localizes the fault (the 3rd mutant output), to give some idea of the information our method presents.
 %
-%                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        %
-% The precise form an explanation based on our techniques should take is not clear.
-One possibility, suggested by the results here, is a method focused on repairs, but using common coverage changes.
-For each repairing mutant, in rank order, the highest ranking coverage changes for that mutant could also be shown, to give an idea of unusual (and potentially localizing) changes induced by repair.
-%
-% \end{comment}                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-01-20 journal/localizations.tex
more text

--- 2018-07-16 localizations.tex
@@
 
-\comment{
-By definition, when a mutant repairs a test case, it changes the
-behavior of the test.  In some cases, this behavioral change is
-limited to changes in program state and data values.  However, the
-vast majority of mutant repairs also modify the code coverage of a
-test; in our experiments, there were no repairs that did not change
-coverage.  If a fault can be avoided by, for example, disabling a
-problematic optimization in a compiler, this naturally leads to
-changed coverage.  If the fault is due to a bad conditional, where a
-program path should include additional code, the repair will often
-change the execution to include the omitted code.  As a localization
-for a fault, then, we can examine the source statements whose
-  coverage changed most often in repairs.  Early experiments with this
-  approach showed that a useful coverage change localization must  account for the frequency of a
-change.  Some coverage changes are extremely common across all
-failures, faults, and repairing mutants,
-and therefore likely not related to any specific fault.
-
-\begin{quote}
-The {\bf Coverage} localization for a failing test case $f \in F$ ($F$
-is the set of all failing test cases) ranks
-statements by the function $r$ (for rank) where:
-%$$U = s : \forall m . \forall t \in F . c(m,t,s) = 1$$
-$$C(s) = \sum_{t \in F} (\sum_{m \in R(t)} c(m,t,s))$$
-%\[
-$$r(s,f) = \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s)$$
-%\begin{cases}
-% \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s) & s \not\in U\\
-%   0 & s \in U
-%\end{cases}
-%\]
-$c(m,t,s)$ is 1 if coverage of $s$ changes for test $t$
-with mutant $m$, and 0 otherwise.  $R(f)$ is the set of mutants repairing $f$.
-\end{quote}
-
-That is, the ranking is based on the number of times a statement's
-coverage changes its value (covered or not covered) from the original
-execution of $f$ in executions of mutants that repair $f$.  Changes are weighted by their inverse frequency over all
-failures and all mutants.
-Higher values for $r$ result in a higher ranking (more likely to
-be faulty).  
 %Computing $U$ accounts for the fact that in our
@@
 %r
-}
+
 Repairing mutants can be used for localization and
@@
 
-\comment{
+\subsection{Coverage-Based Localizaton}
+
+By definition, when a mutant repairs a test case, it changes the
+behavior of the test.  In some cases, this behavioral change is
+limited to changes in program state and data values.  However, the
+vast majority of mutant repairs also modify the code coverage of a
+test; in our experiments, there were no repairs that did not change
+coverage.  If a fault can be avoided by, for example, disabling a
+problematic optimization in a compiler, this naturally leads to
+changed coverage.  If the fault is due to a bad conditional, where a
+program path should include additional code, the repair will often
+change the execution to include the omitted code.  As a localization
+for a fault, then, we can examine the source statements whose
+  coverage changed most often in repairs.  Early experiments with this
+  approach showed that a useful coverage change localization must  account for the frequency of a
+change.  Some coverage changes are extremely common across all
+failures, faults, and repairing mutants,
+and therefore likely not related to any specific fault.
+
+\begin{quote}
+The {\bf Coverage} localization for a failing test case $f \in F$ ($F$
+is the set of all failing test cases) ranks
+statements by the function $r$ (for rank) where:
+%$$U = s : \forall m . \forall t \in F . c(m,t,s) = 1$$
+$$C(s) = \sum_{t \in F} (\sum_{m \in R(t)} c(m,t,s))$$
+%\[
+$$r(s,f) = \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s)$$
+%\begin{cases}
+% \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s) & s \not\in U\\
+%   0 & s \in U
+%\end{cases}
+%\]
+$c(m,t,s)$ is 1 if coverage of $s$ changes for test $t$
+with mutant $m$, and 0 otherwise.  $R(f)$ is the set of mutants repairing $f$.
+\end{quote}
+
+That is, the ranking is based on the number of times a statement's
+coverage changes its value (covered or not covered) from the original
+execution of $f$ in executions of mutants that repair $f$.  Changes are weighted by their inverse frequency over all
+failures and all mutants.
+Higher values for $r$ result in a higher ranking (more likely to
+be faulty).
+
 The coverage localization also provides a kind of explanation, though
@@
 on a single failing test case, which in most cases fails due to only one
-fault.  In fact, other faults help both methods rank localizations.}
+fault.  In fact, other faults help both methods rank localizations.

### 2019-01-20 journal/relatedwork.tex
more text

--- 2018-07-17 relatedwork.tex
@@
 
-%There are four primary areas of related work.
+There are four primary areas of related work.
 First, fuzzer taming and related problems of determining the set of faults from a large set of redundant failures have been investigated in a few papers.
@@
 Recently, fault localization researchers have used program mutants to improve statistical \cite{Jones2002} fault localization techniques \cite{MUSE,DebroyMutant,FasterMutant,multilingual,Metallaxis,Papadakis}.
-%The core of this work has been the belief that faulty program locations are (1) more likely to change a failing test to successful when mutated and (2) less likely to change a successful test case to faulty when mutated.
+The core of this work has been the belief that faulty program locations are (1) more likely to change a failing test to successful when mutated and (2) less likely to change a successful test case to faulty when mutated.
 This clearly connects to our general notion that how two test cases respond to a set of mutants is a good measure of their semantic similarity.

### 2019-01-21 journal/experiments.tex
experiments
bring in coverage localization

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{comment}                                               %
 \begin{figure}                                                %
@@
         \}                                                    %
-                                                              %
+
+{\bf Coverage change information:}
+...
+\#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:
+                   a->avail = (jsuword)sp; /* ADDED */                                                            %
 \end{code}                                                    %
@@
 \end{figure}                                                  %
-\end{comment}                                                 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        %
- Figure \ref{fig:explain} shows part of the patch for SpiderMonkey fault 115, and the {\bf Repair} output that successfully localizes the fault (the 3rd mutant output), to give some idea of the information our method presents.
                                                                                                                                                                                                                                                                              %
@@
 
+\subsection{Coverage Localization}
+
+While {\bf Repair} is often a high-quality localization technique, it often fails to localize a fault at all.
+This failure suggested that for the compiler setting, we should consider other localizations methods, even ones we do not consider likely to be helpful in more general settings.
+Compiler faults are often so complex and hard to debug that even a somewhat informative localization might be useful.
+
+By definition, when a mutant repairs a test case, it changes the
+behavior of the test.  In some cases, this behavioral change may be
+limited to changes in program state and data values.  However, the
+vast majority of mutant repairs \emph{for a compiler} can be expected to also modify the code coverage of a
+test; in our compiler experiments, there were no repairs that did not change
+coverage; this would be very unlikely to hold in other software settings, where many bugs related to data values; for well-fuzzed compilers, however, few bugs are simple, e.g., off-by-one bugs in an integer value.
+If a fault can be avoided by disabling a
+problematic optimization in a compiler, this naturally leads to
+changed coverage.  If the fault is due to a bad conditional, where a
+program path should include additional code, the repair will often
+change the execution to include the omitted code.  As a localization
+for a fault, then, we can examine the source statements whose
+coverage changed most often in repairs.  Early experiments with this
+ approach showed that a useful coverage change localization must  account for the frequency of a
+change.  Some coverage changes are extremely common across all
+failures, faults, and repairing mutants,
+and therefore likely not related to any specific fault.
+
+\begin{quote}
+The {\bf Coverage} localization for a failing test case $f \in F$ ($F$
+is the set of all failing test cases) ranks
+statements by the function $r$ (for rank) where:
+%$$U = s : \forall m . \forall t \in F . c(m,t,s) = 1$$
+$$C(s) = \sum_{t \in F} (\sum_{m \in R(t)} c(m,t,s))$$
+%\[
+$$r(s,f) = \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s)$$
+%\begin{cases}
+% \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s) & s \not\in U\\
+%   0 & s \in U
+%\end{cases}
+%\]
+$c(m,t,s)$ is 1 if coverage of $s$ changes for test $t$
+with mutant $m$, and 0 otherwise.  $R(f)$ is the set of mutants repairing $f$.
+\end{quote}
+
+That is, the ranking is based on the number of times a statement's
+coverage changes its value (covered or not covered) from the original
+execution of $f$ in executions of mutants that repair $f$.  Changes are weighted by their inverse frequency over all
+failures and all mutants.
+Higher values for $r$ result in a higher ranking (more likely to
+be faulty).
+
+The coverage localization also provides a kind of explanation, though
+a less direct one than with a repair.  For any
+interesting line of code in the coverage-change-based localization,
+the user can investigate a particular instance of the change,
+by selecting a single mutant and stepping through the execution of the
+failing test case in a debugger to the point of change.  The user can also
+browse through all mutants that produce this coverage change for the test
+case.
+
+\begin{figure}                                                %
+{\scriptsize                                                  %
+\begin{code}                                                  %
+{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} %
+...                                                           %
+<             vp[1] = OBJECT\_TO\_JSVAL(thisp);               %
+<         \} else \{                                          %
+<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    %
+<         \}                                                  %
+...                                                           %
+<                   a->avail = (jsuword)sp;                   %
+<               \}                                            %
+...                                                           %
+---                                                           %
+>         RESTORE\_SP(fp);                                    %
+...                                                           %
+{\bf Mutant information:}                                     %
+Failure repaired by 148 mutants.                              %
+...                                                           %
+\#3: delete statement mutant of jsinterp.c:944:               %
+                                                              %
+            vp[1] = OBJECT\_TO\_JSVAL(thisp);                 %
+        \} else \{                                            %
+/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    %
+        \}                                                    %
+
+{\bf Coverage change information:}
+...
+\#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:
+                   a->avail = (jsuword)sp; /* ADDED */                                                            %
+\end{code}                                                    %
+}                                                             %
+\caption{Patch and explanation for fault 115}                 %
+%\vspace{-0.18in}                                             %
+\label{fig:explain}                                           %
+\end{figure}                                                  %
+
+Figure \ref{fig:explain} shows part of the patch for SpiderMonkey fault 115, and the {\bf Repair} and {\bf Coverage} outputs that successfully localize the fault (the 3rd mutant output, and the 6th coverage change output), to give some idea of the information our methods present.
+For {\bf Coverage}, it is surprising that the sixth highest ranked change only appeared in 5 of 148 repairs.
+This change ranked high out of thousands of changes because it was unusual, appearing for only 10 repairs in the entire SpiderMonkey analysis.
+
 \input{othertools}
@@
 
-\subsection{Mutation Analysis Costs}
+\section{Mutation Analysis Costs}
 
@@
 
+\subsection{Hot and Cold Running Mutants}
+
 \subsection{Threats to Validity}

### 2019-01-21 journal/experiments.tex
coverage and placholder

@@
 
-\begin{figure}                                                %
-{\scriptsize                                                  %
-\begin{code}                                                  %
-{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} %
-...                                                           %
-<             vp[1] = OBJECT\_TO\_JSVAL(thisp);               %
-<         \} else \{                                          %
-<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    %
-<         \}                                                  %
-...                                                           %
-<                   a->avail = (jsuword)sp;                   %
-<               \}                                            %
-...                                                           %
----                                                           %
->         RESTORE\_SP(fp);                                    %
-...                                                           %
-{\bf Mutant information:}                                     %
-Failure repaired by 148 mutants.                              %
-...                                                           %
-\#3: delete statement mutant of jsinterp.c:944:               %
-                                                              %
-            vp[1] = OBJECT\_TO\_JSVAL(thisp);                 %
-        \} else \{                                            %
-/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    %
-        \}                                                    %
+\begin{figure}                                                
+{\scriptsize
+
+\begin{code}                                                  
+{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} 
+...                                                           
+<             vp[1] = OBJECT\_TO\_JSVAL(thisp);               
+<         \} else \{                                          
+<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    
+<         \}                                                  
+...                                                           
+<                   a->avail = (jsuword)sp;                   
+<               \}                                            
+...                                                           
+---                                                           
+>         RESTORE\_SP(fp);                                    
+...                                                           
+{\bf Mutant information:}                                     
+Failure repaired by 148 mutants.                              
+...                                                           
+\#3: delete statement mutant of jsinterp.c:944:
+
+            vp[1] = OBJECT\_TO\_JSVAL(thisp);                 
+        \} else \{                                            
+/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    
+        \}
 
@@
 \#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:
-                   a->avail = (jsuword)sp; /* ADDED */                                                            %
-\end{code}                                                    %
-}                                                             %
-\caption{Patch and explanation for fault 115}                 %
-%\vspace{-0.18in}                                             %
-\label{fig:explain}                                           %
-\end{figure}                                                  %
+                   a->avail = (jsuword)sp; /* ADDED */                                                            
+\end{code}                                                    
+}
+\caption{Patch and explanation for fault 115}                 
+%\vspace{-0.18in}                                             
+\label{fig:explain}                                           
+\end{figure}
 
@@
 This change ranked high out of thousands of changes because it was unusual, appearing for only 10 repairs in the entire SpiderMonkey analysis.
+
+\begin{table}
+\centering
+{\scriptsize
+\begin{tabular}{|c||c|c|}
+\hline
+& \multicolumn{2}{|c|}{Localization Rank} \\
+\hline
+Fault & Coverage & Repair \\
+\hline
+880 & 2,255 & - \\
+95 & 720 & - \\
+1294 & 190 & 1 \\
+60 &  - & - \\
+1543 & 18 & 19 \\
+1172 & 46 & 28 \\
+1561 & - & - \\
+115 & 6 & 3 \\
+\hline
+\end{tabular}
+}
+\caption{Coverage localization results for SpiderMonkey}
+%\vspace{-0.2in}
+\label{covjstable}
+\end{table}
+
+\begin{table}
+\centering
+{\scriptsize
+\begin{tabular}{|c||c|c|}
+\hline
+& \multicolumn{2}{|c|}{Localization Rank} \\
+\hline
+Fault &Coverage & Repair \\
+\hline
+139094 & 44,561 & - \\
+133004 & 75 & 1 \\
+158782 & - & - \\
+156496 & 3,067 & - \\
+143677 & 186 & - \\
+158555 & 490 & - \\
+134321 & - & - \\
+133940 & 260 & 4 \\
+141195 & 465 & - \\
+156795 & 19 & - \\
+139709 & 43,903 & 10 \\
+155698 & 85 & 1 \\
+138646 & 2 & - \\
+140795 & 302 & 34 \\
+136501 & 223 & 5\\
+134322 & 639 & - \\
+\hline
+\end{tabular}
+}
+\caption{Coverage localization results for GCC}
+\label{covgcctable}
+\end{table}
 

### 2019-01-21 journal/localizations.tex
fix localize

@@
 
-\subsection{Coverage-Based Localizaton}
-
-By definition, when a mutant repairs a test case, it changes the
-behavior of the test.  In some cases, this behavioral change is
-limited to changes in program state and data values.  However, the
-vast majority of mutant repairs also modify the code coverage of a
-test; in our experiments, there were no repairs that did not change
-coverage.  If a fault can be avoided by, for example, disabling a
-problematic optimization in a compiler, this naturally leads to
-changed coverage.  If the fault is due to a bad conditional, where a
-program path should include additional code, the repair will often
-change the execution to include the omitted code.  As a localization
-for a fault, then, we can examine the source statements whose
-  coverage changed most often in repairs.  Early experiments with this
-  approach showed that a useful coverage change localization must  account for the frequency of a
-change.  Some coverage changes are extremely common across all
-failures, faults, and repairing mutants,
-and therefore likely not related to any specific fault.
-
-\begin{quote}
-The {\bf Coverage} localization for a failing test case $f \in F$ ($F$
-is the set of all failing test cases) ranks
-statements by the function $r$ (for rank) where:
-%$$U = s : \forall m . \forall t \in F . c(m,t,s) = 1$$
-$$C(s) = \sum_{t \in F} (\sum_{m \in R(t)} c(m,t,s))$$
-%\[
-$$r(s,f) = \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s)$$
-%\begin{cases}
-% \sum_{m \in R(f)}\frac{1}{C(s)}c(m,f,s) & s \not\in U\\
-%   0 & s \in U
-%\end{cases}
-%\]
-$c(m,t,s)$ is 1 if coverage of $s$ changes for test $t$
-with mutant $m$, and 0 otherwise.  $R(f)$ is the set of mutants repairing $f$.
-\end{quote}
-
-That is, the ranking is based on the number of times a statement's
-coverage changes its value (covered or not covered) from the original
-execution of $f$ in executions of mutants that repair $f$.  Changes are weighted by their inverse frequency over all
-failures and all mutants.
-Higher values for $r$ result in a higher ranking (more likely to
-be faulty).
-
-The coverage localization also provides a kind of explanation, though
-a less direct one.  For any
-interesting line of code in the coverage-change-based localization,
-the user can investigate a particular instance of the change,
-by selecting a single mutant and stepping through the execution of the
-failing test case in a debugger to the point of change.  The user can also
-browse through all mutants that produce this coverage change for the test
-case.
-
 Another advantage over many statistical approaches to 
@@
 fault.  In fact, other faults help both methods rank localizations.
+
+%\subsection{Coverage-Based Localizaton}

### 2019-01-22 journal/experiments.tex
move figure
more progress

--- 2019-01-21 journal/experiments.tex
@@
 the more extensive Proteum \cite{Proteum} operators used in MUSE/MUSEUM's evaluations.
-\comment{Even with these restrictions, running all mutants on passing tests until at least a single kill was observed took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.}  In Section \ref{sec:otherfaults} we provide some comparisons with MUSEUM and MUSE fault localization in the context of a full mutation result matrix using their preferred set of operators.
+Even with these restrictions, running all mutants on passing tests until at least a single kill was observed took about 3 times as long as the search for repairs, which we already consider a potentially problematic cost.
+In Section \ref{sec:otherfaults} we provide some comparisons with MUSEUM and MUSE fault localization in the context of a full mutation result matrix using their preferred set of operators.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}                                                %
-{\scriptsize                                                  %
-\begin{code}                                                  %
-{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} %
-...                                                           %
-<             vp[1] = OBJECT\_TO\_JSVAL(thisp);               %
-<         \} else \{                                          %
-<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    %
-<         \}                                                  %
-...                                                           %
-<                   a->avail = (jsuword)sp;                   %
-<               \}                                            %
-...                                                           %
----                                                           %
->         RESTORE\_SP(fp);                                    %
-...                                                           %
-{\bf Mutant information:}                                     %
-Failure repaired by 148 mutants.                              %
-...                                                           %
-\#3: delete statement mutant of jsinterp.c:944:               %
-                                                              %
-            vp[1] = OBJECT\_TO\_JSVAL(thisp);                 %
-        \} else \{                                            %
-/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    %
-        \}                                                    %
 
-{\bf Coverage change information:}
-...
-\#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:
-                   a->avail = (jsuword)sp; /* ADDED */                                                            %
-\end{code}                                                    %
-}                                                             %
-\caption{Patch and explanation for fault 115}                 %
-%\vspace{-0.18in}                                             %
-\label{fig:explain}                                           %
-\end{figure}                                                  %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}                                                %
-{\scriptsize                                                  %
-\begin{code}                                                  %
-{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} %
-...                                                           %
-<             vp[1] = OBJECT\_TO\_JSVAL(thisp);               %
-<         \} else \{                                          %
-<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    %
-<         \}                                                  %
-...                                                           %
-<                   a->avail = (jsuword)sp;                   %
-<               \}                                            %
-...                                                           %
----                                                           %
->         RESTORE\_SP(fp);                                    %
-...                                                           %
-{\bf Mutant information:}                                     %
-Failure repaired by 148 mutants.                              %
-...                                                           %
-\#3: delete statement mutant of jsinterp.c:944:               %
-                                                              %
-            vp[1] = OBJECT\_TO\_JSVAL(thisp);                 %
-        \} else \{                                            %
-/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    %
-        \}                                                    %
+\begin{figure}                                                
+{\scriptsize                                                  
+\begin{code}                                                  
+{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} 
+...                                                           
+<             vp[1] = OBJECT\_TO\_JSVAL(thisp);               
+<         \} else \{                                          
+<             ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);    
+<         \}                                                  
+...                                                           
+<                   a->avail = (jsuword)sp;                   
+<               \}                                            
+...                                                           
+---                                                           
+>         RESTORE\_SP(fp);                                    
+...                                                           
+{\bf Mutant information:}                                     
+Failure repaired by 148 mutants.                              
+...                                                           
+\#3: delete statement mutant of jsinterp.c:944:
+
+            vp[1] = OBJECT\_TO\_JSVAL(thisp);                 
+        \} else \{                                            
+/*MUTANT    ok = OBJ\_GET\_PROPERTY(cx, thisp, id, \&v);*/    
+        \}
 
@@
 \#6: 5 mutants (3.38\% of repairs) added jsinterp.c:992:
-                   a->avail = (jsuword)sp; /* ADDED */                                                            %
-\end{code}                                                    %
-}                                                             %
-\caption{Patch and explanation for fault 115}                 %
-%\vspace{-0.18in}                                             %
-\label{fig:explain}                                           %
-\end{figure}                                                  %
+                   a->avail = (jsuword)sp; /* ADDED */                                                            
+\end{code}                                                    
+}                                                             
+\caption{Patch and explanation for fault 115}                 
+%\vspace{-0.18in}                                             
+\label{fig:explain}                                           
+\end{figure}
 
@@
 This change ranked high out of thousands of changes because it was unusual, appearing for only 10 repairs in the entire SpiderMonkey analysis.
+
+\begin{table}
+\centering
+{\scriptsize
+\begin{tabular}{|c||c|c||c||c|c|}
+\hline
+\multicolumn{3}{|c|}{SpiderMonkey}&\multicolumn{3}{|c|}{GCC}\\
+\hline
+ID & Coverage & Repair & ID & Coverage & Repair\\
+\hline
+1 & 2,255 & - & 1 & 44,561 & -\\
+2 & 720 & - & 2 & 75 & 1\\
+3 & 190 & 1 & 3 & - & -\\
+4 &  - & - & 4 & 3,067 & -\\
+5 & 18 & 19 & 5 & 186 & - \\
+6 & 46 & 28 & 6 & 490 & - \\
+7 & - & - & 7 & - & -\\
+8 & 6 & 3 & 8 &260 & 4\\
+   &    &    & 9 & 465 & - \\
+   &    &    & 10 & 19 & - \\
+   &    &    & 11 & 43,903 & 10 \\
+   &    &    & 12 & 85 & 1 \\
+   &    &    & 13 & 2 & - \\
+   &    &    & 14 & 302 & 34 \\
+   &    &    & 15 & 223 & 5 \\
+   &    &    & 16 & 639 & - \\
+\hline
+\end{tabular}
+}
+\caption{Coverage localization results}
+%\vspace{-0.2in}
+\label{covtable}
+\end{table}
+
+Table \ref{covtable} compares the {\bf Coverage} localization with the {\bf Repair} localization.
+In general, obviously, {\bf Coverage} is not as effective as {\bf Repair}.
+In fact, in cases where both methods provide any localization information, there is only one instance where {\bf Coverage} is better than {\bf Repair}.
+On the other hand, {\bf Coverage} provides a localization for 10 faults where {\bf Repair} offers no assistance.
+The rankings for these suggestions are often poor (indeed, in some GCC cases, they are almost ludicrously bad).
+However, for GCC faults 10 and 13, {\bf Coverage} produced a useful localization when all of the other methods (including MUSE and MUSEUM) failed to localize at all.
+Because compiler faults are so difficult to understand and fix, it might be useful for developers to examine the top few localizations (that differ) for multiple methods.
+While MUSE and MUSEUM both generally perform better than {\bf Coverage}, they almost never provide better information than {\bf Repair}.
+In every case where MUSE or MUSEUM provides a useful localization (within top-25) and performs better than {\bf Repair}, the difference in ranking between MUSE or MUSEUM and {\bf Repair} is five or fewer ranking positions.
+{\bf Coverage} uses a sufficiently different approach that it offers potentially useful information, even when {\bf Repair} performs badly.
 
@@
 In practice, their work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique, since trying multiple metrics is impractical: a user has to examine each curve.
-\comment{It also included methods such as comparing output signatures that are inapplicable to differential testing in most cases, the most critical compiler application.}  This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work; our improvements over any single ``reasonable'' method applied to both problems would be considerably larger (for instance, using the most obvious basis, line coverage, we see more than 17\% improvement).
+The set of metrics also included methods such as comparing output signatures that are inapplicable to differential testing for wrong-code faults, the most critical compiler application.
+This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work; our improvements over any single ``reasonable'' method applied to both problems would be considerably larger (for instance, using the most obvious basis, line coverage, we see more than 17\% improvement).
 Moreover, the improvement is, for our subject programs, reliable:  choosing random subsets of the full data set, the difference in mean APFD from best-previous method is around 3\%, and statistically significant.
@@
 
-\comment{
- Why does {\bf Repair} perform best in this setting?
+%\comment{
+ Why does {\bf Repair} perform best in the \emph{compiler} setting in particular?
 It is impossible to be sure, but one  possibility is power to distinguish repairs.
@@
 For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.
-We applied our repair localization to the MUSEUM data set, using the full Proteum mutant set and added MUSEUM-specific mutants, and for all but one of the 8 bugs in the data set, our results were equivalent to those for MUSEUM (in one case our approach fails to produce a localization).}
+Our suspicion is somewhat confirmed by the results below where we used our methods with the MUSEUM mutant result set.
+
+Finally, in contrast to many previously reported results in fault localization (which are typically over much simpler programs and faults than subtle compiler semantics bugs) all methods we applied very frequently failed to provide a useful localization at all for our compiler faults.
+This is presumably due to the sheer difficulty of compiler bugs:  our approach, and MUSEUM, performed very well on complex multi-lingual faults in the MUSEUM data set.
+In one sense, {\bf Coverage} is the most intriguing method presented here.
+While it did not perform nearly as well as {\bf Repair}, it did manage, in many cases, to rank \emph{some} faulty line highly, relative to the extremely large number of statements executed in each failing GCC run (on the order of 50K-100K+ statements).
+In future work, we plan to experiment with different methods for weighting frequency of coverage, perhaps using methods from machine learning, or of pruning changed coverage that is not likely to be relevant.
+For example, perhaps the {\bf Repair} approach can be applied, and coverage changes present in very distant failures can be removed, though the argument is less clear, since a coverage change isn't directly associated with the onion-ring structure, and does not directly repair a set of failures.
+In fact, since a study of the importance of delta-debugging in localization \cite{ReduceLocalize} showed that even traditional Spectrum-Based Fault Localization can \emph{sometimes} produce good results on SpiderMonkey faults (though often performing worse than even {\bf Coverage}), we suspect that combining multiple methods, using some method to cross-assess confidence might be important for obtaining reliably good compiler bug localizations.
+There is much to be done if automated debugging is ever to make the lives of compiler developers easier in most cases \cite{AutoHelp}.
 
@@
 
-\comment{
-Finally, in contrast to many previously reported results in fault localization, typically over much simpler programs and faults than subtle compiler semantics bugs, all methods frequently failed to provide a useful localization at all.
-This is presumably due to the difficulty of compiler bugs:  our approach, and MUSEUM, performed very well on complex multi-lingual faults in the MUSEUM data set.
-}
 \comment{ In one sense, {\bf Coverage} is the most intriguing method presented here.
@@
 Running each test case under each mutated version of the compiler is cheap: each execution requires around 0.05 seconds for SpiderMonkey, and 0.12 seconds for GCC, on average; moreover, executions can be done in parallel, most failures do not cover most mutants, and vice versa, so the total number of repair checks needed is far less than the product of mutants and failures, and can be spread over many machines.
-\comment{The practical approach to fuzzer taming for a large, ongoing project is likely to maintain a set of ``hot'' pre-compiled mutants.
-These can be pruned at will by running them on passing test cases, and removing mutants that fail on any test case.
-}
 
@@
 E.g., trivial compiler equivalence (TCE) \cite{TCE} can reject equivalent and redundant mutants, removing almost 30\% of mutants for benchmark subjects.
-Other techniques, such as using mutant schemas to avoid having to compile and store each mutant, are also applicable --- almost all of the techniques characterized by Offutt and Untch \cite{offutt2001mutation} as \emph{do faster} approaches apply.
-Whether sampling mutants \cite{RahulISSRE} is applicable is less clear.
-Even if the FPF curve remains the same, a critical mutant for localization might be omitted.
-We performed some limited experiments, and found that for a few sample sizes ranging from 13\% of mutants to 50\% of mutants or using only some mutation operators (with the exception of operator-replacement-only), APFD was always better than benchmark set's best metric.
-However, the improvement in some cases was \emph{much} smaller than for the full set, and varied considerably (in some cases it was up to 2\% better than our full results).
-Fault localization results were much worse, however (with relative rankings preserved) for both our techniques and MUSE/MUSEUM.
-\comment{Finally, guided sampling may provide good fuzzer taming (if not fault localization):  compute vectors over a small constant-sized \cite{RahulISSRE} sample of mutants, then report highly distinct failing tests.
-Other metrics, such as coverage-based comparisons, could provide hints as to likely distinct tests (and thus mutants covered by them) to focus on first.
-These are likely to be different faults.
-After repairing some of these, choose mutants to evaluate that are covered by multiple tests with few or no repairing mutants (since such tests are the ones most likely to be distinct but not yet identified).
-Such an approach could be quite cost-effective, requiring only a relatively small number of mutant evaluations before providing useful results.}
+Other techniques, such as using mutant schemas to avoid having to compile and store each mutant, are also applicable --- almost all of the techniques characterized by Offutt and Untch \cite{offutt2001mutation} as \emph{do faster} approaches apply, without any loss of ability to identify or localize faults.
 
@@
 
+Note that the largest cost in our approach, by far, is compiling the mutants.
+While running all the GCC tests on the covered, compilable, test-passing mutants takes nearly 24 days of total execution time, the process is embarrasingly parallel:  no mutant/test execution depends on any other.
+On the other hand, all executions of a mutant/test pair depend on having a compiler that includes that mutant built, and compiling GCC requires more time than test executions by a large margin.
+
+We expect that fuzzer taming will mostly be performed when someone is fuzzing a compiler either for the first time or using a new technique.
+If fuzzing is performed on a well-fuzzed compiler, with only incremental changes from previous versions, there is not likely to be a significant fuzzer taming problem:  fuzzing will often only detect a small number of faults, and manual triage efforts will probably suffice.
+In the case of ``new'' fuzzing efforts, however, they will usually be aimed at release versions of a compiler.
+For major projects such as GCC, LLVM, or V8, producing and publishing a set of ``hot'' mutants might be one solution.
+\emph{Hot} mutants are mutants that 1) compile 2) pass all current tests and 3) are not known to be equivalent mutants.
+Producing such mutants should be an easily automatable task without prohibitive cost, at the level of a major open source project.
+Release versions are only produced a few times a year (e.g., both GCC and LLVM had 5 releases in 2018 and 4 in 2017).
+In addition to fuzzer taming, such mutants would provide a useful quick benchmark for new compiler-fuzzing techniques:  if a technique can kill mutants not detectable by the compiler's test suite, it has possibilities for bug detection as well.
+
+\subsection{Lossy Techniques}
+
+Mutation cost reduction techniques classified as \emph{do fewer} on the other hand involve a loss of signal, not just more efficient production of the same signal.
+Figures \ref{samplejs} and \ref{samplegcc} show, respectively, the results of random sampling of mutants for SpiderMonkey and GCC, by comparing the APFD obtained on a set of 30 random samples of the chosen percentage of mutants with the APFD for the full mutant set.
+Because some mutants are actually misleading with respect to distinguishing faults, sometimes sampling can improve results.
+Sampling 10\% of mutants directly translates to our approach requiring about 90\% less time, or less computing power if run in parallel.
+
+Pure random sampling is fairly effective for GCC, but more mutants are required to obtain a comparable APFD on average, for SpiderMonkey.
+
 \subsection{Threats to Validity}
 
-The largest threat to validity is that this paper's conclusions rely on only two data sets for compilers, and a small number of non-compiler faults.
+The largest threat to validity is that this paper's core conclusions rely on only two data sets for compilers, and a small number of non-compiler faults.
 The compiler ground truths are possibly imperfect, due to the complex change history, and it was not possible to produce good patches for all faults, limiting our analysis of compiler fault localization.

### 2019-01-22 journal/introduction.tex
journal start
mutants16 paper
more progress

--- 2018-03-02 introduction.tex
@@
 
-In a PLDI 2013 paper, Chen et al.
+Chen et al.
 defined the \emph{fuzzer taming problem}: ``Given a potentially large collection of test cases, each of which triggers a bug, rank them in such a way that test cases triggering distinct bugs are early in the list.'' \cite{PLDI13}.
@@
 
-%\subsection{The Onion-Ring and Causality}
+\subsection{The Onion-Ring and Causality}
 
@@
 
-%\subsection{Contributions}
+\subsection{Contributions}
 
-The contributions of this paper include our ideas about how to measure distance between program executions, but more importantly some practical proposals for using such a metric in compiler fuzzer taming and fault localization.
+This paper extends our 2018 ISSRE paper \cite{ISSRE18Mutants} introducing our approach.
+That work, and this, suggested a novel way of measuring distance between program executions, but more importantly provided some practical proposals for using such a metric in compiler fuzzer taming and fault localization.
 Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over best previous results \cite{PLDI13}, in multiple ways.
 Our fault localization approach also compares favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual,Metallaxis}, albeit with a more limited applicability (but also a lower cost).
+We extend the earlier paper by:
+
+\begin{itemize}
+\item providing an extended example to help readers understand the metric, FPF method, and fault localization approach,
+\item introducing and evaluating another compiler fault localization approach, based on code coverage changes,
+\item and including experimental results summarizing the effectiveness and cost reduction provided by various optimizations of our approach.
+\end{itemize}

### 2019-01-22 journal/relatedwork.tex
more progress

@@
 \cite{multilingual} on MUSEUM and Moon et al.
-\cite{MUSE} on MUSE provides a good summary of other work along these lines.
-\comment{, and is perhaps most similar to ours in assumptions, in particular ``Conjecture 1'' that a test case that used to fail on a program is more likely to pass on a mutation of a faulty statement.
-While we suggest some localizations as a welcome and interesting side effect of our distance metric and FPF analysis, our primary goal is the discovery of all faults.} Much fault localization work is largely (in many cases exclusively) based on single-fault scenarios; e.g., the MUSE \cite{MUSE} evaluation includes only 2 of 14 scenarios with multiple faults, and those scenarios include only a few faults.
+\cite{MUSE} on MUSE provides a good summary of other work along these lines, and is perhaps most similar to ours in assumptions.
+In particular we agree with ``Conjecture 1'' that a test case that used to fail on a program is more likely to pass on a mutation of a faulty statement.
+However, while we suggest some localizations as a welcome and interesting side effect of our distance metric and FPF analysis, our primary goal is the discovery of all faults.
+Much fault localization work is largely (in many cases exclusively) based on single-fault scenarios; e.g., the MUSE \cite{MUSE} evaluation includes only 2 of 14 scenarios with multiple faults, and those scenarios include only a few faults.
 The statistical approaches may owe some of their superiority in experiments over other methods to the use of mostly single-fault evaluations \cite{Jones05}.
 By focusing on a single failing test case (for us, selected by fuzzer taming) as the locus of analysis, our work, like some early efforts \cite{NearNeighbor,Cleve05,GroceError} may be more suited for situations involving many faults.
-\comment{MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single failure.
-}
+MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single failure.
 
@@
 
-\comment{
 Some approaches to localization or patching assume the possibility of a mutant/change \emph{actually} fixing a fault.
-In our experiments, we found this to be highly unlikely -- most of our patches were far larger than the largest higher-order-mutants \cite{jia2009higher} anyone usually considers applying, and none were equivalent to a repair.
+In our experiments, we found this to be highly unlikely -- most of our compiler patches were far larger than the largest higher-order-mutants \cite{jia2009higher} anyone usually considers applying, and \emph{none} were equivalent to a repair.
 Previous work has shown that, across a large body of open source programs in many languages, most fault fixes were larger in size than even low-order higher-order mutants \cite{GopinathMutants}.
@@
 Our primary assumption is simply that similar executions respond similarly to mutants, and in particular executions due to the same fault can be ``repaired'' by similar program changes.
-These changes are likely to be \emph{related} in some causal way to the actual fault, but are unlikely to be the actual fault, for reasons of comparative complexity of mutants and real fixes.}
+These changes are likely to be \emph{related} in some causal way to the actual fault, but are unlikely to be the actual fault, for reasons of comparative complexity of mutants and real fixes.
 

### 2019-01-22 journal/experiments.tex
closer to shippable, need all sample results

@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{comparejs}
+  \includegraphics[width=0.95\columnwidth]{comparejs}
   \caption{APFD values for Spidermonkey suite slices}
@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{comparegcc}
+  \includegraphics[width=0.95\columnwidth]{comparegcc}
   \caption{APFD values for GCC suite slices}
@@
 \subsection{Coverage Localization}
+\label{sec:covloc}
 
@@
 \begin{code}                                                  
-{\bf Diff of old ($<$) vs new ($>$) for 115 patch (portion):} 
+{\bf Diff of old ($<$) vs new ($>$) for SpiderMonkey fault 8 patch (portion):} 
 ...                                                           
@@
 For example, perhaps the {\bf Repair} approach can be applied, and coverage changes present in very distant failures can be removed, though the argument is less clear, since a coverage change isn't directly associated with the onion-ring structure, and does not directly repair a set of failures.
-In fact, since a study of the importance of delta-debugging in localization \cite{ReduceLocalize} showed that even traditional Spectrum-Based Fault Localization can \emph{sometimes} produce good results on SpiderMonkey faults (though often performing worse than even {\bf Coverage}), we suspect that combining multiple methods, using some method to cross-assess confidence might be important for obtaining reliably good compiler bug localizations.
-There is much to be done if automated debugging is ever to make the lives of compiler developers easier in most cases \cite{AutoHelp}.
+In fact, a study of the importance of delta-debugging in localization \cite{ReduceLocalize} showed that even traditional Spectrum-Based Fault Localization (SBFL) can \emph{sometimes} produce good results on SpiderMonkey faults.
+Such methods, of course, often perform worse than even {\bf Coverage}; e.g., for our fault \#8 (their fault ID 115), {\bf Repair} and {\bf Coverage} produced localizations at rank 3 and 6, respectively, while the best SBFL result was well over rank 32.
+We suspect that combining multiple methods, using some method to cross-assess confidence might be required to obtain reliably good compiler bug localizations, and assign good confidence scores to localizations.
+There is much to be done if automated debugging is ever to make the lives of compiler developers easier in most cases \cite{AutoHelp}, and a human study using real, complex compiler bugs would be ideal, since previous studies of localization effectiveness tend to focus on simple tasks even moderately experienced developers can potentially succeed at.
+Fixing deep optimizing compiler bugs is a very specialized skill, in contrast to general debugging skills.
 
@@
 
-\section{Mutation Analysis Costs}
+\section{Mutation Analysis Costs and Potential Optimizations}
+\label{sec:highcost}
 
-The compiler experiments required a large computational budget.
+Our compiler experiments required a very large computational budget.
 However, our approach is, by design, cheaper than MUSE or Metallaxis-FL, in that, for a program with up-to-date mutation testing results (just knowing which mutants are killed, not a full matrix), it only requires executing mutants over failing tests.
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\columnwidth]{samplejs}
+  \caption{APFD values for Spidermonkey samples}
+  \label{samplejs}
+\end{figure}%
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\columnwidth]{samplegcc}
+  \caption{APFD values for GCC samples}
+  \label{samplegcc}
+\end{figure}%
+
 Mutation cost reduction techniques classified as \emph{do fewer} on the other hand involve a loss of signal, not just more efficient production of the same signal.
@@
 Because some mutants are actually misleading with respect to distinguishing faults, sometimes sampling can improve results.
-Sampling 10\% of mutants directly translates to our approach requiring about 90\% less time, or less computing power if run in parallel.
+Sampling 10\% of mutants directly translates to our approach requiring about 90\% less time (or less computing power if run in parallel).
+We assume a ``hot'' mutants approach in that our sampling is of compilable, test-passing mutants.
+A larger sample would be required, in proportion to how many mutants compile and pass tests, to achieve the same results in a scenario with only ``cold'' mutants.
 
 Pure random sampling is fairly effective for GCC, but more mutants are required to obtain a comparable APFD on average, for SpiderMonkey.
+The reason for this difference is not easy to understand.
+At a high level, every mutant that repairs at least one failing test test case provides at most one bit of information for fuzzer taming.
+Some repairing mutants repair all test cases (thus providing no information to distinguish faults), and others provide potentially redundant information in that they repair the same set of mutants as another mutant.
+Mutants that repair no test cases provide no information, of course.
+Given that more SpiderMonkey mutants (12.6\%) repaired a test case than did GCC mutants (7.8\%), it would seem that a \emph{smaller} sample of mutants would be required to produce the ``same number of bits of information'' with SpiderMonkey.
+However, SpiderMonkey mutants are more redundant than GCC mutants, and, perhaps more importantly, the SpiderMonkey code is less well-structured, and changes behavior for more causally complex (and unrelated to faults and clear optimization levels) reasons than the more mature GCC code, making each mutant's bit of information inherently more ``noisy.''
+For GCC, using only 30\% of the mutants produces results quite close to the optimal results, and almost never drops below 90\% effectiveness.
+For SpiderMonkey, similar results come with 60\% of mutants.
+Sampling is an inherently incremental process:  mutants can be processed in a random order, and a developer can compute the curve at any point in time.
+We observed that the early curve is reliably good for both SpiderMonkey and GCC, even with as few as 10\% of mutants; most variance is in the further reaches of the curve.
+If developers can identify and fix bugs based on early mutant results, these can be fed back into the process, and mutants not covered by any remaining unfixed tests can be removed from the set.
+For fault localization, with hot mutants a single failing test's complete repair set can be found with relatively few evaluations, and then the repairing mutants can be evaluated on all other tests, to compute a localization that is essentially as accurate as the best ones we produced (distances may be slightly off due to mutants repairing other tests, but not the test being localized, but the sampling is likely to largely handle this issue, since exact distances are not important, only relative ones).
 

### 2019-01-22 journal/introduction.tex
closer to shippable, need all sample results

@@
 \begin{itemize}
-\item providing an extended example to help readers understand the metric, FPF method, and fault localization approach,
-\item introducing and evaluating another compiler fault localization approach, based on code coverage changes,
-\item and including experimental results summarizing the effectiveness and cost reduction provided by various optimizations of our approach.
+\item generally expanding discussion of our approach, its' conceptual basis, and related work,
+\item providing an extended example (in Section \ref{sec:example} to help readers better understand the metric, FPF-based fuzzer taming, and our fault localization approach,
+\item introducing and evaluating another compiler fault localization approach (in Section \ref{sec:covloc}), based on code coverage changes,
+\item and including a much more in-depth examination of the computational costs of our approach and various ways of reducing that cost, including experimental evaluations (in Section \ref{sec:highcost}).
 \end{itemize}

### 2019-01-22 journal/example.tex
closer to shippable, need all sample results
example started, bib fixed

@@
 \section{A Simple Example}
+\label{sec:example}
 
 To better understand the causal approach taken in this paper, it is useful to consider a simplified, abstract example, and understand how the approach proposed would apply to this example.
+
+\begin{figure*}
+\begin{code}
+ 01  void compile(program: feature list, optimizeLevel: int) \{
+ 02    internalRepresentation = initInternalRep(program);
+ 03    if optimizeLevel > 0: \{
+ 04       if feature1 in program:
+ 05          if feature2 in program:
+ 06             corrupt(internalRepresentation);
+ 07       if optimizeLevel > 1: \{
+ 08          if feature3 in program:
+ 09             if feature4 not in program:
+ 10                corrupt(internalRepresentation);
+ 11          if feature5 in program:
+ 12             if feature6 in program:
+ 13                corrupt(internalRepresentation);
+ 14       \}
+ 15    \}
+ 16    /* Will produce wrong code if structure has been corrupted */
+ 17    generateCode(internalRepresentation);
+\end{code}
+\caption{A toy ``optimizing compiler'' with multiple bugs}
+\label{fig:example}
+\end{figure*}
+
+\begin{figure}
+\begin{code}
+Test 1:  (1, 2)
+Test 2:  (3)
+Test 3:  (5, 6)
+\end{code}
+\caption{Failing tests for the toy compiler.}
+\label{fig:toybugs}
+\end{figure}
+
+Consider the pseudo-code ``compiler'' in Figure \ref{fig:example}.
+We can imagine that there is a large amount of code (not shown) in the {\tt initInternalRep} and {\tt generateCode} functions.
+Assume, however, that the only bugs in the compiler are the {\tt corrupt} calls.
+Traditional Spectrum-Based Fault Localization (SBFL) techniques can be easily ``confused'' in multiple-fault settings, and by cases such as this, where there may be strong correlations between code covered by the (fault-free) {\tt initInternalRep} and {\tt generateCode} functions and the features contained in programs, making lines of code far from the fault locations appear suspicous.
+
+Assume that all of our test cases run the compiler with {\tt optimizeLevel} of 2, so as to potentially apply all optimizations.
+A popular way to fuzz a compiler is to compare program behavior when code is compiled with all optimizations with code compiled with no optimizations, so this reflects real-world compiler testing well.
+A set of delta-debugged failing tests, then, might be those shown in Figure \ref{fig:toybugs}.
+
+Further, let us only consider two kinds of mutants:  condition-negation and statement deletion mutants.
+Obviously, negating the condition on line 3 will make all of our failing test cases pass, and negating the condition on line 7 will make test cases 2 and 3 pass, but not test case 1.
+Negating the conditions on lines 4-5, 8-9, and 11-12 will repair our test cases, but is also likely to make some passing test case fail, so these mutants will not be used in our analysis.
+Finally, the only statements that can be deleted without breaking the code are deletions of the {\tt corrupt} calls on lines 6, 10, and 13.
+We can identify these mutants, which will form the basis of our distance metric and our fault localization, by the line number of the change, since there is only one mutant per line in the set.
+Each test can then be associated with a vector of 5 elements:  {\tt (repaired by line 3, repaired by line 6, repaired by line 7, repaired by line 10, repaired by line 13)}.

### 2019-01-22 journal/experiments.tex
better text, better graphs

@@
 We assume a ``hot'' mutants approach in that our sampling is of compilable, test-passing mutants.
-A larger sample would be required, in proportion to how many mutants compile and pass tests, to achieve the same results in a scenario with only ``cold'' mutants.
+A larger sample would be required, in proportion to how many mutants compile and pass tests, to achieve the same results in a scenario with only ``cold'' mutants, though the difference is not huge, since most mutants are ruled out by the fact that no passing test covers the mutant, rather than that they cause some passing test to fail.
+The more significant difference in a cold-mutant scenario is that each mutant/test evaluation is of course much cheaper without the cost (even amortized over many tests) of compilation, not the total number of mutants considered.
 

### 2019-01-22 journal/experiments.tex
small change

@@
 
+Alternatively, a radical way to produce ``hot'' mutants at low cost (the only significant cost being that of executing the test suite to throw out mutants killed by some passing test) is to produce mutants at the \emph{object code} level, which does not require recompiling a system.
+This is how SQLite applies mutation testing:  \url{https://www.sqlite.org/testing.html#mutation_testing}.
+Object level mutants are obviously not useful (or at least, certainly not \emph{as} useful) for fault localization as source-level mutants, but they mutating branches alone probably captures many compiler repairs that are useful in fuzzer taming.
+
 \subsection{Lossy Techniques}

### 2019-01-23 journal/example.tex
ok

@@
 We can identify these mutants, which will form the basis of our distance metric and our fault localization, by the line number of the change, since there is only one mutant per line in the set.
-Each test can then be associated with a vector of 5 elements:  {\tt (repaired by line 3, repaired by line 6, repaired by line 7, repaired by line 10, repaired by line 13)}.
+Each test can then be associated with a bit-vector of 5 elements:  {\tt (repaired by line 3, repaired by line 6, repaired by line 7, repaired by line 10, repaired by line 13)}.
+The vectors for our tests, then, are:
+
+\begin{code}
+Test 1: 1 1 0 0 0
+Test 2: 1 0 1 1 0
+Test 3: 1 0 1 0 1
+\end{code}
+
+Test 1 thus has a distance of 0.75 from Test 2 and Test 3.
+Test 2 has a distance of only 0.5 from Test 3.
+If the Furthest-Point-First algorithm starts from Test 1 it will arbitrarily select either Test 2 or 3 next, then the other test.
+If it starts with Test 2 or 3, it will select Test 1 next, then the other of the two closer tests.
+Of course, taming a set of three failing tests is not difficult.
+However, if we imagine a much larger set of failing tests, which contain irrelevant features that cannot be removed by delta-debugging without causing the programs to become syntactically invalid, so long as the basic structure of the cause for failure remains the same, then the FPF algorithm will select tests by the distance between vectors and still essentially produce Test 2, then 1, then 3, for example.

### 2019-01-23 journal/abstract.tex
Example is ready to go

--- 2017-05-05 abstract.tex
@@
 The metric is based on a causal concept of distance where executions are similar to the degree that changes in the program itself, introduced by mutation, cause similar changes in the correctness of the executions.
-Specifically, if two failing test cases (for the original compiler) become successful due to the same mutant, they are more likely to be due to the same fault.
+Specifically, if two failing test cases for a compiler can be made to pass by applying the same mutant, those two tests are more likely to be due to the same fault.
 We evaluate our metric using more than 50 faults and 2,800 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.
+We additionally show that this approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs.

### 2019-01-23 journal/example.tex
Example is ready to go

--- 2019-01-22 journal/example.tex
@@
 \begin{code}
-Test 1:  (1, 2)
-Test 2:  (3)
-Test 3:  (5, 6)
+T1: (1, 2)
+T2: (3)
+T3: (5, 6)
 \end{code}
@@
 
-Assume that all of our test cases run the compiler with {\tt optimizeLevel} of 2, so as to potentially apply all optimizations.
-A popular way to fuzz a compiler is to compare program behavior when code is compiled with all optimizations with code compiled with no optimizations, so this reflects real-world compiler testing well.
-A set of delta-debugged failing tests, then, might be those shown in Figure \ref{fig:toybugs}.
+Assume that all of our test cases run the compiler with {\tt
+  optimizeLevel} of 2, so as to potentially apply all optimizations.
+A popular way to fuzz a compiler is to compare program behavior when
+code is compiled with all optimizations with program behavior when
+code is compiled with no optimizations, so this reflects real-world compiler testing.  A set of delta-debugged failing tests, then, might be those shown in Figure \ref{fig:toybugs}.
 
-Further, let us only consider two kinds of mutants:  condition-negation and statement deletion mutants.
-Obviously, negating the condition on line 3 will make all of our failing test cases pass, and negating the condition on line 7 will make test cases 2 and 3 pass, but not test case 1.
-Negating the conditions on lines 4-5, 8-9, and 11-12 will repair our test cases, but is also likely to make some passing test case fail, so these mutants will not be used in our analysis.
+Further, let us only consider two kinds of mutants:
+condition-negation and statement-deletion mutants.  Obviously,
+negating the condition on line 3 will make all of our failing test
+cases pass, and negating the condition on line 7 will make test cases
+2 and 3 pass, but not test case 1.  Negating the conditions on lines
+4-5, 8-9, and 11-12 will repair our test cases, but these changes are
+also likely to make some passing test case fail (they represent
+preconditions for compiler optimizations), so these mutants will not
+be used in our analysis.
 Finally, the only statements that can be deleted without breaking the code are deletions of the {\tt corrupt} calls on lines 6, 10, and 13.
 We can identify these mutants, which will form the basis of our distance metric and our fault localization, by the line number of the change, since there is only one mutant per line in the set.
-Each test can then be associated with a vector of 5 elements:  {\tt (repaired by line 3, repaired by line 6, repaired by line 7, repaired by line 10, repaired by line 13)}.
+Each test can then be associated with a bitvector of 5 elements:  {\tt (repaired by line 3, repaired by line 6, repaired by line 7, repaired by line 10, repaired by line 13)}.
+The vectors for our tests, then, are:
+
+\begin{code}
+T1: 11000
+T2: 10110
+T3: 10101
+\end{code}
+
+T1 thus has a distance of 0.75 from T2 and T3.  T2 has
+a distance of only 0.5 from T3, however.
+
+\subsection{Fuzzer Taming for the Toy Compiler}
+
+If the Furthest-Point-First
+algorithm starts from T1 it will arbitrarily select either T2
+or T3 next, then the other test.  If it starts with T2 or T3, it
+will select T1 next, then the other of the two closer tests.  Of
+course, taming a set of three failing tests is not difficult!
+However, if we imagine a much larger set of failing tests, which
+contain irrelevant features that cannot be removed by delta-debugging
+without causing the programs to become syntactically invalid, so long
+as the basic structure of the cause for failure remains the same, the
+the FPF algorithm will still select tests with the same bitvectors as
+T1, T2, and T3.  Concretely, if we add these tests (shown with associated bitvectors):
+
+\begin{code}
+T4: (1, 2, 7)    11000
+T5: (3, 6)       10110
+T6: (1, 5, 6)    10101
+T7: (1, 2, 5, 6) 10000
+\end{code}
+
+\noindent then FPF from T1 will proceed to a maximally distant
+test, such as T5, at distance 0.75.  The next ranked test could not have
+been T4 (distance 0.0) or T7 (distance 0.5), but could have
+been T2, T3, or T6.  After T1 and T5, the next test in the ranking will be the test whose nearest
+neighbor among already chosen tests is furthest away.  The distances
+to the two tests already chosen are:
+
+{\scriptsize
+\begin{code}
+                         11000  10110
+T2: (3)          10110   0.75   0.00
+T3: (5, 6)       10101   0.75   0.50
+T4: (1, 2, 7)    11000   0.00   0.75
+T6: (1, 5, 6)    10101   0.75   0.50
+T7: (1, 2, 5, 6) 10000   0.50   0.67
+\end{code}
+}
+
+\noindent The FPF algorithm could choose either T3, T6, or T7 all of
+which have a nearest neighbor at distance 0.5.  Let
+us assume it chooses T6.  At this point, the three distinct bugs
+are all ranked.  However, there is one interesting pattern remaining.
+The distances to tests are now:
+
+{\scriptsize
+\begin{code}
+                         11000  10110  10101
+T2: (3)          10110   0.75   0.00   0.50
+T3: (5, 6)       10101   0.75   0.50   0.00
+T4: (1, 2, 7)    11000   0.00   0.75   0.75
+T7: (1, 2, 5, 6) 10000   0.50   0.67   0.67
+\end{code}
+}
+
+\noindent The algorithm is forced to choose T7, the only remaining
+test that does not have the same bitvector as another test.  T7 is
+worth ranking, as it shows that a program can be written that will
+result in wrong code unless both bugs it triggers are fixed.
+
+\subsection{Fault Localization for the Toy Compiler}
+
+Now let us consider the {\bf Repair} fault localization algorithm
+applied to this example.  If we are trying to fix the problem
+exhibited by T2, what will the system suggest to us as the
+location of the fault?  Recall that 1) in {\bf Repair}, a localization
+is the location (and content) of a repairing mutant.  T2 is repaired
+by negating the condition on line 3, negating the condition on line 7,
+or deleting the statement on line 10.  Which of these is best?  {\bf
+  Repair} distinguishes between repairing mutants by ordering them by
+their distance to the most-distant (by bitvector) test also repaired
+by that mutant.  The only other tests repaired by deletion of line 10
+are the other instances of the same bug, all at distance 0.  Of the
+other two repairs, negating the condition on line 3 also repairs T1,
+at distance 0.75, while negating the condition on line 2 repairs
+nothing farther away than T3 at 0.5.
+Therefore, the localization for T2 is (in rank order):
+
+{\scriptsize {\tt 1: DELETE  10                corrupt(internalRepresentation);}}
+
+{\scriptsize {\tt 2: NEGATE  07       if optimizeLevel > 1: \{}}
+
+{\scriptsize {\tt 3: NEGATE  03    if optimizeLevel > 0: \{}}
+
+\subsection{The Onion Ring}
+
+In this simple toy example, the onion ring will only have a few
+layers, one for each possible distance from the test at the center.
+For example, if we take T2 as the center, then the layers will be:
+
+\begin{code}
+ 0.00:  T2 T5
+ 0.50:  T3 T6
+ 0.67:  T7
+ 0.75:  T1 T4
+\end{code}

### 2019-01-23 journal/introduction.tex
Example is ready to go

@@
 Repairs can also provide  information for understanding and debugging discovered faults.
-\comment{The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.
-Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).}
+The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.
+Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).
 
@@
 This leads to the general idea of comparing any two executions by their response to mutants, since a mutant can cause a passing or failing test to violate more/different properties.   
-However, for practical purposes the most interesting implication is that if two failing test cases can be made to differ as to whether they fail by a mutant, this provides evidence that they fail due to different faults.
-
-\comment{
-%Where there are multiple, independent, correctness properties, this even allows us to compare failing and successful executions, since a mutant can cause a failure to violate a different property.
-{\bf Hypothesis 2} at first may suggest that this idea subsumes past notions of similarity, such as spectra of code coverage, since if two executions do not cover a statement, they both cannot alter correctness if that statement is altered.
+This generalization to any two executions may at first glance suggest that our approach subsumes past notions of similarity, such as spectra of code coverage, since if two executions do not cover a statement, they both cannot alter correctness if that statement is altered.
 The metrics are in fact simply incomparable:  for example, differences in coverage that never have an effect on correctness cannot contribute to similarity in our approach; with weak correctness properties most executions are very similar.
 This is similar to the idea of \emph{checked coverage}:  code that cannot impact correctness is not interesting \cite{CheckCov}.
-Hypothesis 2 also suggests an additional hypothesis related to hypothesis 1.
-This is {\bf Hypothesis 1a:} If two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.
-}
+However, for practical purposes the most interesting implication of our proposal is that if two failing test cases can be made to differ as to whether they fail by applying a program mutant, this provides evidence that they fail due to different faults.
 

### 2019-01-23 journal/introduction.tex
submittable journal version

@@
 This paper extends our 2018 ISSRE paper \cite{ISSRE18Mutants} introducing our approach.
-That work, and this, suggested a novel way of measuring distance between program executions, but more importantly provided some practical proposals for using such a metric in compiler fuzzer taming and fault localization.
+That paper introduced a novel way of measuring distance between program executions, but more importantly provided some practical proposals for using such a metric in compiler fuzzer taming and fault localization.
 Our experimental results show that, for two realistic complex compilers, Mozilla's SpiderMonkey 1.6 JIT for JavaScript and GCC 4.3.0, using mutant response metrics improves FPF-based fuzzer taming over best previous results \cite{PLDI13}, in multiple ways.
-Our fault localization approach also compares favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual,Metallaxis}, albeit with a more limited applicability (but also a lower cost).
+Our fault localization approach also compares favorably to state-of-the-art fault localization techniques \cite{MUSE,multilingual,Metallaxis} for non-compiler programs, albeit with a more limited applicability (but also a lower cost).
 We extend the earlier paper by:
@@
 \begin{itemize}
-\item generally expanding discussion of our approach, its' conceptual basis, and related work,
-\item providing an extended example (in Section \ref{sec:example} to help readers better understand the metric, FPF-based fuzzer taming, and our fault localization approach,
+\item generally expanding discussion of our approach, including the conceptual basis for using mutants to measure similarity of executions, and related work,
+\item providing an extended example (in Section \ref{sec:example}) to help readers better understand the metric, FPF-based fuzzer taming, and our fault localization approach,
 \item introducing and evaluating another compiler fault localization approach (in Section \ref{sec:covloc}), based on code coverage changes,

### 2019-01-25 journal/experiments.tex
in

@@
 This is how SQLite applies mutation testing:  \url{https://www.sqlite.org/testing.html#mutation_testing}.
-Object level mutants are obviously not useful (or at least, certainly not \emph{as} useful) for fault localization as source-level mutants, but they mutating branches alone probably captures many compiler repairs that are useful in fuzzer taming.
+Object level mutants are obviously not useful (or at least, certainly not \emph{as} useful) for fault localization as source-level mutants, but mutating branches alone probably captures many compiler repairs that are useful in fuzzer taming.
 
@@
 
+\subsubsection{Random Sampling}
+
 Mutation cost reduction techniques classified as \emph{do fewer} on the other hand involve a loss of signal, not just more efficient production of the same signal.
+One of the most popular and simple approaches to reducing the cost of mutation testing is \emph{random sampling} \cite{budd1980mutation,acree1980mutation,RahulISSRE,MutRand}.
  Figures \ref{samplejs} and \ref{samplegcc} show, respectively, the results of random sampling of mutants for SpiderMonkey and GCC, by comparing the APFD obtained on a set of 30 random samples of the chosen percentage of mutants with the APFD for the full mutant set.
@@
 
+\subsubsection{Operator-Based Selection}
+
+A deterministic alternative to random sampling, long used in the mutation testing community, is operator-based selection, where only some mutation operators are applied \cite{wong1995reducing,untch2009onreduced}.
+Using \emph{only statement deletion and condition negation mutants}, as in our toy example, on the grounds that we expect most compiler ``repairs'' to simply avoid an optimization, not actually fix a simple defect such as an off-by-one error in a loop, is an intuitively appealing and trivial to implement approach.
+The idea is clearly related to the object-level branch-mutation proposed above, conceptually, but operating at the (more understandable to humans, and thus useful in localization) source level.
+The resulting cost reduction is significant:  67.8\% of mutants can be ignored for SpiderMonkey, and 73\% for GCC.
+More importantly, the APFD actually \emph{improves} with this change.
+For SpiderMonkey, the APFD improves by almost 3\%, and for GCC, it improves by 0.86\%.
+For SpiderMonkey, the early curve (first ten items) is improved, and for GCC it is very slightly worse.
+Most importantly, in both cases, two additional faults are revealed in the first 50 examined tests.
+
 \subsection{Threats to Validity}

### 2019-01-25 journal/experiments.tex
ok

@@
 Most importantly, in both cases, two additional faults are revealed in the first 50 examined tests.
+Operator selection can make a difference for fault localization, but with hot mutants, it is easy, if localization using only negations and deletions is not helpful, run all mutants on the test to be localized, then run only the small set of repairing mutants on other tests to compute the same localization as with the full set of operators, since most tests are repaired by only a few mutants.
 

### 2019-01-25 journal/experiments.tex
operator based

--- 2019-01-25 journal/experiments.tex
@@
 Most importantly, in both cases, two additional faults are revealed in the first 50 examined tests.
+Operator selection can make a (negative) difference for fault localization, but with hot mutants, it is easy, if localization using only negations and deletions is not helpful, run all mutants on the test to be localized, then run only the small set of repairing mutants on other tests to compute the same localization as with the full set of operators, since most tests are repaired by only a few mutants.
+This should usually not be required, however:  we noted that in cases where a mutant involved in a localization did use a operator replacement it was usually equivalent in behavior to a condition negation.
+Constant replacement repairs were simply very rare and not involved in any of the useful compiler localizations.
 

### 2019-01-25 journal/experiments.tex
change operator text

--- 2019-01-25 journal/experiments.tex
@@
 Most importantly, in both cases, two additional faults are revealed in the first 50 examined tests.
+Operator selection can, in principle (and in one case in our data set, in practice) reduce the effectiveness of fault localization.
+However, with hot mutants, it is easy, if localization using only negations and deletions is not helpful, to use the same method as proposed above for random selection to compute a more precise localization for a test.
+This should usually not be required, however:  we noted that in most cases where a mutant involved in a localization did use a operator replacement, the mutant was usually equivalent in behavior to a condition negation.
+Constant replacement repairs were simply very rare and not involved in any of the useful compiler localizations.
+The problem observed with random sampling, that the metric of distances to other tests may be inaccurate, is less of concern with operator selection:  as far as we can tell, based on FPF discovery curves and APFD, distances are \emph{more} accurate using only the mutants we propose.
 

### 2019-01-26 journal/conclusions.tex
journal start
just change format, proofread, and ship

--- 2017-05-05 conclusions.tex
@@
 
+More practically, we would like to tightly integrate our approach with
+automated test generation tools, such as TSTL \cite{tstlsttt,NFM15},
+DeepState \cite{DeepState,DeepStateTutorial}, Echidna \cite{Echidna}, and
+Manticore \cite{Manticore}.  Automated test generation tools tend to produce a
+large number of passing and failing tests, and are often used in
+``overnight'' runs where adding a step to use mutants to triage
+failing tests and provide localization suggestions for the highly
+ranked tests may not even impose a noticable overhead on current usage
+practices.  The latter two tools (Echidna and Manticore), which target smart
+contracts for the Ethereum blockchain \cite{wood2014ethereum}, are
+particularly promising opportunities, because the gas-limited computation bounds of the Ethereum
+Virtual Machine essentially guarantee that smart contract code is
+small enough to make mutation analysis cheap.  The availability of a
+multi-language mutation analysis tool \cite{RegExpMut} supporting
+Python, C/C++, and Solidity (the language used for most smart
+contracts) means that a common infrastructure base can be shared by
+all of these tool integrations.
+
 We also plan to explore other applications of mutant-based metrics.
@@
 history \cite{KlausRajeev,scriptstospecs}.
+
+{\scriptsize {\bf Acknowledgements:} The authors would like to thank the authors of
+the MUSEUM \cite{multilingual} and Metallaxis-FL vs. MUSE 
+\cite{Papadakis} studies for access to their data sets.  In
+particular, we thank Shin Hong and Thierry Chekam for prompt
+assistance beyond the call of duty.}

### 2019-01-26 journal/experiments.tex
just change format, proofread, and ship

@@
 
+We tried various ways of improving the effectiveness of random sampling by forcing the inclusion of certain mutants that were deemed more likely to be useful.
+For example, it is very important to include mutants that repair hard-to-repair faults represented by a small number of tests.
+Finding faults represented by few tests is, in some sense, what users want from a fuzzer taming algorithm.
+We cannot, without evaluating them over tests, predict which mutants will repair only a small number of tests (and thus, likely, faults with only a few representatives).
+However, making sure to include mutants that are only \emph{covered} by a small number of tests lets us ensure we cover mutants that, if they repair any failing tests, only repair a small number of failing tests.
+However, due to the likelyhood that such mutants repair no tests at all, including mutants covering at most $k$ tests (for small $k$) turned out to be either actually harmful (statistically significantly worse APFD) or statistically indistinguishable from pure random selection.
+We were unable to find any mixes of random and biased or chosen sampling that performed better than pure random sampling.
+Is this the best we can do?
+
 \subsubsection{Operator-Based Selection}
@@
 The resulting cost reduction is significant:  67.8\% of mutants can be ignored for SpiderMonkey, and 73\% for GCC.
+
 More importantly, the APFD actually \emph{improves} with this change.
-For SpiderMonkey, the APFD improves by almost 3\%, and for GCC, it improves by 0.86\%.
+For SpiderMonkey, the APFD improves by almost 3\%, and for GCC it improves by 0.86\%.
 For SpiderMonkey, the early curve (first ten items) is improved, and for GCC it is very slightly worse.
-Most importantly, in both cases, two additional faults are revealed in the first 50 examined tests.
+Most importantly, in both cases, \emph{two additional faults} are revealed in the first 50 examined tests.
+We speculate that the accidental nature of many repairs produced by operator and constant replacements may introduce harmful noise into the distance metric, and thus the FPF-based fuzzer taming process.
+Changes that simply sidestep faulty functionality are, perhaps, more likely to be meaningful \cite{achour}.
+
 Operator selection can, in principle (and in one case in our data set, in practice) reduce the effectiveness of fault localization.
@@
 Constant replacement repairs were simply very rare and not involved in any of the useful compiler localizations.
-The problem observed with random sampling, that the metric of distances to other tests may be inaccurate, is less of concern with operator selection:  as far as we can tell, based on FPF discovery curves and APFD, distances are \emph{more} accurate using only the mutants we propose.
+The problem observed with random sampling, that the metric of distances to other tests may be inaccurate, is of much less concern with operator selection:  as far as we can tell, based on FPF discovery curves and APFD, distances are \emph{more} accurate using only the mutants we propose.
+
+While using only some mutation operators \emph{may} degrade fault localization performance for non-compiler faults, we propose it as the default way to use our approaches in compiler fuzzer taming  and debugging scenarios.
+Furthermore, we speculate, based on the results of Qi et al.
+\cite{achour} showing that most actual correct fixes in program repair are functionality deletions, that limiting mutants used to condition negations and statement deletions (or even adding an operator that converts conditionals to a {\tt False} constant, as is provided by some tools \cite{RegExpMut}), might be the best practice for complex programs and ``deep'' (hard to debug) faults in general.
+It seems unlikely that mutation operators will produce many fixes that do not simply avoid broken functionality (vs.
+actually fix it), for the kinds of faults that are actually hard for programmers to locate, understand, and correct.
 

### 2019-01-26 journal/stvrdoc.tex
stvr stuff

% stvrdoc.tex V2.0, 13 May 2010

\documentclass[times]{stvrauth}

\usepackage{moreverb}

\usepackage[dvips,colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2010}

\begin{document}

\runningheads{A.~N.~Other}{A demonstration of the \journalabb\
class file}

\title{A demonstration of the \LaTeXe\ class file for
\itshape{\journalnamelc}\footnotemark[2]}

\author{A.~N.~Other\corrauth}

\address{John Wiley \& Sons, Ltd, The Atrium, Southern Gate, Chichester,
West Sussex, PO19~8SQ, UK}

\corraddr{Journals Production Department, John Wiley \& Sons, Ltd,
The Atrium, Southern Gate, Chichester, West Sussex, PO19~8SQ, UK.}

\begin{abstract}
This paper describes the use of the \LaTeXe\
\textsf{\journalclass} class file for setting papers for
\emph{\journalnamelc}.
\end{abstract}

\keywords{class file; \LaTeXe; \emph{\journalabb}}

\maketitle

\footnotetext[2]{Please ensure that you use the most up to date
class file,
available from the STVR Home Page at\\
\href{http://www3.interscience.wiley.com/journal/13635/home}{\texttt{http://www3.interscience.wiley.com/journal/13635/home}}}

\vspace{-6pt}

\section{Introduction}
\vspace{-2pt}
Many authors submitting to research journals use \LaTeXe\ to
prepare their papers. This paper describes the
\textsf{\journalclass} class file which can be used to convert
articles produced with other \LaTeXe\ class files into the correct
form for publication in \emph{\journalnamelc}.

The \textsf{\journalclass} class file preserves much of the
standard \LaTeXe\ interface so that any document which was
produced using the standard \LaTeXe\ \textsf{article} style can
easily be converted to work with the \textsf{\journalclassshort}
style. However, the width of text and typesize will vary from that
of \textsf{article.cls}; therefore, \emph{line breaks will change}
and it is likely that displayed mathematics and tabular material
will need re-setting.

In the following sections we describe how to lay out your code to
use \textsf{\journalclass} to reproduce the typographical look of
\emph{\journalnamelc}. However, this paper is not a guide to
using \LaTeXe\ and we would refer you to any of the many books
available (see, for example, \cite{R1,R2,R3}).

\vspace{-6pt}

\section{The Three Golden Rules}
\vspace{-2pt}

Before we proceed, we would like to stress \emph{three golden
rules} that need to be followed to enable the most efficient use
of your code at the typesetting stage:
\begin{enumerate}
\item[(i)] keep your own macros to an absolute minimum;

\item[(ii)] as \TeX\ is designed to make sensible spacing
decisions by itself, do \emph{not} use explicit horizontal or
vertical spacing commands, except in a few accepted (mostly
mathematical) situations, such as \verb"\," before a
differential~d, or \verb"\quad" to separate an equation from its
qualifier;

\item[(iii)] follow the \emph{\journalnamelc} reference style.
\end{enumerate}

\pagebreak

\section{Getting Started} The \textsf{\journalclassshort} class file should run
on any standard \LaTeXe\ installation. If any of the fonts, style
files or packages it requires are missing from your installation,
they can be found on the \emph{\TeX\ Collection} DVDs or from
CTAN.

\emph{\journalnamelc} is published using Times fonts and this is
achieved by using the \verb"times"
option as\\
\verb"\documentclass[times]{stvrauth}".

\noindent If for any reason you have a problem using Times you can
easily resort to Computer Modern fonts by removing the
\verb"times" option.

\begin{figure}
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}
\documentclass[times]{stvrauth}
%\documentclass[times,doublespace]{stvrauth}%For paper submission

\begin{document}

\runningheads{<Initials and Surnames>}{<Short title>}

\title{<Initial cap, lower case>}

\author{<An Author\affil{1},
Someone Else\affil{2}\corrauth\ and Perhaps Another\affil{1}>}

\address{<\affilnum{1}First author's address
(in this example it is the same as the third author)\break
\affilnum{2}Second author's address>}

\corraddr{<Corresponding author's address (the second author in
this example)>. E-mail: <corresponding author's email address>}

%\cgs{<Contract/grant sponsor name (no number)>}
%\cgsn{<Contract/grant sponsor name>}{<number>}

\begin{abstract}
<Text>
\end{abstract}

\keywords{<List keywords>}

\maketitle

\section{Introduction}
.
.
.
\end{boxedverbatim}
\end{center}
%\vspace{-12pt}
\caption{Example header text.\label{F1}}
\end{figure}

\section{The Article Header Information}
The heading for any file using \textsf{\journalclass} is shown in
Figure~\ref{F1}.

\subsection{Remarks}
\begin{enumerate}
\item[(i)] In \verb"\runningheads" use `\emph{et~al.}' if there
are three or more authors.

\item[(ii)] Note the use of \verb"\affil" and \verb"\affilnum" to
link names and addresses. The author for correspondence is marked
by \verb"\corrauth" and \verb"\corraddr" is used to give that
author's address, which will be printed as a footnote, prefaced by
`Correspondence to:'.

\item[(iii)] For submitting a double-spaced manuscript, add
\verb"doublespace" as an option to the documentclass line.

\item[(iv)] Use \verb"\cgs" for giving details of financial
sponsors; alternatively use \verb"\cgsn" if the grant number is
also to be included. These details will be printed as a footnote,
with `Contract/grant sponsor:' and `contract/grant number:'
inserted in the appropriate places.

\item[(v)] The abstract should be capable of standing by itself,
in the absence of the body of the article and of the bibliography.
Therefore, it must not contain any reference citations.

\item[(vi)] Keywords are separated by semicolons.
\end{enumerate}

\begin{figure}
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}
\begin{table}
\caption{<Table caption>}
\centering
\tabsize
\begin{tabular}{<table alignment>}
\toprule
<column headings>\\
\midrule
<table entries
(separated by & as usual)>\\
<table entries>\\
.
.
.\\
\bottomrule
\end{tabular}
\end{table}
\end{boxedverbatim}
\end{center}
\vspace{-6pt}
\caption{Example table layout.\label{F2}}
\vspace{-6pt}
\end{figure}

\section{The Body of the Article}

\subsection{Mathematics} \textsf{\journalclass} makes the full
functionality of \AmS\/\TeX\ available. We encourage the use of
the \verb"align", \verb"gather" and \verb"multline" environments
for displayed mathematics. \textsf{amsthm} is used for setting
theorem-like and proof environments. The usual \verb"\newtheorem"
command needs to be used to set up the environments for your
particular document.

\subsection{Figures and Tables} \textsf{\journalclass} includes the
\textsf{graphicx} package for handling figures.

Figures are called in as follows:
\begin{verbatim}
\begin{figure}
\centering
\includegraphics{<figure name>}
\caption{<Figure caption>}
\end{figure}
\end{verbatim}

For further details on how to size figures, etc., with the
\textsf{graphicx} package see, for example, \cite{R1}
or \cite{R3}. If figures are available in an
acceptable format (for example, .eps, .ps) they will be used but a
printed version should always be provided. \medbreak

The standard coding for a table is shown in Figure~\ref{F2}.

\subsection{Cross-referencing}
The use of the \LaTeX\ cross-reference system
for figures, tables, equations, etc., is encouraged
(using \verb"\ref{<name>}" and \verb"\label{<name>}").

\subsection{Acknowledgements} An Acknowledgements section is started with \verb"\ack" or
\verb"\acks" for \textit{Acknowledgement} or
\textit{Acknowledgements}, respectively. It must be placed just
before the References.

\subsection{Bibliography}
The normal commands for producing the reference list are:
\begin{verbatim}
\begin{thebibliography}{99}
\bibitem{<x-ref label>}
         <Reference details>
.
.
.
\end{thebibliography}
\end{verbatim}
where \verb"\bibitem{x-ref label}"
corresponds to \verb"\cite{x-ref label}" in the body of the article
and \verb"{99}" is the widest such number expected and determines
the width of the number column in the reference list.

Please note that the file \textsf{wileyj.bst} is available from
the same download page for those authors using \BibTeX.

\subsection{Double Spacing}
If you need to double space your document for submission please
use the \verb+doublespace+ option as shown in the sample layout in
Figure~\ref{F1}.

\section{Support for \textsf{\journalclass}}
We offer on-line support to participating authors. Please contact
us via e-mail at\\
\href{mailto:stvrauth-cls@wiley.co.uk}{\texttt{stvrauth-cls@wiley.co.uk}}.

We would welcome any feedback, positive or otherwise, on your
experiences of using \textsf{\journalclass}.

\section{Copyright Statement}
Please  be  aware that the use of  this \LaTeXe\ class file is
governed by the following conditions.

\subsection{Copyright}
Copyright \copyright\ \volumeyear\ John Wiley \& Sons, Ltd, The
Atrium, Southern Gate, Chichester, West Sussex, PO19~8SQ, UK. All
rights reserved.

\subsection{Rules of Use}
This class file is made available for use by authors who wish to
prepare an article for publication in \emph{\journalnamelc}
published by John Wiley \& Sons, Ltd. The user may not exploit any
part of the class file commercially.

This class file is provided on an \emph{as is}  basis, without
warranties of any kind, either express or implied, including but
not limited to warranties of title, or implied  warranties of
merchantablility or fitness for a particular purpose. There will
be no duty on the author[s] of the software or  John Wiley \&
Sons, Ltd to correct any errors or defects in the software. Any
statutory  rights you may have remain unaffected by your
acceptance of these rules of use.

\ack This class file was developed by Sunrise Setting Ltd,
Torquay, Devon, UK. Website:\\
\href{http://www.sunrise-setting.co.uk}{\texttt{www.sunrise-setting.co.uk}}

\begin{thebibliography}{9}

\bibitem{R1} Kopka~H, Daly~PW. 2003. \emph{A Guide to \LaTeX} (4th~edn).
Addison-Wesley.

\bibitem{R2} Lamport~L. 1994. \emph{\LaTeX: a Document Preparation System} (2nd~edn).
Addison-Wesley.

\bibitem{R3} Mittelbach~F, Goossens~M. 2004. \emph{The \LaTeX\ Companion}
(2nd~edn). Addison-Wesley.
\end{thebibliography}
\end{document}

### 2019-01-26 journal/conclusions.tex
right format

@@
 \label{conc}
-\comment{
 When are two failing test cases due to the same fault?  This paper
@@
 has long been tied to a notion of similarity of
-executions \cite{NearNeighbor,GroceError}. } By using program changes
+executions \cite{NearNeighbor,GroceError}.  By using program changes
 (produced by mutants) as
@@
 compiler debugging can consistently be made less onerous through automated
-assistance.  \comment{One idea is to combine the metric approach and fall
-the ideas
-in MUSE and MUSEUM, or other mutant-based localizations such as
-Metallaxis \cite{Metallaxis} in some way, to exploit passing tests and perhaps
-additional coverage information.  We would also be very interested to
-compare our localizations with the Metallaxis
-\cite{Metallaxis,Papadakis} approach.}
+assistance.
 

### 2019-01-26 journal/experiments.tex
right format

@@
 Faults that produce a crash are also likely to be much easier to debug than semantic problems such as all of the GCC wrong code bugs and most of the JavaScript faults --- e.g., following a dynamic slice might well suffice.
-
-\comment{
-We applied our metric to two of the datasets used in the previous study of compiler fuzzer taming \cite{PLDI13}: faults and tests cases for SpiderMonkey 1.6, Mozilla's JavaScript engine, with tests generated by {\tt jsfunfuzz} \cite{jsfunfuzz}, and \emph{wrong code} faults and test cases for GCC 4.3.0, with tests generated by Csmith \cite{csmith}.
-We did not use the crash faults from that data set, as these are in practice fairly easy to identify by hand (and perfectly identified by previous metrics), and also typically easy to debug.
-}
 
@@
 They seem likely to be less useful as explanations of the causes of a failure, since they do not impact the program semantics in a way that is only known to be beneficial, and could potentially even mislead a user if used as explanations.
-\comment{Still, it is most useful to compare MUSEUM/MUSE and our approach on the cases where all methods produce a localization.}
 
@@
 Blind, unaided pointers to portions of compiler code without more semantic information about why the statement is suspicious seem unlikely to work well, based on our experience with complex compiler faults, and the results of Parnin and Orso \cite{AutoHelp}.
-
-\comment{
-\begin{table}
-\centering
-{\scriptsize
-\begin{tabular}{|c||c|c|c|}
-\hline
-%& \multicolumn{3}{|c|}{Localization Rank} \\
-%\hline
-Fault & Repair & MUSE & MUSEUM \\
-\hline
-880 & - & - & - \\
-95 & - & 173 & -\\
-1294 & 1 & 22 & 4 \\
-60 & - & - & -\\
-1543 & 19 & 118 & 26\\
-1172 & 28 & 87 & 23\\
-1561 & - & - & - \\
-115 & 3 & 133 & 5 \\
-\hline
-\end{tabular}
-}
-\caption{Fault localization results for SpiderMonkey}
-%\vspace{-0.2in}
-\label{jstable}
-\end{table}
-}
 
@@
 \subsubsection{Fault Localization}
-
-\comment{
-\begin{table}
-\centering
-{\scriptsize
-\begin{tabular}{|c||c||c|c|}
-\hline
-%& & \multicolumn{4}{|c|}{Localization Rank} \\
-%\hline
-Fault & Repair & MUSE & MUSEUM \\
-\hline
-139094 & - & 157 & -\\
-133004 & 1 & 99 & 14\\
-158782 & - & 146 & -\\
-156496 &  - & - & -\\
-143677 & - & 109 & -\\
-158555 & - & - & -\\
-134321 & - & - & -\\
-133940 & 4 & 158 & 18\\
-141195 & - & - & -\\
-156795 & - & - & -\\
-139709 & 10 & 167 & 6\\
-155698 & 1 & 172 & 195\\
-138646 & - & - & -\\
-140795 & 34 & 170 & 125\\
-136501 &  5 & 159 & 31\\
-134322 & - & - & -\\
-\hline
-\end{tabular}
-}
-\caption{Fault localization results for GCC}
-\label{gcctable}
-\end{table}
-}
 
@@
 
-%\comment{
  Why does {\bf Repair} perform best in the \emph{compiler} setting in particular?
@@
 Fixing deep optimizing compiler bugs is a very specialized skill, in contrast to general debugging skills.
-
-\comment{ For the compiler faults, if any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.
-MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.
-Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this is unlikely due to the nature of the repairs.
-}
-
-\comment{ In one sense, {\bf Coverage} is the most intriguing method presented here.
-While it did not perform nearly as well as {\bf Repair}, it did manage, in many cases, to rank \emph{some} faulty line highly relative to the extremely large number of statements executed in each failing GCC run (on the order of 50K-100K+ statements).
-In future work, we plan to experiment with different methods for weighting frequency of coverage, perhaps using methods from machine learning, or of pruning changed coverage that is not likely to be relevant.
-For example, perhaps the {\bf Repair} approach can be applied, and coverage changes present in very distant failures can be removed, though the argument is less clear, since a coverage change isn't directly associated with the onion-ring structure, and does not directly repair a set of failures.
-There is much to be done if automated debugging is ever to make the lives of compiler developers easier in most cases \cite{AutoHelp}.
-}
 

### 2019-01-26 journal/localizations.tex
right format

@@
 information about the likely location of a fault in a program's source
-code.  \comment{Error explanation \cite{GroceError} also provides information
+code.  Error explanation \cite{GroceError} also provides information
 about the causal structure of a fault --- how it can be avoided, what
-triggers it --- a kind of ``fault story.''}  Parnin and Orso recently
+triggers it --- a kind of ``fault story.''  Parnin and Orso recently
 suggested that fault localization methods are not actually providing a

### 2019-01-26 journal/metrics.tex
right format

@@
 are due to the same fault, even if not all repairs match.
-\comment{
-\footnote{In Section \ref{conc} we discuss uses of $d$
-requiring comparison of both failing and passing test cases:  in this 
-case, we would use three-valued vectors, where 1 indicates a 
-failing property succeeds and -1 that a succeeding property 
-fails.  For such an application, we could remove the requirement that
-a repair not be a killed mutant, since the passing test vectors would encode
-that information.}}
 

### 2019-01-26 journal/example.tex
done

@@
  15    \}
- 16    /* Will produce wrong code if structure has been corrupted */
+ 16    /* Will produce wrong code if corrupt called */
  17    generateCode(internalRepresentation);

### 2019-01-27 journal/experiments.tex
done
data location, put threats in own section

@@
 
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.5\textwidth}
+\begin{figure}
     \centering
@@
         \label{jscurves}
-    \end{subfigure}%
+ \end{figure}%
     ~ 
-    \begin{subfigure}[t]{0.5\textwidth}
+    \begin{figure}
         \centering
@@
         \label{gcccurves}
-    \end{subfigure}
     \caption{Discovery curves compared to best curves from Chen et. al PLDI 2013 data (FPF benchmark)}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsection{Threats to Validity}
+\section{Threats to Validity}
 
@@
 The comparison with other fault localization methods is also limited:  our compiler experiments limit the mutation budget in ways that may weaken MUSE and MUSEUM effectiveness, and the results using fault and mutant data sets from other studies provide few faults, no multi-fault problems, and very few multi-failure problems, limiting {\bf Repair} effectiveness.
-It is also possible that faults in the framework used for experiments introduced errors into the data.
-In order to check for these problems, we will make the raw data for a re-analysis available on request.
+It is also possible that faults in the framework used for experiments introduced errors into our results.
+In order to make it possible to check for these problems, we have made the raw data for a re-analysis available: \url{https://github.com/agroce/mutants16/tree/master/tests}.
+The data required for analysis of the non-compiler faults is available from the authors of the respective studies (see the acknowledgements at the end of the paper).

### 2019-02-03 journal/conclusions.tex
add one more tstl ref

@@
 More practically, we would like to tightly integrate our approach with
-automated test generation tools, such as TSTL \cite{tstlsttt,NFM15},
+automated test generation tools, such as TSTL \cite{tstlsttt,NFM15,ISSTA15},
 DeepState \cite{DeepState,DeepStateTutorial}, Echidna \cite{Echidna}, and

### 2019-02-04 journal/abstract.tex
more small fixes

@@
 We evaluate our metric using more than 50 faults and 2,800 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.
-We additionally show that this approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs.
+A simple operator-selection approach to reducing the number of mutants can reduce the cost of our approach by more than 70\%, while producing a gain in fault identification accuracy.
+We additionally show that our approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs.

### 2019-02-04 journal/introduction.tex
done
more small fixes

@@
 When a failure is detected by differential testing \cite{Differential}, where two compilers or optimization levels disagree, the problem is so hard that it may turn potential users away from automated test generation \cite{PLDI13}.
+Such \emph{wrong code} bugs do not produce a crash, only two compiled programs with different, incorrect (in at least one case) behavior.
 Even once a fault is identified, fixing the problem can be very difficult.
@@
 This paper proposes the use of mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.
-The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault.
+The underlying idea is that if two failures can be repaired, by applying the same program mutant, \emph{even if the repair only avoids the failure and does not truly fix it}, the two failures are likely due to the same fault.
 Repairs can also provide  information for understanding and debugging discovered faults.

### 2019-02-06 journal/experiments.tex
shipping

@@
 Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this is unlikely due to the nature of the repairs.
-For the CoREBench faults, {\bf Repair} produced more than twice as many perfect localizations as the other methods, though it also did not provide any ranking for a fault in a large number of cases.
+For the CoREBench faults, {\bf Repair} produced more than twice as many perfect localizations as the other methods, though it did not provide any ranking for a fault in a large number of cases.
 
@@
 {\bf Repair} can distinguish repairing mutants in fine-grained ways that MUSE and MUSEUM cannot by relying on the onion-ring structure:  when a repair also repairs failures that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults, and so ranks it lower than more ``relevant'' repairs.
+MUSEUM relies on a large set of mutation operators, and assumes that faulty statements will have multiple repairs, compared to non-faulty statements, or that faulty statements will be less likely to cause passing tests to fail when mutated.
+Restricting ourselves to cases where {\bf Repair} and MUSEUM both provided a localization, we know that some faulty statement was in a repairing mutant, but that passing tests could not distinguish this mutant  from other repairs (since we only use repairs that cause no passing tests to fail).
+Many of the faulty statement repairs (as expected for optimizing compilers) were statement deletions and conditional negations.
+For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.
+Our suspicion is somewhat confirmed by the results above where we used our methods with the MUSEUM mutant result set, and obtained similar results.
 
@@
 
- Why does {\bf Repair} perform best in the \emph{compiler} setting in particular?
-It is impossible to be sure, but one  possibility is power to distinguish repairs.
-If there are many repairs for a failure, how can the most important ones be distinguished?
-MUSEUM relies on a large set of mutation operators, and assumes that faulty statements will have multiple repairs, compared to non-faulty statements, or that faulty statements will be less likely to cause passing tests to fail when mutated.
-Restricting ourselves to cases where {\bf Repair} and MUSEUM both provided a localization, we know that some faulty statement was in a repairing mutant, but that passing tests could not distinguish this mutant  from other repairs (since we only use repairs that cause no passing tests to fail).
-Many of the faulty statement repairs (as expected for optimizing compilers) were statement deletions and conditional negations.
-For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.
-Our suspicion is somewhat confirmed by the results below where we used our methods with the MUSEUM mutant result set.
-
 Finally, in contrast to many previously reported results in fault localization (which are typically over much simpler programs and faults than subtle compiler semantics bugs) all methods we applied very frequently failed to provide a useful localization at all for our compiler faults.
@@
 Such methods, of course, often perform worse than even {\bf Coverage}; e.g., for our fault \#8 (their fault ID 115), {\bf Repair} and {\bf Coverage} produced localizations at rank 3 and 6, respectively, while the best SBFL result was well over rank 32.
-We suspect that combining multiple methods, using some method to cross-assess confidence might be required to obtain reliably good compiler bug localizations, and assign good confidence scores to localizations.
+We suspect that combining multiple methods, using some method to cross-assess confidence, might be required to obtain reliably good compiler bug localizations, and assign good confidence scores to localizations.
 There is much to be done if automated debugging is ever to make the lives of compiler developers easier in most cases \cite{AutoHelp}, and a human study using real, complex compiler bugs would be ideal, since previous studies of localization effectiveness tend to focus on simple tasks even moderately experienced developers can potentially succeed at.
@@
 We assume a ``hot'' mutants approach in that our sampling is of compilable, test-passing mutants.
-A larger sample would be required, in proportion to how many mutants compile and pass tests, to achieve the same results in a scenario with only ``cold'' mutants, though the difference is not huge, since most mutants are ruled out by the fact that no passing test covers the mutant, rather than that they cause some passing test to fail.
+A larger sample would be required, in proportion to how many mutants compile and pass tests, to achieve the same results in a scenario with only ``cold'' mutants, though the difference is not huge, since most mutants are ruled out by the fact that no failing test covers the mutant, rather than that they cause some passing test to fail.
 The more significant difference in a cold-mutant scenario is that each mutant/test evaluation is of course much cheaper without the cost (even amortized over many tests) of compilation, not the total number of mutants considered.
@@
 However, with hot mutants, it is easy, if localization using only negations and deletions is not helpful, to use the same method as proposed above for random selection to compute a more precise localization for a test.
-This should usually not be required, however:  we noted that in most cases where a mutant involved in a localization did use a operator replacement, the mutant was usually equivalent in behavior to a condition negation.
+This should usually not be required, however:  we noted that in most cases where a mutant involved in a localization was an operator replacement, the mutant was usually equivalent in behavior to a condition negation.
 Constant replacement repairs were simply very rare and not involved in any of the useful compiler localizations.

### 2019-02-06 journal/relatedwork.tex
shipping

@@
 
-The most closely related work is that of Chen et.
+The most closely related work (other than the ISSRE 2018 \cite{ISSRE18Mutants} paper that we extend) is that of Chen et.
 al \cite{PLDI13}, whose FPF \cite{Gonzalez} algorithm and benchmark we use.
-We extend their work with a better, and more universal distance metric, as well as an approach to fault localization in addition to fuzzer taming.Prior to the FPF-based work, Francis et al.
+We extend their work with a better, and more universal distance metric, as well as an approach to fault localization in addition to fuzzer taming.
+Prior to the FPF-based work, Francis et al.
 \cite{Podgurski04}, Podgurski et al.

### 2019-02-06 journal/abstract.tex
small fix

@@
 We evaluate our metric using more than 50 faults and 2,800 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.
-A simple operator-selection approach to reducing the number of mutants can reduce the cost of our approach by more than 70\%, while producing a gain in fault identification accuracy.
+A simple operator-selection approach to reducing the number of mutants can reduce the cost of our approach by 70\%, while producing a gain in fault identification accuracy.
 We additionally show that our approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs.

### 2019-09-13 journal/relatedwork.tex
SCB

@@
 \cite{OneTest} provides an alternative approach to fault identification, based on term rewriting to normalize failures, but is not currently applicable to compiler fuzzing, as it only supports tests that are sequences of method calls in Python \cite{ISSTA15,tstlsttt}.
+
+One recent approach does, at a conceptual level, resemble our methods:  \emph{semantic crash bucketing} (SCB) improves the fuzzer taming ability of AFL \cite{aflfuzz}, Honggfuzz \cite{honggfuzz}, and CERT BFF \cite{BFF}, by approximating real bug fixes with lightweight program modifications \cite{SCB}.
+The primary difference between SCB and our approach is that we generate many more potential approximate fixes, via mutation, but use no semantic feedback from failing inputs other than code coverage.
+The selectivity of SCB dramatically improves its scalability and ease-of-use, but limits applicability to cases where a patch template has been manually constructed, e.g., buffer overruns and null pointer dereferences.
+While these cases are by far the most common in traditional security/crash-based fuzzing, crashes are usually the \emph{uninteresting} cases in compiler fuzzing, or fuzzing/symbolic-execution based property-based \cite{ClaessenH00} testing \cite{DeepState,DeepStateTutorial,deepstaterepo}.
+It is unlikely that enough patch templates for the interesting changes to compiler code, or invariant-preserving code in other fuzzing, can be manually produced to make SCB alone generalize to these settings.
+Another difference is that SCB provides bucketing, rather than ranking; this is in some cases ideal (a user of AFL usually just wants to count ``different'' vulnerabilities that are memory-safety based), but in settings where the fixes are even more approximate and overlapping, and the bugs more complex, ranking is most useful.
 

### 2019-09-19 journal/introduction.tex
start response, edits

@@
 $S$ of items to rank, and a distance metric $d$, FPF computes $r_i$ as $s \in S: \forall s' \in S: min_{ j < i}(d(s,r_j)) \geq min_{j < i}(d(s',r_j))$.
+The condition on $s$ is obviously true when $s = s'$, or when $s' = r_j$ for some $j < i$; the other cases for $s'$, however, force selection of some (possibly non-unique) max-min-distance $s$.
 
@@
 
-This not a claim that being repaired by the same mutant perfectly predicts fault equivalence.
+This is not a claim that being repaired by the same mutant perfectly predicts fault equivalence.
 To understand why that seems highly unlikely, consider why a mutant of a compiler might cause a test case to stop failing.

### 2019-09-19 journal/localizations.tex
more small fixes
start response, edits

@@
 \begin{quote}
-For a set $F$ of failing test cases and $m$ of mutants, 
-the {\bf Repair} localization for $f \in F$, where $R(f) \in m$ is the set
+For a set $F$ of failing test cases and $M$ of mutants, 
+the {\bf Repair} localization for $f \in F$, where $R(f) \subseteq M$ is the set
 of mutants repairing $f$, ranks
-statements by the function $r$:
+statements by the function $r_f(m)$:
 %$$dmax(m,f) = max_{t : m \in R(t)} d(t,f)$$
 \[
-r(f) = 
+r_f(m) = 
 \begin{cases}
@@
 from $f$ to the most distant failing test also repaired by $m$, adding 1
-so that if $m$ only repairs failures at distance 0 $r$ is well defined
+so that if $m$ only repairs failures at distance 0 $r_f$ is well defined
 (and maximal).
@@
 
-An advantage of this method over some statistical approaches to
-localization is that it is not potentially confused in
-settings with multiple faults.  Each localization/explanation is based
-on a single failing test case, which in most cases fails due to only one
-fault.  In fact, other faults (and their failures) theoretically \emph{help} {\bf Repair} rank the mutants.
-
 In general, we expect {\bf Repair} to be used in the context of
@@
 
-Another advantage over many statistical approaches to 
-localization is that our methods are not potentially confused in 
+An advantage of this method over some statistical approaches to 
+localization is that it is not potentially confused in 
 settings with multiple faults.  Each localization/explanation is based 
 on a single failing test case, which in most cases fails due to only one 
-fault.  In fact, other faults help both methods rank localizations.
+fault.  In fact, other faults (and their failures) theoretically \emph{help} {\bf Repair} rank the mutants.
 

### 2019-09-19 journal/metrics.tex
start response, edits

@@
 
-This metric, the portion of mismatches over bits that are 1 in either
-vector, matches our intuitions about the above comparisons.  It
+\noindent In other words, the Jaccard distance is $\frac{\mathit{bitcount}(u
+\veebar v)}{\mathit{bitcount}(u \vee v)}$ (the number of 1 bits in the
+XOR over the number of 1 bits in the OR for $u$ and $v$); that is, the
+distance is the portion of mismatches over repairing mutants for
+either test.  This metric agrees with our intuitions about the above comparisons.  It
 essentially formalizes the intuition of the onion-ring model, where

### 2019-09-19 journal/stvrresponse.tex
start response, edits

\documentclass{letter}
%\def\today{September 10, 2019}
\usepackage[letterpaper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}

\begin{document}
\begin{letter}{}

  \address{Alex Groce\\School of Informatics, Computing \& Cyber Systems\\Northern Arizona University}

  \opening{Dear editor and reviewers:}

\noindent {\bf Reviewer 2:}

\begin{itemize}
\item {\bf pg 2. Line 27. Is it valid for $s'$ to range over all $S$? Should it instead range over $S \ { r_j | j < i }$? Otherwise, isn't $min_{j<i}(d(s', r_j))$ guaranteed to be 0? For example, for $d(r_0, r_0))$.}  If $s'$ is an already-ranked item, then indeed the $min$ will be zero.  That means that any $s$ will satisfy the inequality.  But for $s'$ not already ranked, the $min$ is only over distance to already-ranked items ($r_0 \ldots r_{i-1}$), and could be larger, thus falsifying the inequality.  In other words, the condition means that ranking any item next is as good as ranking an already-ranked item, but this only matters when all items are either already ranked or at distance zero from some already ranked item.  The condition can be re-worded as follows:  ``for all $s'$, the closest item to $s$ is at least as far away as the closest item to $s'$.''  This is trivially true when $s' = s$ and when $s' = r_j$ for some $j<i$.  But the other possible $s'$, not-ranked and not equal to $s$, force the ranking of the ``maximum minimum'' item.
\end{itemize}

  \closing{Sincerely,\\Alex Groce (agroce@gmail.com)\\Associate
    Professor\\
    School of Informatics, Computing \& Cyber Systems\\Northern
    Arizona University \vspace{0.1in}\\Josie Holmes\\Affiliated Researcher\\
    School of Informatics, Computing \& Cyber Systems\\Northern Arizona University}

\end{letter}
\end{document}

### 2019-09-19 journal/experiments.tex
initial work

@@
         \label{gcccurves}
-    \caption{Discovery curves compared to best curves from Chen et. al PLDI 2013 data (FPF benchmark)}
+    %\caption{Discovery curves compared to best curves from Chen et. al PLDI 2013 data (FPF benchmark)}
 \end{figure}
@@
 The next three columns show localization rankings.
+A ranking of 1 means that a code location \emph{in} the faulty code was presented to the user as the most likely location of the fault by the methodxs.
+A ranking of, e.g., 22 however, would mean that a user examining code locations proposed as faulty by the method would not reach any faulty code until examining the 22nd such location proposed by the method.
 A dash for a column means that localization did not rank any faulty statements, or assigned all faulty statements suspiciousness of 0.0.

### 2019-09-20 journal/conclusions.tex
start on fuzzgoat

@@
 particular, we thank Shin Hong and Thierry Chekam for prompt
-assistance beyond the call of duty.}
+assistance beyond the call of duty.  We would like to thank the
+authors of the Semantic Crash Bucketing \cite{SCB} study, in
+particular Rijnard van Tonder and Claire Le Goues, for their assistance in
+comparing our results with SCB.

### 2019-09-20 journal/experiments.tex
start on fuzzgoat

--- 2019-02-06 journal/experiments.tex
@@
 The next three columns show localization rankings.
+A ranking of 1 means that a code location \emph{contained in} the actual faulty code locations was presented to the user as the most likely location of the fault by the method: the method has precisely identified the fault location.
+A ranking of, for instance, 22, on the other hand would mean that a user examining code locations proposed as faulty by the method would not reach any faulty code until examining the 22nd such location proposed by it.
 A dash for a column means that localization did not rank any faulty statements, or assigned all faulty statements suspiciousness of 0.0.
@@
 
+\subsubsection{Combinatorial Control of Optimizations}
+
+Finally, if we assume that in the setting of compilers, faults are due to optimizations, it may be possible to use a combinatorial approach to turning on/off optimizations as a cheap equivalent of mutants; Ghandehari et.
+al \cite{CombTestFaultLoc} proposed a combinatorial approach to fault localization and such a method might work fairly well in this setting.
+Further experimental effort would be required to determine how useful such a method would be; a major concern is that rather than simply ``turning off'' optimizations as one might at the command line, some mutants change the result of a check as to whether an optimization is valid for a piece of code.
+GCC and LLVM also do not, unfortunately\footnote{Unfortunately, at least, from this combinatorial point of view; as developers we may suspect that modern compilers already have plenty of command-line options.}, expose all the optimizations they perform as individually controllable by users.
+Based on our examination, the level of control over optimizations provided by GCC 4.3.0 was simply too coarse-grained to enable effective replacement of mutants, and SpiderMonkey 1.6 offers almost no control over optimizations, as would be expected of a JIT usually used in a browser context; it does not work as a command-line tool that takes an input file and produces a binary.
+There may be API calls to control optimization in some way, but these were not easy to discover, or practical for a developer to use.
+
+\section{Fuzzer Taming for A Non-Compiler Target: Fuzzgoat}
+
+%\subsection{Fuzzgoat}
+
+\begin{figure}
+   \centering
+        \includegraphics[width=1.0\textwidth]{afl_fuzzgoat.png}
+        \caption{AFL run for fuzzgoat}
+        \label{aflfuzzgoat}  
+  \end{figure}
+
+  Fuzzgoat (\url{https://github.com/fuzzstati0n/fuzzgoat}) is a well-known backdoored C program used to test fuzzers and other analysis tools.
+It is an interesting subject to consider non-compiler fuzzer taming because it is a single program version with 4 distinct bugs with complete ground truth, and widely viewed as representative of a typical fuzzing use-case.
+We ran AFL \cite{aflfuzz} for over 7 hours on the program, generating 4,244 crashing inputs (Figure \ref{aflfuzzgoat} shows a screenshot from just before we terminated the run).
+AFL's internal triaging methods reduced this to a set of 38 crashes it considered distinct.
+We terminated the run once it had been nearly three hours since a new distinct crash was detected.
+
+  These crashes exhibit a classic fuzzer taming structure.
+Three of the bugs are easy to identify: they are triggered by 15, 14, and 11 of the 38 AFL-produced crashing inputs, respectively, and all have multiple inputs that uniquely detect that bug, and no other.
+The fourth bug, however, is only triggered by \emph{three} of the 38 inputs, and is never uniquely present (any input triggering it also triggers at least one other bug).
+This bug was only found by AFL after more than an hour of fuzzing, in fact.
+
+  This is a classic fuzzer taming problem, albeit on a smaller scale than with the compilers:  a user will probably not want to inspect the highly-redundant set of 38 inputs, but will want to see all four distinct faults, while examining as few inputs as possible.
+
+  Applying the mutation-repair based approach is relatively easy here.
+We used the {\bf universalmutator} tool \cite{RegExpMut} to generate 18,489 mutants for fuzzgoat.
+This tool is far more aggressive than Andrews' tool, in terms of operators used, and is more widely available to typical users.
+Of these 6,809 mutants actually compile, and trivial compiler equivalance \cite{TCE} can further reduce the set to 2,899 mutants to be used in producing repair bitvectors.
+Computing the full bitvector set and all distances for the 38 crashes required just over 10 minutes.
+While compiling and running mutants is a serious computational burden in the setting of a production compiler, it is negligible compared to the time required for fuzzing, in a typical fuzzing session, we suspect, even without using such simple additional reductions as checking the mutants for code covered by some crashing input.
+This is partly because the number of mutants to consider is smaller, and partly because compiling GCC or SpiderMonkey is simply much slower than compiling more typical small-to-medium-sized fuzzing targets.
+Fuzzing a very complex target, such as a database (e.g., LevelDB or RocksDB) might be more like compiler fuzzing, however.
+Input runtime is also smaller for most non-compiler targets, we suspect.
+
+\begin{table}
+\centering
+{\scriptsize
+\begin{tabular}{|l||r|r|r|r|}
+\hline
+Method & Cutoff & Mean bugs & \# Perfect & \% Perfect\\
+  \hline
+  Repair & 3 & 3.00 & 0 & 0.00\% \\
+  Coverage & 3 &  3.10 & 13 & 34.21\% \\
+  Random & 3 & 2.58 & 3 & 7.89\%\\
+  \hline
+  Repair & 4 & 3.42 & 16 & 42.10\%\\
+  Coverage & 4 &  3.39 & 15 & 39.47\% \\
+  Random & 4 & 2.74 & 6 & 15.79\%\\  
+  \hline 
+    Repair & 5 & 4.00 & 38 & 100.00\%\\
+  Coverage & 5 &  3.50 & 19 & 50.00\%\\
+  Random & 5 & 3.29 & 16 & 42.10\%\\  
+  \hline
+    Repair & 6 & 4.00 & 38 & 100.00\% \\
+  Coverage & 6 &  3.50 & 19 & 50.00\%\\
+  Random & 6 & 3.18 & 11 & 28.95\%\\
+  \hline
+  \multicolumn{5}{c}{$\ldots$} \\
+\hline 
+    Repair & 10 & 4.00 & 38 & 100.00\% \\
+  Coverage & 10 &  3.63 & 24 & 63.16\%\\
+  Random & 10 & 3.58 & 23 & 60.53\%\\  
+  \hline
+    Repair & 11 & 4.00 & 38 & 100.00\% \\
+  Coverage & 11 &  3.63 & 24 & 63.16\%\\
+  Random & 11 & 3.66 & 25 & 65.79\%\\  
+  \hline
+    \multicolumn{5}{c}{$\ldots$} \\
+  \hline
+    Repair & 17 & 4.00 & 38 & 100.00\% \\
+  Coverage & 17 &  3.97 & 37 & 97.37\%\\
+  Random & 17 & 3.84 & 32 & 84.21\%\\  
+  \hline
+    Repair & 18 & 4.00 & 38 & 100.00\% \\
+  Coverage & 18 &  4.00 & 38 & 100.00\% \\ 
+  Random & 18 & 3.82 & 31 & 81.58\%\\  
+  \hline
+\end{tabular}
+}
+\caption{Fuzzgoat triage results}
+%\vspace{-0.2in}
+\label{table:fuzzgoat}
+\end{table}
+
+Table \ref{table:fuzzgoat} shows fuzzer taming results for fuzzgoat, using our repairing-mutant-based approach, Chen et.
+al's \cite{PLDI13} statement coverage method, and (as a baseline) random ordering.
+For repair-based ranking and statement coverage-based ranking, we show the mean result over all 38 possible choices for first test to rank.
+Random is the mean over 38 random rankings.
+{\bf Cutoff} indicates the point at which the user stops examining crashing inputs, and {\bf Mean bugs} indicates how many of the four distinct faults, on average, the method, with that cutoff, allowed the user to see within that cutoff.
+{\bf \# Perfect} describes how many of the 38 experiments produced a perfect triage, where all of the bugs were seen before the cutoff.
+We begin {\bf Cutoff} at 3, since 3 is the minimum possible number of inputs a user must examine to find all four bugs (some inputs cause a crash due to either of two bugs, so are able to identify two faults).
+Coverage-based FPF is better than repair-based FPF, for fuzzer taming, if a user only examines 3 of the 38 crashes.
+However, once a user is willing to look at 4 inputs, mutation-repair is better, and if the user is willing to look at 5 of the 38 inputs, our approach guarantees all of the distinct bugs will be seen.
+Statement coverage-based FPF, in contrast, requires a user to look at {\bf 18} inputs before providing such a guarantee.
+Indeed, there is a more than 25\% chance of missing one bug even if the user looks at 10 inputs, and random ranking performs almost as well as coverage-based FPF, once a user is willing to examine 10 or more inputs.
+In the table, cutoffs for which statement-based coverage performance is identical to the previous cutoff are omitted.
+That the coverage-based ranking does not improve its performance between cutoff of 6 and 9, and 11 and 16, shows how weak the signal from the distinct faults is, in terms of code coverage.
+
 \section{Threats to Validity}

### 2019-09-20 journal/experiments.tex
small add

@@
   Applying the mutation-repair based approach is relatively easy here.
-We used the {\bf universalmutator} tool \cite{RegExpMut} to generate 18,489 mutants for fuzzgoat.
-This tool is far more aggressive than Andrews' tool, in terms of operators used, and is more widely available to typical users.
-Of these 6,809 mutants actually compile, and trivial compiler equivalance \cite{TCE} can further reduce the set to 2,899 mutants to be used in producing repair bitvectors.
+We used the universalmutator tool \cite{RegExpMut} (\url{ https://github.com/agroce/universalmutator}) to generate 18,489 mutants for fuzzgoat.
+This tool is far more aggressive than Andrews' tool, in terms of operators used, and, more importantly is more widely available to typical users and easy to apply to source code in a large set of languages, unlike, e.g., Proteum.
+Of these mutants, only 6,809 actually compile, and trivial compiler equivalance \cite{TCE} can further reduce the set to 2,899 mutants to be used in producing repair bitvectors.
 Computing the full bitvector set and all distances for the 38 crashes required just over 10 minutes.
-While compiling and running mutants is a serious computational burden in the setting of a production compiler, it is negligible compared to the time required for fuzzing, in a typical fuzzing session, we suspect, even without using such simple additional reductions as checking the mutants for code covered by some crashing input.
+While compiling and running mutants is a serious computational burden in the setting of a production compiler, it is negligible compared to the time required for fuzzing, in a typical fuzzing effort, we suspect, even without using such simple additional reductions as checking the mutants for code covered by some crashing input.
 This is partly because the number of mutants to consider is smaller, and partly because compiling GCC or SpiderMonkey is simply much slower than compiling more typical small-to-medium-sized fuzzing targets.
-Fuzzing a very complex target, such as a database (e.g., LevelDB or RocksDB) might be more like compiler fuzzing, however.
-Input runtime is also smaller for most non-compiler targets, we suspect.
+Fuzzing a very complex target, such as a complex production-grade database (e.g., LevelDB or RocksDB) or the Linux kernel itself might be more like compiler fuzzing, however, in terms of mutation analysis cost.
+Input runtime is also smaller for most non-compiler targets, we suspect, even including ``hard'' targets like databases and kernels.
 

### 2019-09-21 journal/conclusions.tex
much more

@@
 (with effect size of about 3\% improvement), over
-the \emph{best} (for each subject program) of over 16 metrics considered in previous work.
+the \emph{best} (for each subject program) of over 16 metrics
+considered in previous work.
+
+We also demonstrated that our approach can be cheap and effective in
+fuzzer taming not targeting compilers.  It performed much better than
+a coverage-based ranking approach, reliably revealing distinct faults
+in the first few inputs presented, and is more general in application
+than the (to our knowledge) best competing method, semantic crash
+bucketing \cite{SCB}.
+
 Preliminary investigation also shows improvement (for a shared
@@
 particular Rijnard van Tonder and Claire Le Goues, for their assistance in
-comparing our results with SCB.
+comparing our results with SCB.}

### 2019-09-21 journal/experiments.tex
much more

@@
 The fourth bug, however, is only triggered by \emph{three} of the 38 inputs, and is never uniquely present (any input triggering it also triggers at least one other bug).
-This bug was only found by AFL after more than an hour of fuzzing, in fact.
+The bug frees a pointer that can subsequently be used, and it is (as often with use-after-frees \cite{DangSan}) hard to produce inputs that actually cause this to happen.
+This bug was only found by AFL after nearly an hour of fuzzing, in fact.
 
@@
 
+To our knowledge, the state-of-the-art in research on advancing fuzzing triage over the tools built into fuzzers like AFL is represented by the \emph{semantic crash bucketing} (SCB) approach of van Tonder, Kotheimer, and Le Goues \cite{SCB}.
+This approach uses patch templates based on common types of bugs/semi-fixes, combined with semantic information obtained by observing the execution of a crashing input, to bucket crashing inputs discovered by fuzzers.
+We contacted the SCB authors, and with their cooperation investigated how well SCB handled the fuzzgoat crashes.
+The fuzzgoat code highlights the primary limitation of SCB: namely, if a kind of failure is not handled by a (manually constructed) patch template and identified by the semantic feedback mechanisms, then SCB does not really handle it.
+The SCB tool was able to produce a patch for one of the four fuzzgoat bugs, a null pointer dereference.
+Unfortunately, at present, SCB does not handle bugs related to misuse of {\tt free} (use-after-free, double frees, or frees of invalid pointers), and the other three fuzzgoat bugs are all {\tt free}-related.
+SCB therefore can very effectively identify the inputs associated with one of the four bugs (when it is the only one triggeered by the input), but provides no assistance with finding the other three bugs.
+This is not to say that SCB is useless here; the null pointer bug alone is responsible for 13 of the 38 crashes, so SCB does make it easy to see one example of the null pointer bug, and ignore another 12 inputs.
+It does not, however, help find the rare bug with only a few instances among the remaining 25 inputs.
+While mutation repair also has no knowledge of {\tt free} bugs, it is able to find mutants that avoid their symptoms or their causes, and thus triage all four faults effectively.
+The SCB team added the fuzzgoat example to their published VM, and plan to use it to drive development of patch templates and analysis for use-after-frees; this will, at some point, likely make it possible for SCB to handle fuzzgoat effectively.
+However, in the long term, if fuzzing is applied to find assertion violations, differential testing divergences \cite{Differential,ICSEDiff}, or other more complex ``custom'' properties, as is likely with the appearance of tools such as DeepState \cite{DeepState} that merge fuzzing and property-based testing, it is hard to see how SCB will handle these ``custom'' bug types.
+
+Fuzzgoat is a poor candidate for fault localization, due to the lack of a reasonable set of passing tests, and due to the fact that all four bugs are fairly easy to understand once an appropriate input is run under a sanitizer.
+We do note that repair manages to localize the null pointer dereference perfectly (SCB's patch is also an excellent localization for this bug).
+Given passing tests, we suspect all fault localization methods discussed in this paper could handle the localization problem easily, in contrast to a typical compiler bug.
+
 \section{Threats to Validity}

### 2019-09-21 journal/relatedwork.tex
much more

@@
 
+Explaining, rather than merely localizing, bugs is a less common goal, though all mutation-based approaches can probably be adapted to this idea (since a mutation is, itself, a kind of explanation).
+Key work on explaining bugs in addition to localizing them includes Groce's early model-checking work \cite{GroceError,ChakiLev} and the MintHint \cite{MintHint} approach, which bears some resemblance to the use of mutants as ``hints'' for fault repair, but is more general and backed by user studies to show its utility.
+The real-world value of MintHint explanations supports the idea that mutant-based explanations might be useful to developers.
+
 Using program modifications or mutants broadly defined to repair programs has been a popular topic of recent work as well \cite{GenProg,AutoRep,WeiFix}.

### 2019-09-24 journal/abstract.tex
related

@@
 A simple operator-selection approach to reducing the number of mutants can reduce the cost of our approach by 70\%, while producing a gain in fault identification accuracy.
-We additionally show that our approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs.
+We additionally show that our approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs, and can help triage crashes found in fuzzing of non-compiler programs more effectively than a state-of-the-art technique.

### 2019-09-24 journal/experiments.tex
related

@@
 In fact, on examination, only two of the 11 crash bugs are not distinguished by simply examining the crash message.
-Faults that produce a crash are also likely to be much easier to debug than semantic problems such as all of the GCC wrong code bugs and most of the JavaScript faults --- e.g., following a dynamic slice might well suffice.
+Faults that produce a crash are also likely to be much easier to debug than semantic problems such as all of the GCC wrong code bugs and most of the JavaScript faults-- e.g., following a dynamic slice might well suffice.
 
@@
 For GCC, the oracle is a differential check on a hash code:  failure was defined as producing a executable with the -O3 flag that, when executed, produced a different checksum than code compiled by either of GCC 4.9.3 or clang 7.0.0, both using -O0.
-A repairing mutant must enable GCC 4.3.0 to actually compile code correctly, which is a very strict correctness property, making coincidental correctness \cite{CCT} highly unlikely --- only missed optimizations are allowed.
+A repairing mutant must enable GCC 4.3.0 to actually compile code correctly, which is a very strict correctness property, making coincidental correctness \cite{CCT} highly unlikely---only missed optimizations are allowed.
 
@@
 An additional measure of effectiveness is to consider how effective a metric is in producing matched nearest neighbors.
-That is, how often is the nearest neighbor of a failure (that is not due to a singleton fault --- a fault detected by only one test case) due to the same fault?
+That is, how often is the nearest neighbor of a failure (that is not due to a singleton fault---a fault detected by only one test case) due to the same fault?
 For reduced \cite{DD,PLDI13,
@@
 
-No method performs extremely well --- for half the faults, no methods produced what we consider a useful localization.
+No method performs extremely well---for half the faults, no methods produced what we consider a useful localization.
 Compiler faults are hard to debug, and any help is useful, but it seems unlikely developers will really benefit from a localization when it does not rank at least one faulty statement in the first 25-30 statements.
@@
 A few mutants fixed a very large number of tests, up to 1,050.
-These all appear to be turning off all optimizations --- essentially running gcc in -O0 mode.
+These all appear to be turning off all optimizations---essentially running gcc in -O0 mode.
 Most mutants fixed only a few tests, to a greater extent than was true with SpiderMonkey.
@@
 
-Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both benchmark metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement --- even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults) --- is more than 3.8\% better than either benchmark metric.
+Slicing the GCC tests into 20 random equal-sized test subsets and comparing with function and line coverage metrics (Figure \ref{comparegcc}) we find that differences in APFD values are statistically significant between our metric and both benchmark metrics, by Wilcoxon test, with $p < 0.0005$, and the mean APFD improvement---even for only 55 tests, exposing only 6-11 faults (mean of 7.75 faults)---is more than 3.8\% better than either benchmark metric.
 
@@
 This is basically a trade-off.
-MUSE, MUSEUM, and Metallaxis-FL provide localization for many more faults, in a more diverse range of settings (with single failing tests, no repairs, etc.), and using more diverse information --- but, in cases where {\bf Repair} provides a localization, it seems to often be higher quality due to its restriction to a high-signal source of information (mutants repairing failing tests, only).
+MUSE, MUSEUM, and Metallaxis-FL provide localization for many more faults, in a more diverse range of settings (with single failing tests, no repairs, etc.), and using more diverse information---but, in cases where {\bf Repair} provides a localization, it seems to often be higher quality due to its restriction to a high-signal source of information (mutants repairing failing tests, only).
 In settings, unlike compilers, where there are relatively few repairs, {\bf Repair} also produces very little incorrect information, and so has a low cost to the user in terms of wasted attention.
-Note that our heuristic to stop reading after 5 wrong localizations is not even needed for all but one CoREBench fault --- {\bf Repair} suggests 1-3 statements at most.
+Note that our heuristic to stop reading after 5 wrong localizations is not even needed for all but one CoREBench fault---{\bf Repair} suggests 1-3 statements at most.
 
@@
 E.g., trivial compiler equivalence (TCE) \cite{TCE} can reject equivalent and redundant mutants, removing almost 30\% of mutants for benchmark subjects.
-Other techniques, such as using mutant schemas to avoid having to compile and store each mutant, are also applicable --- almost all of the techniques characterized by Offutt and Untch \cite{offutt2001mutation} as \emph{do faster} approaches apply, without any loss of ability to identify or localize faults.
+Other techniques, such as using mutant schemas to avoid having to compile and store each mutant, are also applicable---almost all of the techniques characterized by Offutt and Untch \cite{offutt2001mutation} as \emph{do faster} approaches apply, without any loss of ability to identify or localize faults.
 
@@
 It is also possible that faults in the framework used for experiments introduced errors into our results.
-In order to make it possible to check for these problems, we have made the raw data for a re-analysis available: \url{https://github.com/agroce/mutants16/tree/master/tests}.
+In order to make it possible to check for these problems, we have made the raw data for a re-analysis of the PLDI13 benchmark available, including the tests, ground truth, and mutants, at \url{https://github.com/agroce/compilermutants}; the fuzzgoat example is also available, including source code for a simple version of our FPF algorithms, at \url{https://github.com/agroce/fuzzgoattriage}.
 The data required for analysis of the non-compiler faults is available from the authors of the respective studies (see the acknowledgements at the end of the paper).

### 2019-09-30 journal/abstract.tex
more on fuzzgoat and SCB

--- 2019-02-06 journal/abstract.tex
@@
 A simple operator-selection approach to reducing the number of mutants can reduce the cost of our approach by 70\%, while producing a gain in fault identification accuracy.
-We additionally show that our approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs.
+We additionally show that our approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs, and can help triage certain types of crashes found in fuzzing non-compiler programs more effectively than a state-of-the-art technique.

### 2019-09-30 journal/conclusions.tex
more on fuzzgoat and SCB

@@
 
+Finally, the results of our comparison with semantic crash bucketing
+(SCB) \cite{SCB},
+which we consider the state-of-the-art in general-purpose fuzzer
+taming, suggest what may be the most important line of future work.
+When SCB applies to a set of crashes, it is cheap, effective, and
+provides a useful bucketing.  However, it is very hard to imagine the
+manual-patch template approach of SCB ever applying to complex compiler
+bugs, or even to all the kinds of faults that might be found in
+fuzzing of other targets, as discussed in Secton \ref{fuzzgoat}.  We
+believe, however, that a
+combination of SCB with our FPF-and-mutant-repair approach, perhaps using operator
+selection to approximate a large, imprecise, set of ``patch
+templates,'' and using semantic guidance to select among these, might offer a fuzzer taming method that is both
+extremely practical (like SCB) but also truly general-purpose (like
+our mutation-based approach).  In Section \ref{fuzzgoat} we propose a
+simple two-phase approach, using SCB first, then our approach, but we
+believe a more principled combination of the methods might provide
+truly effective fuzzer taming, at a reasonable cost.
+
 {\scriptsize {\bf Acknowledgements:} The authors would like to thank the authors of

### 2019-09-30 journal/experiments.tex
more on fuzzgoat and SCB

@@
-\section{Experimental Results}
+\section{Core Experimental Results: Compiler Triage and Debugging}
 
@@
 
-\subsection{Discussion}
+\section{Discussion}
 
@@
 \section{Fuzzer Taming for A Non-Compiler Target: Fuzzgoat}
-
+\label{fuzzgoat}
 %\subsection{Fuzzgoat}
@@
 
-  These crashes exhibit a classic fuzzer taming structure.
+  These crashes exhibit the usual fuzzer taming structure, where some bugs are far more frequently found than others.
 Three of the bugs are easy to identify: they are triggered by 15, 14, and 11 of the 38 AFL-produced crashing inputs, respectively, and all have multiple inputs that uniquely detect that bug, and no other.
@@
 SCB therefore can very effectively identify the inputs associated with one of the four bugs (when it is the only one triggeered by the input), but provides no assistance with finding the other three bugs.
-This is not to say that SCB is useless here; the null pointer bug alone is responsible for 13 of the 38 crashes, so SCB does make it easy to see one example of the null pointer bug, and ignore another 12 inputs.
-It does not, however, help find the rare bug with only a few instances among the remaining 25 inputs.
+This is not to say that SCB is useless here; the null pointer bug is responsible for 15 of the 38 crashes, so SCB does make it easy to see an instance of the null pointer bug, and ignore 14 additional inputs.
+It does not, however, help find the rare bug with only a few instances among the remaining 23 inputs.
 While mutation repair also has no knowledge of {\tt free} bugs, it is able to find mutants that avoid their symptoms or their causes, and thus triage all four faults effectively.
@@
 Given passing tests, we suspect all fault localization methods discussed in this paper could handle the localization problem easily, in contrast to a typical compiler bug.
+
+Unsurprisingly, the repairing-mutants approach does not improve on SCB on the set of bugs from the SCB paper \cite{SCB}.
+These bugs were intentionally selected to consist of null pointers and buffer overflows; SCB in this setting operates as a bucketing version of repair that only has to use a selected, known-relevant set of mutants.
+Repair is good at ranking the crashes for these subjects, but unable to provide bucketing, and more difficult to set up and more expensive (though still relatively cheap---it requires minutes for taming, vs.
+the seconds of SCB) than SCB.
+When SCB applies, it is likely to be faster and more useful than mutant-based FPF ranking.
+However, a user will obviously seldom know for certain that all faults in a set of crashes correlate to bug types that SCB supports.
+One approach would be to use SCB to bucket crashes, and then apply our approach to any remaining inputs, and examine the first few inputs in repair-FPF order.
+In Section \ref{conc} we will return to this point.
 

### 2019-09-30 journal/introduction.tex
more on fuzzgoat and SCB

@@
 \item introducing and evaluating another compiler fault localization approach (in Section \ref{sec:covloc}), based on code coverage changes,
-\item and including a much more in-depth examination of the computational costs of our approach and various ways of reducing that cost, including experimental evaluations (in Section \ref{sec:highcost}).
+\item including a much more in-depth examination of the computational costs of our approach and various ways of reducing that cost, including experimental evaluations (in Section \ref{sec:highcost}),
+\item and showing how our approach works for fuzzer taming of non-compiler bugs, highlighting the limitations of both our technique and the state-of-the-art semantic crash bucketing (SCB) \cite{SCB} approach.
 \end{itemize}

### 2019-09-30 journal/relatedwork.tex
related
ready to finish response

@@
 Finally, our work fits into the general framework of efforts to define distances between program executions.
+
+\subsection{Fuzzer Taming/Bug Triage}
 
@@
 
+\subsection{Fault Localization}
+
 Fault localization \cite{FaultSurvey,NearNeighbor,Liblit03,Liblit05,Cleve05,Jones2002,Jones05,Jones07,GroceError,ChakiLev,Liu06,SPIN03,Santelices:ICSE:2009,Abreu:2006:PRDC,xuan2014test,DD}, even specifically for compilers \cite{Whalley94}, is a long-standing field of research.
-Recently, fault localization researchers have used program mutants to improve statistical \cite{Jones2002} fault localization techniques \cite{MUSE,DebroyMutant,FasterMutant,multilingual,Metallaxis,Papadakis}.
-The core of this work has been the belief that faulty program locations are (1) more likely to change a failing test to successful when mutated and (2) less likely to change a successful test case to faulty when mutated.
+Many fault localization researchers have used program mutants to improve statistical \cite{Jones2002} fault localization techniques \cite{MUSE,DebroyMutant,FasterMutant,multilingual,Metallaxis,MetalSelect,Papadakis,MechLocal,MurtazaTrace,Musco}.
+One  core aspect of this work has been the belief that faulty program locations are (1) more likely to change a failing test to successful when mutated and (2) less likely to change a successful test case to faulty when mutated.
 This clearly connects to our general notion that how two test cases respond to a set of mutants is a good measure of their semantic similarity.
+
 The work of Hong et al.
@@
 
+Metallaxis \cite{Metallaxis,MetalSelect,Papadakis}, in contrast to MUSE/MUSEUM (and our own work), relies on the idea that mutants that are (relative to other killed mutants) mostly killed by failing tests likely have a connection to a program defect or defects.
+This idea relies on a fundamentally different notion of ``killing'' a mutant than the one primarily used in this paper.
+We generally speak of a mutant being killed when the mutant makes a previously passing test fail; in this sense, a failing test cannot kill a mutant.
+However, these ``correctness kills'' are not the only possible notion of mutant kill used in the literature; a different approach is to consider a mutant killed whenever it produces a different output than the original program.
+This point of view, unfortunately, is not obviously useful for compiler mutants, where almost all mutants, however benign, of covered code, will result in some (perhaps meaningless) change in the output binary.
+We have therefore consider a mutant killed only when it causes semantically wrong output from a compiled program, a fairly complex specification-based notion of difference.
+Compilers may therefore be an unusually poor fit for the Metallaxis approach, which, as our results above show, is quite competitive for more ``normal'' localization, where output (vs.
+specification-satisfaction/correctness) differences are likely to be meaningful.
+Extending our approach beyond ``correctness kills'' (at least for non-compiler contexts) is an interesting possible line of future research.
+
+Musco, Monperrus, and Preux offer yet another alternative mutant-based approach to localization, one that (like ours) is explicitly causal \cite{Musco}.
+Their approach, quite different than MUSE, MUSEUM, Metallaxis, or our own, uses mutants to mark likely causal relationships in call graphs.
+A recent survey of advances in mutation testing places all of these approaches using mutants in fault localization within the larger context of mutation testing research \cite{MutationSurvey}.
+
 Explaining, rather than merely localizing, bugs is a less common goal, though all mutation-based approaches can probably be adapted to this idea (since a mutation is, itself, a kind of explanation).
@@
 The real-world value of MintHint explanations supports the idea that mutant-based explanations might be useful to developers.
+
+\subsection{Automated Program Repair}
 
@@
 
+\subsection{Distance Metrics}
+
 The use of distance metrics in software engineering includes some of the localization and fault identification efforts discussed above (e.g., \cite{NearNeighbor,GroceError,ChakiLev,Liu06}).

### 2019-09-30 journal/stvrresponse.tex
initial work
response

@@
 
+We addressed all reviewer comments where this seeemed feasible within the turnaround time for the special issue, and in particular added a whole new experimental section on non-compiler fuzzer taming to address the generalizability threat.
+This new section includes use of mutants to ``fix'' use-after-free faults, which are quite different in character than wrong-code compiler faults, and suggests a possible more general, but low-cost, solution for fuzzer taming/triage.
+We thank the reviewers for their numerous comments, which allowed us to clarify many subtle points and to considerably extend the scope of related work that is addressed (adding 8 new references, plus much more discussion of some previously cited work).
+This new version is nearly 4 pages (over 10\%) longer than the previous version.
+
+  \noindent {\bf Reviewer 1:}
+
+  While the proposed approach is not very useful for fuzzing GCC or LLVM (unless new advances in compiler fuzzing appear, which does happen), it is likely very useful in fuzzing, e.g., Swift, Solidity, or any other compiler less aggressively fuzzed to date.
+Building a high quality fuzzer and establishing ground truths in such a setting, unfortunately, is beyond the scope of this paper.
+
+  We have added a number of comments to address these general concerns, and show that our technique can be both useful and (fairly) cheap in non-compiler fuzzing with a new case study, comparing to semantic crash bucketing.
+
+  A link to the released compiler mutants and all tests we used is now included in the paper (as is a link to the SCB comparison example).  We welcome other uses of our mutants and the tests!
+
 \noindent {\bf Reviewer 2:}
+
+We added an application beyond compiler bugs, and associated discussion, as suggested.
+The comparison with SCB also suggests a promising avenue of future work, which we are discussing now with the SCB team.
 
@@
 But the other possible $s'$, not-ranked and not equal to $s$, force the ranking of the ``maximum minimum'' item.
+We added a clarifying sentence, though it's still a somewhat subtle formula, we agree.
+\item {\bf pg 5. Line 19.} Thanks!  This makes the metric much easier to understand.
+  \item {\bf pg 6. Line 23.} Again, thanks, we fixed the notation to make this easier to follow, as proposed.
+  \item All other detailed comments have been addressed, and indeed were errors/possible improvements.
   \end{itemize}
+
+  \noindent {\bf Reviewer 3:}
+
+  We added a discussion of the idea of using combinatorial approaches to handle compiler optimizations in the section on possible speed improvements.
+
+  While unfortunately mutant vs.
+correctness kills is hard to explore experimentally, in the compiler context, we added a discussion of the distinction to related work, in the process of considerably extending our examiniation of related work using mutants to localize faults.
 

### 2019-10-01 journal/metrics.tex
small fixes' git push

@@
 
-For every failing test case, its response to mutants defines a bit-vector, with
-length equal to the number of mutants that repaired any test case,
+For every failing test case, the response (fixing or not fixing the
+failure) to all mutants defines a bitvector, with
+length equal to the number of mutants that repaired \emph{any} test
+case (or all mutants; it does not matter for our proposed metric),
 where a 1 bit indicates that the mutant in question repairs this test
@@
 causes the test to pass, \emph{and} the mutant does not cause
-other, previously passing tests, to fail.}.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
+other, previously passing tests, to fail.}.  These bitvectors cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
 by 100 mutants.  If these two test cases share 90 common repairing

### 2019-10-01 journal/experiments.tex
missing word

@@
 test; in our compiler experiments, there were no repairs that did not change
-coverage; this would be very unlikely to hold in other software settings, where many bugs related to data values; for well-fuzzed compilers, however, few bugs are simple, e.g., off-by-one bugs in an integer value.
+coverage; this would be very unlikely to hold in other software settings, where many bugs are related to data values; for well-fuzzed compilers, however, few bugs are simple, e.g., off-by-one bugs in an integer value.
 If a fault can be avoided by disabling a

### 2019-10-01 journal/experiments.tex
dash

@@
 In fact, on examination, only two of the 11 crash bugs are not distinguished by simply examining the crash message.
-Faults that produce a crash are also likely to be much easier to debug than semantic problems such as all of the GCC wrong code bugs and most of the JavaScript faults-- e.g., following a dynamic slice might well suffice.
+Faults that produce a crash are also likely to be much easier to debug than semantic problems such as all of the GCC wrong code bugs and most of the JavaScript faults --- e.g., following a dynamic slice might well suffice.
 

### 2019-10-01 journal/experiments.tex
text fix

@@
  Fuzzgoat (\url{https://github.com/fuzzstati0n/fuzzgoat}) is a well-known backdoored C program used to test fuzzers and other analysis tools.
-It is an interesting subject to consider non-compiler fuzzer taming because it is a single program version with 4 distinct bugs with complete ground truth, and widely viewed as representative of a typical fuzzing use-case.
+It is an useful subject for demonstrating non-compiler fuzzer taming because it is a single program version with 4 distinct bugs with complete ground truth, and widely viewed as representative of a typical fuzzing use-case.
 We ran AFL \cite{aflfuzz} for over 7 hours on the program, generating 4,244 crashing inputs (Figure \ref{aflfuzzgoat} shows a screenshot from just before we terminated the run).

### 2019-10-01 journal/experiments.tex
typo fix

@@
 Unfortunately, at present, SCB does not handle bugs related to misuse of {\tt free} (use-after-free, double frees, or frees of invalid pointers), and the other three fuzzgoat bugs are all {\tt free}-related.
-SCB therefore can very effectively identify the inputs associated with one of the four bugs (when it is the only one triggeered by the input), but provides no assistance with finding the other three bugs.
+SCB therefore can very effectively identify the inputs associated with one of the four bugs, but provides no assistance with finding the other three bugs.
 This is not to say that SCB is useless here; the null pointer bug is responsible for 15 of the 38 crashes, so SCB does make it easy to see an instance of the null pointer bug, and ignore 14 additional inputs.
@@
 These bugs were intentionally selected to consist of null pointers and buffer overflows; SCB in this setting operates as a bucketing version of repair that only has to use a selected, known-relevant set of mutants.
-Repair is good at ranking the crashes for these subjects, but unable to provide bucketing, and more difficult to set up and more expensive (though still relatively cheap---it requires minutes for taming, vs.
+Repair is good at ranking the crashes for these subjects, but unable to provide bucketing, and more difficult to set up and more expensive (though still relatively cheap --- it requires minutes for taming, vs.
 the seconds of SCB) than SCB.

### 2019-10-01 journal/relatedwork.tex
typo

@@
 This point of view, unfortunately, is not obviously useful for compiler mutants, where almost all mutants, however benign, of covered code, will result in some (perhaps meaningless) change in the output binary.
-We have therefore consider a mutant killed only when it causes semantically wrong output from a compiled program, a fairly complex specification-based notion of difference.
+We have therefore considered a mutant killed only when it causes semantically wrong output from a compiled program, a fairly complex specification-based notion of difference.
 Compilers may therefore be an unusually poor fit for the Metallaxis approach, which, as our results above show, is quite competitive for more ``normal'' localization, where output (vs.
